### 2020-11-20 plottex/a.tex
Add plottex.

\documentclass{standalone}
        \usepackage{tikz,pgfplots}

        \pgfplotsset{compat=1.10}

        \usepgfplotslibrary{fillbetween,external}
        \tikzexternalize

        \begin{document}

        \begin{tikzpicture}
            \begin{axis}[
                thick,smooth,no markers,
                grid=both,
                grid style={line width=.1pt, draw=gray!10},
                xlabel={Steps},
                ylabel={Train/Tot_Reward}]
            ]
            \addplot+[name path=A,black,line width=1pt] coordinates {(0, 0.0)(1, 0.0)(2, 0.0)(3, 0.0)(4, 0.0)(5, 0.0)(6, 0.0)(7, 0.0)(8, 0.0)(9, 0.0)(10, 0.0)(11, 0.0)(12, 0.0)(13, 0.0)(14, 0.0)(15, 0.0)(16, 0.0)(17, 0.0)(18, 0.0)(19, 0.0)(20, 0.0)(21, 0.0)(22, 0.0)(23, 0.0)(24, 0.0)(25, 0.0)(26, 0.0)(27, 0.0)(28, 0.0)(29, 0.0)(30, 0.0)(31, 0.0)(32, 0.0)(33, 0.0)(34, 0.0)(35, 0.0)(36, 0.0)(37, 0.0)(38, 0.0)(39, 0.0)(40, 0.0)(41, 0.0)(42, 0.0)(43, 0.0)(44, 0.0)(45, 0.0)(46, 0.0)(47, 0.0)(48, 0.0)(49, 0.0)(50, 0.0)(51, 0.0)(52, 0.0)(53, 0.0)(54, 0.0)(55, 0.0)(56, 0.0)(57, 0.0)(58, 0.0)(59, 0.0)(60, 0.0)(61, 0.0)(62, 0.0)(63, 0.0)(64, 0.0)(65, 0.0)(66, 0.0)(67, 0.0)(68, 0.0)(69, 0.0)(70, 0.0)(71, 0.0)(72, 0.0)(73, 0.0)(74, 0.0)(75, 0.0)(76, 0.0)(77, 0.0)(78, 0.0)(79, 0.0)(80, 0.0)(81, 0.0)(82, 0.0)(83, 0.0)(84, 0.0)(85, 0.0)(86, 0.0)(87, 0.0)(88, 0.0)(89, 0.0)(90, 0.0)(91, 0.0)(92, 0.0)(93, 0.0)(94, 0.0)(95, 0.0)(96, 0.0)(97, 0.0)(98, 0.0)(99, 0.0)(100, 0.0)(101, 0.0)(102, 0.0)(103, 0.0)(104, 0.0)(105, 0.0)(106, 0.0)(107, 0.0)(108, 0.0)(109, 0.0)(110, 0.0)(111, 0.0)(112, 0.0)(113, 0.0)(114, 0.0)(115, 0.0)(116, 0.0)(117, 0.0)(118, 0.0)(119, 0.0)(120, 0.0)(121, 0.0)(122, 0.0)(123, 0.0)(124, 0.0)(125, 0.0)(126, 0.0)(127, 0.0)(128, 0.0)(129, 0.0)(130, 0.0)(131, 0.0)(132, 0.0)(133, 0.0)(134, 0.0)(135, 0.0)(136, 0.0)(137, 0.0)(138, 0.0)(139, 0.0)(140, 0.0)(141, 0.0)(142, 0.0)(143, 0.0)(144, 0.0)(145, 0.0)(146, 0.0)(147, 0.0)(148, 0.0)(149, 0.0)(150, 0.0)(151, 0.0)(152, 0.0)(153, 0.0)(154, 0.0)(155, 0.0)(156, 0.0)(157, 0.0)(158, 0.0)(159, 0.0)(160, 0.0)(161, 0.0)(162, 0.0)(163, 0.0)(164, 0.0)(165, 0.0)(166, 0.0)(167, 0.0)(168, 0.0)(169, 0.0)(170, 0.0)(171, 0.0)(172, 0.0)(173, 0.0)(174, 0.0)(175, 0.0)(176, 0.0)(177, 0.0)(178, 0.0)(179, 0.0)(180, 0.0)(181, 0.0)(182, 0.0)(183, 0.0)(184, 0.0)(185, 0.0)(186, 0.0)(187, 0.0)(188, 0.0)(189, 0.0)(190, 0.0)(191, 0.0)(192, 0.0)(193, 0.0)(194, 0.0)(195, 0.0)(196, 0.0)(197, 0.0)(198, 0.0)(199, 0.0)(200, 0.0)(201, 0.0)(202, 0.0)(203, 0.0)(204, 0.0)(205, 0.0)(206, 0.0)(207, 0.0)(208, 0.0)(209, 0.0)(210, 0.0)(211, 0.0)(212, 0.0)(213, 0.0)(214, 0.0)(215, 0.0)(216, 0.0)(217, 0.0)(218, 0.0)(219, 0.0)(220, 0.0)(221, 0.0)(222, 0.0)(223, 0.0)(224, 0.0)(225, 0.0)(226, 0.0)(227, 0.0)(228, 0.0)(229, 0.0)(230, 0.0)(231, 0.0)(232, 0.0)(233, 0.0)(234, 0.0)(235, 0.0)(236, 0.0)(237, 0.0)(238, 0.0)(239, 0.0)(240, 0.0)(241, 0.0)(242, 0.0)(243, 0.0)(244, 0.0)(245, 0.0)(246, 0.0)(247, 0.0)(248, 0.0)(249, 0.0)(250, 0.0)(251, 0.0)(252, 0.0)(253, 0.0)(254, 0.0)(255, 0.0)(256, 0.0)(257, 0.0)(258, 0.0)(259, 0.0)(260, 0.0)(261, 0.0)(262, 0.0)(263, 0.0)(264, 0.0)(265, 0.0)(266, 0.0)(267, 0.0)(268, 0.0)(269, 0.0)(270, 0.0)(271, 0.0)(272, 0.0)(273, 0.0)(274, 0.0)(275, 0.0)(276, 0.0)(277, 0.0)(278, 0.0)(279, 0.0)(280, 0.0)(281, 0.0)(282, 0.0)(283, 0.0)(284, 0.0)(285, 0.0)(286, 0.0)(287, 0.0)(288, 0.0)(289, 0.0)(290, 0.0)(291, 0.0)(292, 0.0)(293, 0.0)(294, 0.0)(295, 0.0)(296, 0.0)(297, 0.0)(298, 0.0)(299, 0.0)(300, 0.0)(301, 0.0)(302, 0.0)(303, 0.0)(304, 0.0)(305, 0.0)(306, 0.0)(307, 0.0)(308, 0.0)(309, 0.0)(310, 0.0)(311, 0.0)(312, 0.0)(313, 0.0)(314, 0.0)(315, 0.0)(316, 0.0)(317, 0.0)(318, 0.0)(319, 0.0)(320, 0.0)(321, 0.0)(322, 0.0)(323, 0.0)(324, 0.0)(325, 0.0)(326, 0.0)(327, 0.0)(328, 0.0)(329, 0.0)(330, 0.0)(331, 0.0)(332, 0.0)(333, 0.0)(334, 0.0)(335, 0.0)(336, 0.0)(337, 0.0)(338, 0.0)(339, 0.0)(340, 0.0)(341, 0.0)(342, 0.0)(343, 0.0)(344, 0.0)(345, 0.0)(346, 0.0)(347, 0.0)(348, 0.0)(349, 0.0)(350, 0.0)(351, 0.0)(352, 0.0)(353, 0.0)(354, 0.0)(355, 0.0)(356, 0.0)(357, 0.0)(358, 0.0)(359, 0.0)(360, 0.0)(361, 0.0)(362, 0.0)(363, 0.0)(364, 0.0)(365, 0.0)(366, 0.0)(367, 0.0)(368, 0.0)(369, 0.0)(370, 0.0)(371, 0.0)(372, 0.0)(373, 0.0)(374, 0.0)(375, 0.0)(376, 0.0)(377, 0.0)(378, 0.0)(379, 0.0)(380, 0.0)(381, 0.0)(382, 0.0)(383, 0.0)(384, 0.0)(385, 0.0)(386, 0.0)(387, 0.0)(388, 0.0)(389, 0.0)(390, 0.0)(391, 0.0)(392, 0.0)(393, 0.0)(394, 0.0)(395, 0.0)(396, 0.0)(397, 0.0)(398, 0.0)(399, 0.0)(400, 0.0)(401, 0.0)(402, 0.0)(403, 0.0)(404, 0.0)(405, 0.0)(406, 0.0)(407, 0.0)(408, 0.0)(409, 0.0)(410, 0.0)(411, 0.0)(412, 0.0)(413, 0.0)(414, 0.0)(415, 0.0)(416, 0.0)(417, 0.0)(418, 0.0)(419, 0.0)(420, 0.0)(421, 0.0)(422, 0.0)(423, 0.0)(424, 0.0)(425, 0.0)(426, 0.0)(427, 0.0)(428, 0.0)(429, 0.0)(430, 0.0)(431, 0.0)(432, 0.0)(433, 0.0)(434, 0.0)(435, 0.0)(436, 0.0)(437, 0.0)(438, 0.0)(439, 0.0)(440, 0.0)(441, 0.0)(442, 0.0)(443, 0.0)(444, 0.0)(445, 0.0)(446, 0.0)(447, 0.0)(448, 0.0)(449, 0.0)(450, 0.0)(451, 0.0)(452, 0.0)(453, 0.0)(454, 0.0)(455, 0.0)(456, 0.0)(457, 0.0)(458, 0.0)(459, 0.0)(460, 0.0)(461, 0.0)(462, 0.0)(463, 0.0)(464, 0.0)(465, 0.0)(466, 0.0)(467, 0.0)(468, 0.0)(469, 0.0)(470, 0.0)(471, 0.0)(472, 0.0)(473, 0.0)(474, 0.0)(475, 0.0)(476, 0.0)(477, 0.0)(478, 0.0)(479, 0.0)(480, 0.0)(481, 0.0)(482, 0.0)(483, 0.0)(484, 0.0)(485, 0.0)(486, 0.0)(487, 0.0)(488, 0.0)(489, 0.0)(490, 0.0)(491, 0.0)(492, 0.0)(493, 0.0)(494, 0.0)(495, 0.0)(496, 0.0)(497, 0.0)(498, 0.0)(499, 0.0)(500, 0.0)(501, 0.0)(502, 0.0)(503, 0.0)(504, 0.0)(505, 0.0)(506, 0.0)(507, 0.0)(508, 0.0)(509, 0.0)(510, 0.0)(511, 0.0)(512, 0.0)(513, 0.0)(514, 0.0)(515, 0.0)(516, 0.0)(517, 0.0)(518, 0.0)(519, 0.0)(520, 0.0)(521, 0.0)(522, 0.0)(523, 0.0)(524, 0.0)(525, 0.0)(526, 0.0)(527, 0.0)(528, 0.0)(529, 0.0)(530, 0.0)(531, 0.0)(532, 0.0)(533, 0.0)(534, 0.0)(535, 0.0)(536, 0.0)(537, 0.0)(538, 0.0)(539, 0.0)(540, 0.0)(541, 0.0)(542, 0.0)(543, 0.0)(544, 0.0)(545, 0.0)(546, 0.0)(547, 0.0)(548, 0.0)(549, 0.0)(550, 0.0)(551, 0.0)(552, 0.0)(553, 0.0)(554, 0.0)(555, 0.0)(556, 0.0)(557, 0.0)(558, 0.0)(559, 0.0)(560, 0.0)(561, 0.0)(562, 0.0)(563, 0.0)(564, 0.0)(565, 0.0)(566, 0.0)(567, 0.0)(568, 0.0)(569, 0.0)(570, 0.0)(571, 0.0)(572, 0.0)(573, 0.0)(574, 0.0)(575, 0.0)(576, 0.0)(577, 0.0)(578, 0.0)(579, 0.0)(580, 0.0)(581, 0.0)(582, 0.0)(583, 0.0)(584, 0.0)(585, 0.0)(586, 0.0)(587, 0.0)(588, 0.0)(589, 0.0)(590, 0.0)(591, 0.0)(592, 0.0)(593, 0.0)(594, 0.0)(595, 0.0)(596, 0.0)(597, 0.0)(598, 0.0)(599, 0.0)(600, 0.0)(601, 0.0)(602, 0.0)(603, 0.0)(604, 0.0)(605, 0.0)(606, 0.0)(607, 0.0)(608, 0.0)(609, 0.0)(610, 0.0)(611, 0.0)(612, 0.0)(613, 0.0)(614, 0.0)(615, 0.0)(616, 0.0)(617, 0.0)(618, 0.0)(619, 0.0)(620, 0.0)(621, 0.0)(622, 0.0)(623, 0.0)(624, 0.0)(625, 0.0)(626, 0.0)(627, 0.0)(628, 0.0)(629, 0.0)(630, 0.0)(631, 0.0)(632, 0.0)(633, 0.0)(634, 0.0)(635, 0.0)(636, 0.0)(637, 0.0)(638, 0.0)(639, 0.0)(640, 0.0)(641, 0.0)(642, 0.0)(643, 0.0)(644, 0.0)(645, 0.0)(646, 0.0)(647, 0.0)(648, 0.0)(649, 0.0)(650, 0.0)(651, 0.0)(652, 0.0)(653, 0.0)(654, 0.0)(655, 0.0)(656, 0.0)(657, 0.0)(658, 0.0)(659, 0.0)(660, 0.0)(661, 0.0)(662, 0.0)(663, 0.0)(664, 0.0)(665, 0.0)(666, 0.0)(667, 0.0)(668, 0.0)(669, 0.0)(670, 0.0)(671, 0.0)(672, 0.0)(673, 0.0)(674, 0.0)(675, 0.0)(676, 0.0)(677, 0.0)(678, 0.0)(679, 0.0)(680, 0.0)(681, 0.0)(682, 0.0)(683, 0.0)(684, 0.0)(685, 0.0)(686, 0.0)(687, 0.0)(688, 0.0)(689, 0.0)(690, 0.0)(691, 0.0)(692, 0.0)(693, 0.0)(694, 0.0)(695, 0.0)(696, 0.0)(697, 0.0)(698, 0.0)(699, 0.0)(700, 0.0)(701, 0.0)(702, 0.0)(703, 0.0)(704, 0.0)(705, 0.0)(706, 0.0)(707, 0.0)(708, 0.0)(709, 0.0)(710, 0.0)(711, 0.0)(712, 0.0)(713, 0.0)(714, 0.0)(715, 0.0)(716, 0.0)(717, 0.0)(718, 0.0)(719, 0.0)(720, 0.0)(721, 0.0)(722, 0.0)(723, 0.0)(724, 0.0)(725, 0.0)(726, 0.0)(727, 0.0)(728, 0.0)(729, 0.0)(730, 0.0)(731, 0.0)(732, 0.0)(733, 0.0)(734, 0.0)(735, 0.0)(736, 0.0)(737, 0.0)(738, 0.0)(739, 0.0)(740, 0.0)(741, 0.0)(742, 0.0)(743, 0.0)(744, 0.0)(745, 0.0)(746, 0.0)(747, 0.0)(748, 0.0)(749, 0.0)(750, 0.0)(751, 0.0)(752, 0.0)(753, 0.0)(754, 0.0)(755, 0.0)(756, 0.0)(757, 0.0)(758, 0.0)(759, 0.0)(760, 0.0)(761, 0.0)(762, 0.0)(763, 0.0)(764, 0.0)(765, 0.0)(766, 0.0)(767, 0.0)(768, 0.0)(769, 0.0)(770, 0.0)(771, 0.0)(772, 0.0)(773, 0.0)(774, 0.0)(775, 0.0)(776, 0.0)(777, 0.0)(778, 0.0)(779, 0.0)(780, 0.0)(781, 0.0)(782, 0.0)(783, 0.0)(784, 0.0)(785, 0.0)(786, 0.0)(787, 0.0)(788, 0.0)(789, 0.0)(790, 0.0)(791, 0.0)(792, 0.0)(793, 0.0)(794, 0.0)(795, 0.0)(796, 0.0)(797, 0.0)(798, 0.0)(799, 0.0)(800, 0.0)(801, 0.0)(802, 0.0)(803, 0.0)(804, 0.0)(805, 0.0)(806, 0.0)(807, 0.0)(808, 0.0)(809, 0.0)(810, 0.0)(811, 0.0)(812, 0.0)(813, 0.0)(814, 0.0)(815, 0.0)(816, 0.0)(817, 0.0)(818, 0.0)(819, 0.0)(820, 0.0)(821, 0.0)(822, 0.0)(823, 0.0)(824, 0.0)(825, 0.0)(826, 0.0)(827, 0.0)(828, 0.0)(829, 0.0)(830, 0.0)(831, 0.0)(832, 0.0)(833, 0.0)(834, 0.0)(835, 0.0)(836, 0.0)(837, 0.0)(838, 0.0)(839, 0.0)(840, 0.0)(841, 0.0)(842, 0.0)(843, 0.0)(844, 0.0)(845, 0.0)(846, 0.0)(847, 0.0)(848, 0.0)(849, 0.0)(850, 0.0)(851, 0.0)(852, 0.0)(853, 0.0)(854, 0.0)(855, 0.0)(856, 0.0)(857, 0.0)(858, 0.0)(859, 0.0)(860, 0.0)(861, 0.0)(862, 0.0)(863, 0.0)(864, 0.0)(865, 0.0)(866, 0.0)(867, 0.0)(868, 0.0)(869, 0.0)(870, 0.0)(871, 0.0)(872, 0.0)(873, 0.0)(874, 0.0)(875, 0.0)(876, 0.0)(877, 0.0)(878, 0.0)(879, 0.0)(880, 0.0)(881, 0.0)(882, 0.0)(883, 0.0)(884, 0.0)(885, 0.0)(886, 0.0)(887, 0.0)(888, 0.0)(889, 0.0)(890, 0.0)(891, 0.0)(892, 0.0)(893, 0.0)(894, 0.0)(895, 0.0)(896, 0.0)(897, 0.0)(898, 0.0)(899, 0.0)(900, 0.0)(901, 0.0)(902, 0.0)(903, 0.0)(904, 0.0)(905, 0.0)(906, 0.0)(907, 0.0)(908, 0.0)(909, 0.0)(910, 0.0)(911, 0.0)(912, 0.0)(913, 0.0)(914, 0.0)(915, 0.0)(916, 0.0)(917, 0.0)(918, 0.0)(919, 0.0)(920, 0.0)(921, 0.0)(922, 0.0)(923, 0.0)(924, 0.0)(925, 0.0)(926, 0.0)(927, 0.0)(928, 0.0)(929, 0.0)(930, 0.0)(931, 0.0)(932, 0.0)(933, 0.0)(934, 0.0)(935, 0.0)(936, 0.0)(937, 0.0)(938, 0.0)(939, 0.0)(940, 0.0)(941, 0.0)(942, 0.0)(943, 0.0)(944, 0.0)(945, 0.0)(946, 0.0)(947, 0.0)(948, 0.0)(949, 0.0)(950, 0.0)(951, 0.0)(952, 0.0)(953, 0.0)(954, 0.0)(955, 0.0)(956, 0.0)(957, 0.0)(958, 0.0)(959, 0.0)(960, 0.0)(961, 0.0)(962, 0.0)(963, 0.0)(964, 0.0)(965, 0.0)(966, 0.0)(967, 0.0)(968, 0.0)(969, 0.0)(970, 0.0)(971, 0.0)(972, 0.0)(973, 0.0)(974, 0.0)(975, 0.0)(976, 0.0)(977, 0.0)(978, 0.0)(979, 0.0)(980, 0.0)(981, 0.0)(982, 0.0)(983, 0.0)(984, 0.0)(985, 0.0)(986, 0.0)(987, 0.0)(988, 0.0)(989, 0.0)(990, 0.0)(991, 0.0)(992, 0.0)(993, 0.0)(994, 0.0)(995, 0.0)(996, 0.0)(997, 0.0)(998, 0.0)(999, 0.0)(1000, 0.0)(1001, 0.0)(1002, 0.0)(1003, 0.0)(1004, 0.0)(1005, 0.0)(1006, 0.0)(1007, 0.0)(1008, 0.0)(1009, 0.0)(1010, 0.0)(1011, 0.0)(1012, 0.0)(1013, 0.0)(1014, 0.0)(1015, 0.0)(1016, 0.0)(1017, 0.0)(1018, 0.0)(1019, 0.0)(1020, 0.0)(1021, 0.0)(1022, 0.0)(1023, 0.0)(1024, 0.0)(1025, 0.0)(1026, 0.0)(1027, 0.0)(1028, 0.0)(1029, 0.0)(1030, 0.0)(1031, 0.0)(1032, 0.0)(1033, 0.0)(1034, 0.0)(1035, 0.0)(1036, 0.0)(1037, 0.0)(1038, 0.0)(1039, 0.0)(1040, 0.0)(1041, 0.0)(1042, 0.0)(1043, 0.0)(1044, 0.0)(1045, 0.0)(1046, 0.0)(1047, 0.0)(1048, 0.0)(1049, 0.0)(1050, 0.0)(1051, 0.0)(1052, 0.0)(1053, 0.0)(1054, 0.0)(1055, 0.0)(1056, 0.0)(1057, 0.0)(1058, 0.0)(1059, 0.0)(1060, 0.0)(1061, 0.0)(1062, 0.0)(1063, 0.0)(1064, 0.0)(1065, 0.0)(1066, 0.0)(1067, 0.0)(1068, 0.0)(1069, 0.0)(1070, 0.0)(1071, 0.0)(1072, 0.0)(1073, 0.0)(1074, 0.0)(1075, 0.0)(1076, 0.0)(1077, 0.0)(1078, 0.0)(1079, 0.0)(1080, 0.0)(1081, 0.0)(1082, 0.0)(1083, 0.0)(1084, 0.0)(1085, 0.0)(1086, 0.0)(1087, 0.0)(1088, 0.0)(1089, 0.0)(1090, 0.0)(1091, 0.0)(1092, 0.0)(1093, 0.0)(1094, 0.0)(1095, 0.0)(1096, 0.0)(1097, 0.0)(1098, 0.0)(1099, 0.0)(1100, 0.0)(1101, 0.0)(1102, 0.0)(1103, 0.0)(1104, 0.0)(1105, 0.0)(1106, 0.0)(1107, 0.0)(1108, 0.0)(1109, 0.0)(1110, 0.0)(1111, 0.0)(1112, 0.0)(1113, 0.0)(1114, 0.0)(1115, 0.0)(1116, 0.0)(1117, 0.0)(1118, 0.0)(1119, 0.0)(1120, 0.0)(1121, 0.0)(1122, 0.0)(1123, 0.0)(1124, 0.0)(1125, 0.0)(1126, 0.0)(1127, 0.0)(1128, 0.0)(1129, 0.0)(1130, 0.0)(1131, 0.0)(1132, 0.0)(1133, 0.0)(1134, 0.0)(1135, 0.0)(1136, 0.0)(1137, 0.0)(1138, 0.0)(1139, 0.0)(1140, 0.0)(1141, 0.0)(1142, 0.0)(1143, 0.0)(1144, 0.0)(1145, 0.0)(1146, 0.0)(1147, 0.0)(1148, 0.0)(1149, 0.0)(1150, 0.0)(1151, 0.0)(1152, 0.0)(1153, 0.0)(1154, 0.0)(1155, 0.0)(1156, 0.0)(1157, 0.0)(1158, 0.0)(1159, 0.0)(1160, 0.0)(1161, 0.0)(1162, 0.0)(1163, 0.0)(1164, 0.0)(1165, 0.0)(1166, 0.0)(1167, 0.0)(1168, 0.0)(1169, 0.0)(1170, 0.0)(1171, 0.0)(1172, 0.0)(1173, 0.0)(1174, 0.0)(1175, 0.0)(1176, 0.0)(1177, 0.0)(1178, 0.0)(1179, 0.0)(1180, 0.0)(1181, 0.0)(1182, 0.0)(1183, 0.0)(1184, 0.0)(1185, 0.0)(1186, 0.0)(1187, 0.0)(1188, 0.0)(1189, 0.0)(1190, 0.0)(1191, 0.0)(1192, 0.0)(1193, 0.0)(1194, 0.0)(1195, 0.0)(1196, 0.0)(1197, 0.0)(1198, 0.0)(1199, 0.0)(1200, 0.0)(1201, 0.0)(1202, 0.0)(1203, 0.0)(1204, 0.0)(1205, 0.0)(1206, 0.0)(1207, 0.0)(1208, 0.0)(1209, 0.0)(1210, 0.0)(1211, 0.0)(1212, 0.0)(1213, 0.0)(1214, 0.0)(1215, 0.0)(1216, 0.0)(1217, 0.0)(1218, 0.0)(1219, 0.0)(1220, 0.0)(1221, 0.0)(1222, 0.0)(1223, 0.0)(1224, 0.0)(1225, 0.0)(1226, 0.0)(1227, 0.0)(1228, 0.0)(1229, 0.0)(1230, 0.0)(1231, 0.0)(1232, 0.0)(1233, 0.0)(1234, 0.0)(1235, 0.0)(1236, 0.0)(1237, 0.0)(1238, 0.0)(1239, 0.0)(1240, 0.0)(1241, 0.0)(1242, 0.0)(1243, 0.0)(1244, 0.0)(1245, 0.0)(1246, 0.0)(1247, 0.0)(1248, 0.0)(1249, 0.0)(1250, 0.0)(1251, 0.0)(1252, 0.0)(1253, 0.0)(1254, 0.0)(1255, 0.0)(1256, 0.0)(1257, 0.0)(1258, 0.0)(1259, 0.0)(1260, 0.0)(1261, 0.0)(1262, 0.0)(1263, 0.0)(1264, 0.0)(1265, 0.0)(1266, 0.0)(1267, 0.0)(1268, 0.0)(1269, 0.0)(1270, 0.0)(1271, 0.0)(1272, 0.0)(1273, 0.0)(1274, 0.0)(1275, 0.0)(1276, 0.0)(1277, 0.0)(1278, 0.0)(1279, 0.0)(1280, 0.0)(1281, 0.0)(1282, 0.0)(1283, 0.0)(1284, 0.0)(1285, 0.0)(1286, 0.0)(1287, 0.0)(1288, 0.0)(1289, 0.0)(1290, 0.0)(1291, 0.0)(1292, 0.0)(1293, 0.0)(1294, 0.0)(1295, 0.0)(1296, 0.0)(1297, 0.0)(1298, 0.0)(1299, 0.0)(1300, 0.0)(1301, 0.0)(1302, 0.0)(1303, 0.0)(1304, 0.0)(1305, 0.0)(1306, 0.0)(1307, 0.0)(1308, 0.0)(1309, 0.0)(1310, 0.0)(1311, 0.0)(1312, 0.0)(1313, 0.0)(1314, 0.0)(1315, 0.0)(1316, 0.0)(1317, 0.0)(1318, 0.0)(1319, 0.0)(1320, 0.0)(1321, 0.0)(1322, 0.0)(1323, 0.0)(1324, 0.0)(1325, 0.0)(1326, 0.0)(1327, 0.0)(1328, 0.0)(1329, 0.0)(1330, 0.0)(1331, 0.0)(1332, 0.0)(1333, 0.0)(1334, 0.0)(1335, 0.0)(1336, 0.0)(1337, 0.0)(1338, 0.0)(1339, 0.0)(1340, 0.0)(1341, 0.0)(1342, 0.0)(1343, 0.0)(1344, 0.0)(1345, 0.0)(1346, 0.0)(1347, 0.0)(1348, 0.0)(1349, 0.0)(1350, 0.0)(1351, 0.0)(1352, 0.0)(1353, 0.0)(1354, 0.0)(1355, 0.0)(1356, 0.0)(1357, 0.0)(1358, 0.0)(1359, 0.0)(1360, 0.0)(1361, 0.0)(1362, 0.0)(1363, 0.0)(1364, 0.0)(1365, 0.0)(1366, 0.0)(1367, 0.0)(1368, 0.0)(1369, 0.0)(1370, 0.0)(1371, 0.0)(1372, 0.0)(1373, 0.0)(1374, 0.0)(1375, 0.0)(1376, 0.0)(1377, 0.0)(1378, 0.0)(1379, 0.0)(1380, 0.0)(1381, 0.0)(1382, 0.0)(1383, 0.0)(1384, 0.0)(1385, 0.0)(1386, 0.0)(1387, 0.0)(1388, 0.0)(1389, 0.0)(1390, 0.0)(1391, 0.0)(1392, 0.0)(1393, 0.0)(1394, 0.0)(1395, 0.0)(1396, 0.0)(1397, 0.0)(1398, 0.0)(1399, 0.0)(1400, 0.0)(1401, 0.0)(1402, 0.0)(1403, 0.0)(1404, 0.0)(1405, 0.0)(1406, 0.0)(1407, 0.0)(1408, 0.0)(1409, 0.0)(1410, 0.0)(1411, 0.0)(1412, 0.0)(1413, 0.0)(1414, 0.0)(1415, 0.0)(1416, 0.0)(1417, 0.0)(1418, 0.0)(1419, 0.0)(1420, 0.0)(1421, 0.0)(1422, 0.0)(1423, 0.0)(1424, 0.0)(1425, 0.0)(1426, 0.0)(1427, 0.0)(1428, 0.0)(1429, 0.0)(1430, 0.0)(1431, 0.0)(1432, 0.0)(1433, 0.0)(1434, 0.0)(1435, 0.0)(1436, 0.0)(1437, 0.0)(1438, 0.0)(1439, 0.0)(1440, 0.0)(1441, 0.0)(1442, 0.0)(1443, 0.0)(1444, 0.0)(1445, 0.0)(1446, 0.0)(1447, 0.0)(1448, 0.0)(1449, 0.0)(1450, 0.0)(1451, 0.0)(1452, 0.0)(1453, 0.0)(1454, 0.0)(1455, 0.0)(1456, 0.0)(1457, 0.0)(1458, 0.0)(1459, 0.0)(1460, 0.0)(1461, 0.0)(1462, 0.0)(1463, 0.0)(1464, 0.0)(1465, 0.0)(1466, 0.0)(1467, 0.0)(1468, 0.0)(1469, 0.0)(1470, 0.0)(1471, 0.0)(1472, 0.0)(1473, 0.0)(1474, 0.0)(1475, 0.0)(1476, 0.0)(1477, 0.0)(1478, 0.0)(1479, 0.0)(1480, 0.0)(1481, 0.0)(1482, 0.0)(1483, 0.0)(1484, 0.0)(1485, 0.0)(1486, 0.0)(1487, 0.0)(1488, 0.0)(1489, 0.0)(1490, 0.0)(1491, 0.0)(1492, 0.0)(1493, 0.0)(1494, 0.0)(1495, 0.0)(1496, 0.0)(1497, 0.0)(1498, 0.0)(1499, 0.0)(1500, 0.0)(1501, 0.0)(1502, 0.0)(1503, 0.0)(1504, 0.0)(1505, 0.0)(1506, 0.0)(1507, 0.0)(1508, 0.0)(1509, 0.0)(1510, 0.0)(1511, 0.0)(1512, 0.0)(1513, 0.0)(1514, 0.0)(1515, 0.0)(1516, 0.0)(1517, 0.0)(1518, 0.0)(1519, 0.0)};
            \addplot+[name path=B,black,line width=.1pt] coordinates {(0, 0.0)(1, 0.0)(2, 0.0)(3, 0.0)(4, 0.0)(5, 0.0)(6, 0.0)(7, 0.0)(8, 0.0)(9, 0.0)(10, 0.0)(11, 0.0)(12, 0.0)(13, 0.0)(14, 0.0)(15, 0.0)(16, 0.0)(17, 0.0)(18, 0.0)(19, 0.0)(20, 0.0)(21, 0.0)(22, 0.0)(23, 0.0)(24, 0.0)(25, 0.0)(26, 0.0)(27, 0.0)(28, 0.0)(29, 0.0)(30, 0.0)(31, 0.0)(32, 0.0)(33, 0.0)(34, 0.0)(35, 0.0)(36, 0.0)(37, 0.0)(38, 0.0)(39, 0.0)(40, 0.0)(41, 0.0)(42, 0.0)(43, 0.0)(44, 0.0)(45, 0.0)(46, 0.0)(47, 0.0)(48, 0.0)(49, 0.0)(50, 0.0)(51, 0.0)(52, 0.0)(53, 0.0)(54, 0.0)(55, 0.0)(56, 0.0)(57, 0.0)(58, 0.0)(59, 0.0)(60, 0.0)(61, 0.0)(62, 0.0)(63, 0.0)(64, 0.0)(65, 0.0)(66, 0.0)(67, 0.0)(68, 0.0)(69, 0.0)(70, 0.0)(71, 0.0)(72, 0.0)(73, 0.0)(74, 0.0)(75, 0.0)(76, 0.0)(77, 0.0)(78, 0.0)(79, 0.0)(80, 0.0)(81, 0.0)(82, 0.0)(83, 0.0)(84, 0.0)(85, 0.0)(86, 0.0)(87, 0.0)(88, 0.0)(89, 0.0)(90, 0.0)(91, 0.0)(92, 0.0)(93, 0.0)(94, 0.0)(95, 0.0)(96, 0.0)(97, 0.0)(98, 0.0)(99, 0.0)(100, 0.0)(101, 0.0)(102, 0.0)(103, 0.0)(104, 0.0)(105, 0.0)(106, 0.0)(107, 0.0)(108, 0.0)(109, 0.0)(110, 0.0)(111, 0.0)(112, 0.0)(113, 0.0)(114, 0.0)(115, 0.0)(116, 0.0)(117, 0.0)(118, 0.0)(119, 0.0)(120, 0.0)(121, 0.0)(122, 0.0)(123, 0.0)(124, 0.0)(125, 0.0)(126, 0.0)(127, 0.0)(128, 0.0)(129, 0.0)(130, 0.0)(131, 0.0)(132, 0.0)(133, 0.0)(134, 0.0)(135, 0.0)(136, 0.0)(137, 0.0)(138, 0.0)(139, 0.0)(140, 0.0)(141, 0.0)(142, 0.0)(143, 0.0)(144, 0.0)(145, 0.0)(146, 0.0)(147, 0.0)(148, 0.0)(149, 0.0)(150, 0.0)(151, 0.0)(152, 0.0)(153, 0.0)(154, 0.0)(155, 0.0)(156, 0.0)(157, 0.0)(158, 0.0)(159, 0.0)(160, 0.0)(161, 0.0)(162, 0.0)(163, 0.0)(164, 0.0)(165, 0.0)(166, 0.0)(167, 0.0)(168, 0.0)(169, 0.0)(170, 0.0)(171, 0.0)(172, 0.0)(173, 0.0)(174, 0.0)(175, 0.0)(176, 0.0)(177, 0.0)(178, 0.0)(179, 0.0)(180, 0.0)(181, 0.0)(182, 0.0)(183, 0.0)(184, 0.0)(185, 0.0)(186, 0.0)(187, 0.0)(188, 0.0)(189, 0.0)(190, 0.0)(191, 0.0)(192, 0.0)(193, 0.0)(194, 0.0)(195, 0.0)(196, 0.0)(197, 0.0)(198, 0.0)(199, 0.0)(200, 0.0)(201, 0.0)(202, 0.0)(203, 0.0)(204, 0.0)(205, 0.0)(206, 0.0)(207, 0.0)(208, 0.0)(209, 0.0)(210, 0.0)(211, 0.0)(212, 0.0)(213, 0.0)(214, 0.0)(215, 0.0)(216, 0.0)(217, 0.0)(218, 0.0)(219, 0.0)(220, 0.0)(221, 0.0)(222, 0.0)(223, 0.0)(224, 0.0)(225, 0.0)(226, 0.0)(227, 0.0)(228, 0.0)(229, 0.0)(230, 0.0)(231, 0.0)(232, 0.0)(233, 0.0)(234, 0.0)(235, 0.0)(236, 0.0)(237, 0.0)(238, 0.0)(239, 0.0)(240, 0.0)(241, 0.0)(242, 0.0)(243, 0.0)(244, 0.0)(245, 0.0)(246, 0.0)(247, 0.0)(248, 0.0)(249, 0.0)(250, 0.0)(251, 0.0)(252, 0.0)(253, 0.0)(254, 0.0)(255, 0.0)(256, 0.0)(257, 0.0)(258, 0.0)(259, 0.0)(260, 0.0)(261, 0.0)(262, 0.0)(263, 0.0)(264, 0.0)(265, 0.0)(266, 0.0)(267, 0.0)(268, 0.0)(269, 0.0)(270, 0.0)(271, 0.0)(272, 0.0)(273, 0.0)(274, 0.0)(275, 0.0)(276, 0.0)(277, 0.0)(278, 0.0)(279, 0.0)(280, 0.0)(281, 0.0)(282, 0.0)(283, 0.0)(284, 0.0)(285, 0.0)(286, 0.0)(287, 0.0)(288, 0.0)(289, 0.0)(290, 0.0)(291, 0.0)(292, 0.0)(293, 0.0)(294, 0.0)(295, 0.0)(296, 0.0)(297, 0.0)(298, 0.0)(299, 0.0)(300, 0.0)(301, 0.0)(302, 0.0)(303, 0.0)(304, 0.0)(305, 0.0)(306, 0.0)(307, 0.0)(308, 0.0)(309, 0.0)(310, 0.0)(311, 0.0)(312, 0.0)(313, 0.0)(314, 0.0)(315, 0.0)(316, 0.0)(317, 0.0)(318, 0.0)(319, 0.0)(320, 0.0)(321, 0.0)(322, 0.0)(323, 0.0)(324, 0.0)(325, 0.0)(326, 0.0)(327, 0.0)(328, 0.0)(329, 0.0)(330, 0.0)(331, 0.0)(332, 0.0)(333, 0.0)(334, 0.0)(335, 0.0)(336, 0.0)(337, 0.0)(338, 0.0)(339, 0.0)(340, 0.0)(341, 0.0)(342, 0.0)(343, 0.0)(344, 0.0)(345, 0.0)(346, 0.0)(347, 0.0)(348, 0.0)(349, 0.0)(350, 0.0)(351, 0.0)(352, 0.0)(353, 0.0)(354, 0.0)(355, 0.0)(356, 0.0)(357, 0.0)(358, 0.0)(359, 0.0)(360, 0.0)(361, 0.0)(362, 0.0)(363, 0.0)(364, 0.0)(365, 0.0)(366, 0.0)(367, 0.0)(368, 0.0)(369, 0.0)(370, 0.0)(371, 0.0)(372, 0.0)(373, 0.0)(374, 0.0)(375, 0.0)(376, 0.0)(377, 0.0)(378, 0.0)(379, 0.0)(380, 0.0)(381, 0.0)(382, 0.0)(383, 0.0)(384, 0.0)(385, 0.0)(386, 0.0)(387, 0.0)(388, 0.0)(389, 0.0)(390, 0.0)(391, 0.0)(392, 0.0)(393, 0.0)(394, 0.0)(395, 0.0)(396, 0.0)(397, 0.0)(398, 0.0)(399, 0.0)(400, 0.0)(401, 0.0)(402, 0.0)(403, 0.0)(404, 0.0)(405, 0.0)(406, 0.0)(407, 0.0)(408, 0.0)(409, 0.0)(410, 0.0)(411, 0.0)(412, 0.0)(413, 0.0)(414, 0.0)(415, 0.0)(416, 0.0)(417, 0.0)(418, 0.0)(419, 0.0)(420, 0.0)(421, 0.0)(422, 0.0)(423, 0.0)(424, 0.0)(425, 0.0)(426, 0.0)(427, 0.0)(428, 0.0)(429, 0.0)(430, 0.0)(431, 0.0)(432, 0.0)(433, 0.0)(434, 0.0)(435, 0.0)(436, 0.0)(437, 0.0)(438, 0.0)(439, 0.0)(440, 0.0)(441, 0.0)(442, 0.0)(443, 0.0)(444, 0.0)(445, 0.0)(446, 0.0)(447, 0.0)(448, 0.0)(449, 0.0)(450, 0.0)(451, 0.0)(452, 0.0)(453, 0.0)(454, 0.0)(455, 0.0)(456, 0.0)(457, 0.0)(458, 0.0)(459, 0.0)(460, 0.0)(461, 0.0)(462, 0.0)(463, 0.0)(464, 0.0)(465, 0.0)(466, 0.0)(467, 0.0)(468, 0.0)(469, 0.0)(470, 0.0)(471, 0.0)(472, 0.0)(473, 0.0)(474, 0.0)(475, 0.0)(476, 0.0)(477, 0.0)(478, 0.0)(479, 0.0)(480, 0.0)(481, 0.0)(482, 0.0)(483, 0.0)(484, 0.0)(485, 0.0)(486, 0.0)(487, 0.0)(488, 0.0)(489, 0.0)(490, 0.0)(491, 0.0)(492, 0.0)(493, 0.0)(494, 0.0)(495, 0.0)(496, 0.0)(497, 0.0)(498, 0.0)(499, 0.0)(500, 0.0)(501, 0.0)(502, 0.0)(503, 0.0)(504, 0.0)(505, 0.0)(506, 0.0)(507, 0.0)(508, 0.0)(509, 0.0)(510, 0.0)(511, 0.0)(512, 0.0)(513, 0.0)(514, 0.0)(515, 0.0)(516, 0.0)(517, 0.0)(518, 0.0)(519, 0.0)(520, 0.0)(521, 0.0)(522, 0.0)(523, 0.0)(524, 0.0)(525, 0.0)(526, 0.0)(527, 0.0)(528, 0.0)(529, 0.0)(530, 0.0)(531, 0.0)(532, 0.0)(533, 0.0)(534, 0.0)(535, 0.0)(536, 0.0)(537, 0.0)(538, 0.0)(539, 0.0)(540, 0.0)(541, 0.0)(542, 0.0)(543, 0.0)(544, 0.0)(545, 0.0)(546, 0.0)(547, 0.0)(548, 0.0)(549, 0.0)(550, 0.0)(551, 0.0)(552, 0.0)(553, 0.0)(554, 0.0)(555, 0.0)(556, 0.0)(557, 0.0)(558, 0.0)(559, 0.0)(560, 0.0)(561, 0.0)(562, 0.0)(563, 0.0)(564, 0.0)(565, 0.0)(566, 0.0)(567, 0.0)(568, 0.0)(569, 0.0)(570, 0.0)(571, 0.0)(572, 0.0)(573, 0.0)(574, 0.0)(575, 0.0)(576, 0.0)(577, 0.0)(578, 0.0)(579, 0.0)(580, 0.0)(581, 0.0)(582, 0.0)(583, 0.0)(584, 0.0)(585, 0.0)(586, 0.0)(587, 0.0)(588, 0.0)(589, 0.0)(590, 0.0)(591, 0.0)(592, 0.0)(593, 0.0)(594, 0.0)(595, 0.0)(596, 0.0)(597, 0.0)(598, 0.0)(599, 0.0)(600, 0.0)(601, 0.0)(602, 0.0)(603, 0.0)(604, 0.0)(605, 0.0)(606, 0.0)(607, 0.0)(608, 0.0)(609, 0.0)(610, 0.0)(611, 0.0)(612, 0.0)(613, 0.0)(614, 0.0)(615, 0.0)(616, 0.0)(617, 0.0)(618, 0.0)(619, 0.0)(620, 0.0)(621, 0.0)(622, 0.0)(623, 0.0)(624, 0.0)(625, 0.0)(626, 0.0)(627, 0.0)(628, 0.0)(629, 0.0)(630, 0.0)(631, 0.0)(632, 0.0)(633, 0.0)(634, 0.0)(635, 0.0)(636, 0.0)(637, 0.0)(638, 0.0)(639, 0.0)(640, 0.0)(641, 0.0)(642, 0.0)(643, 0.0)(644, 0.0)(645, 0.0)(646, 0.0)(647, 0.0)(648, 0.0)(649, 0.0)(650, 0.0)(651, 0.0)(652, 0.0)(653, 0.0)(654, 0.0)(655, 0.0)(656, 0.0)(657, 0.0)(658, 0.0)(659, 0.0)(660, 0.0)(661, 0.0)(662, 0.0)(663, 0.0)(664, 0.0)(665, 0.0)(666, 0.0)(667, 0.0)(668, 0.0)(669, 0.0)(670, 0.0)(671, 0.0)(672, 0.0)(673, 0.0)(674, 0.0)(675, 0.0)(676, 0.0)(677, 0.0)(678, 0.0)(679, 0.0)(680, 0.0)(681, 0.0)(682, 0.0)(683, 0.0)(684, 0.0)(685, 0.0)(686, 0.0)(687, 0.0)(688, 0.0)(689, 0.0)(690, 0.0)(691, 0.0)(692, 0.0)(693, 0.0)(694, 0.0)(695, 0.0)(696, 0.0)(697, 0.0)(698, 0.0)(699, 0.0)(700, 0.0)(701, 0.0)(702, 0.0)(703, 0.0)(704, 0.0)(705, 0.0)(706, 0.0)(707, 0.0)(708, 0.0)(709, 0.0)(710, 0.0)(711, 0.0)(712, 0.0)(713, 0.0)(714, 0.0)(715, 0.0)(716, 0.0)(717, 0.0)(718, 0.0)(719, 0.0)(720, 0.0)(721, 0.0)(722, 0.0)(723, 0.0)(724, 0.0)(725, 0.0)(726, 0.0)(727, 0.0)(728, 0.0)(729, 0.0)(730, 0.0)(731, 0.0)(732, 0.0)(733, 0.0)(734, 0.0)(735, 0.0)(736, 0.0)(737, 0.0)(738, 0.0)(739, 0.0)(740, 0.0)(741, 0.0)(742, 0.0)(743, 0.0)(744, 0.0)(745, 0.0)(746, 0.0)(747, 0.0)(748, 0.0)(749, 0.0)(750, 0.0)(751, 0.0)(752, 0.0)(753, 0.0)(754, 0.0)(755, 0.0)(756, 0.0)(757, 0.0)(758, 0.0)(759, 0.0)(760, 0.0)(761, 0.0)(762, 0.0)(763, 0.0)(764, 0.0)(765, 0.0)(766, 0.0)(767, 0.0)(768, 0.0)(769, 0.0)(770, 0.0)(771, 0.0)(772, 0.0)(773, 0.0)(774, 0.0)(775, 0.0)(776, 0.0)(777, 0.0)(778, 0.0)(779, 0.0)(780, 0.0)(781, 0.0)(782, 0.0)(783, 0.0)(784, 0.0)(785, 0.0)(786, 0.0)(787, 0.0)(788, 0.0)(789, 0.0)(790, 0.0)(791, 0.0)(792, 0.0)(793, 0.0)(794, 0.0)(795, 0.0)(796, 0.0)(797, 0.0)(798, 0.0)(799, 0.0)(800, 0.0)(801, 0.0)(802, 0.0)(803, 0.0)(804, 0.0)(805, 0.0)(806, 0.0)(807, 0.0)(808, 0.0)(809, 0.0)(810, 0.0)(811, 0.0)(812, 0.0)(813, 0.0)(814, 0.0)(815, 0.0)(816, 0.0)(817, 0.0)(818, 0.0)(819, 0.0)(820, 0.0)(821, 0.0)(822, 0.0)(823, 0.0)(824, 0.0)(825, 0.0)(826, 0.0)(827, 0.0)(828, 0.0)(829, 0.0)(830, 0.0)(831, 0.0)(832, 0.0)(833, 0.0)(834, 0.0)(835, 0.0)(836, 0.0)(837, 0.0)(838, 0.0)(839, 0.0)(840, 0.0)(841, 0.0)(842, 0.0)(843, 0.0)(844, 0.0)(845, 0.0)(846, 0.0)(847, 0.0)(848, 0.0)(849, 0.0)(850, 0.0)(851, 0.0)(852, 0.0)(853, 0.0)(854, 0.0)(855, 0.0)(856, 0.0)(857, 0.0)(858, 0.0)(859, 0.0)(860, 0.0)(861, 0.0)(862, 0.0)(863, 0.0)(864, 0.0)(865, 0.0)(866, 0.0)(867, 0.0)(868, 0.0)(869, 0.0)(870, 0.0)(871, 0.0)(872, 0.0)(873, 0.0)(874, 0.0)(875, 0.0)(876, 0.0)(877, 0.0)(878, 0.0)(879, 0.0)(880, 0.0)(881, 0.0)(882, 0.0)(883, 0.0)(884, 0.0)(885, 0.0)(886, 0.0)(887, 0.0)(888, 0.0)(889, 0.0)(890, 0.0)(891, 0.0)(892, 0.0)(893, 0.0)(894, 0.0)(895, 0.0)(896, 0.0)(897, 0.0)(898, 0.0)(899, 0.0)(900, 0.0)(901, 0.0)(902, 0.0)(903, 0.0)(904, 0.0)(905, 0.0)(906, 0.0)(907, 0.0)(908, 0.0)(909, 0.0)(910, 0.0)(911, 0.0)(912, 0.0)(913, 0.0)(914, 0.0)(915, 0.0)(916, 0.0)(917, 0.0)(918, 0.0)(919, 0.0)(920, 0.0)(921, 0.0)(922, 0.0)(923, 0.0)(924, 0.0)(925, 0.0)(926, 0.0)(927, 0.0)(928, 0.0)(929, 0.0)(930, 0.0)(931, 0.0)(932, 0.0)(933, 0.0)(934, 0.0)(935, 0.0)(936, 0.0)(937, 0.0)(938, 0.0)(939, 0.0)(940, 0.0)(941, 0.0)(942, 0.0)(943, 0.0)(944, 0.0)(945, 0.0)(946, 0.0)(947, 0.0)(948, 0.0)(949, 0.0)(950, 0.0)(951, 0.0)(952, 0.0)(953, 0.0)(954, 0.0)(955, 0.0)(956, 0.0)(957, 0.0)(958, 0.0)(959, 0.0)(960, 0.0)(961, 0.0)(962, 0.0)(963, 0.0)(964, 0.0)(965, 0.0)(966, 0.0)(967, 0.0)(968, 0.0)(969, 0.0)(970, 0.0)(971, 0.0)(972, 0.0)(973, 0.0)(974, 0.0)(975, 0.0)(976, 0.0)(977, 0.0)(978, 0.0)(979, 0.0)(980, 0.0)(981, 0.0)(982, 0.0)(983, 0.0)(984, 0.0)(985, 0.0)(986, 0.0)(987, 0.0)(988, 0.0)(989, 0.0)(990, 0.0)(991, 0.0)(992, 0.0)(993, 0.0)(994, 0.0)(995, 0.0)(996, 0.0)(997, 0.0)(998, 0.0)(999, 0.0)(1000, 0.0)(1001, 0.0)(1002, 0.0)(1003, 0.0)(1004, 0.0)(1005, 0.0)(1006, 0.0)(1007, 0.0)(1008, 0.0)(1009, 0.0)(1010, 0.0)(1011, 0.0)(1012, 0.0)(1013, 0.0)(1014, 0.0)(1015, 0.0)(1016, 0.0)(1017, 0.0)(1018, 0.0)(1019, 0.0)(1020, 0.0)(1021, 0.0)(1022, 0.0)(1023, 0.0)(1024, 0.0)(1025, 0.0)(1026, 0.0)(1027, 0.0)(1028, 0.0)(1029, 0.0)(1030, 0.0)(1031, 0.0)(1032, 0.0)(1033, 0.0)(1034, 0.0)(1035, 0.0)(1036, 0.0)(1037, 0.0)(1038, 0.0)(1039, 0.0)(1040, 0.0)(1041, 0.0)(1042, 0.0)(1043, 0.0)(1044, 0.0)(1045, 0.0)(1046, 0.0)(1047, 0.0)(1048, 0.0)(1049, 0.0)(1050, 0.0)(1051, 0.0)(1052, 0.0)(1053, 0.0)(1054, 0.0)(1055, 0.0)(1056, 0.0)(1057, 0.0)(1058, 0.0)(1059, 0.0)(1060, 0.0)(1061, 0.0)(1062, 0.0)(1063, 0.0)(1064, 0.0)(1065, 0.0)(1066, 0.0)(1067, 0.0)(1068, 0.0)(1069, 0.0)(1070, 0.0)(1071, 0.0)(1072, 0.0)(1073, 0.0)(1074, 0.0)(1075, 0.0)(1076, 0.0)(1077, 0.0)(1078, 0.0)(1079, 0.0)(1080, 0.0)(1081, 0.0)(1082, 0.0)(1083, 0.0)(1084, 0.0)(1085, 0.0)(1086, 0.0)(1087, 0.0)(1088, 0.0)(1089, 0.0)(1090, 0.0)(1091, 0.0)(1092, 0.0)(1093, 0.0)(1094, 0.0)(1095, 0.0)(1096, 0.0)(1097, 0.0)(1098, 0.0)(1099, 0.0)(1100, 0.0)(1101, 0.0)(1102, 0.0)(1103, 0.0)(1104, 0.0)(1105, 0.0)(1106, 0.0)(1107, 0.0)(1108, 0.0)(1109, 0.0)(1110, 0.0)(1111, 0.0)(1112, 0.0)(1113, 0.0)(1114, 0.0)(1115, 0.0)(1116, 0.0)(1117, 0.0)(1118, 0.0)(1119, 0.0)(1120, 0.0)(1121, 0.0)(1122, 0.0)(1123, 0.0)(1124, 0.0)(1125, 0.0)(1126, 0.0)(1127, 0.0)(1128, 0.0)(1129, 0.0)(1130, 0.0)(1131, 0.0)(1132, 0.0)(1133, 0.0)(1134, 0.0)(1135, 0.0)(1136, 0.0)(1137, 0.0)(1138, 0.0)(1139, 0.0)(1140, 0.0)(1141, 0.0)(1142, 0.0)(1143, 0.0)(1144, 0.0)(1145, 0.0)(1146, 0.0)(1147, 0.0)(1148, 0.0)(1149, 0.0)(1150, 0.0)(1151, 0.0)(1152, 0.0)(1153, 0.0)(1154, 0.0)(1155, 0.0)(1156, 0.0)(1157, 0.0)(1158, 0.0)(1159, 0.0)(1160, 0.0)(1161, 0.0)(1162, 0.0)(1163, 0.0)(1164, 0.0)(1165, 0.0)(1166, 0.0)(1167, 0.0)(1168, 0.0)(1169, 0.0)(1170, 0.0)(1171, 0.0)(1172, 0.0)(1173, 0.0)(1174, 0.0)(1175, 0.0)(1176, 0.0)(1177, 0.0)(1178, 0.0)(1179, 0.0)(1180, 0.0)(1181, 0.0)(1182, 0.0)(1183, 0.0)(1184, 0.0)(1185, 0.0)(1186, 0.0)(1187, 0.0)(1188, 0.0)(1189, 0.0)(1190, 0.0)(1191, 0.0)(1192, 0.0)(1193, 0.0)(1194, 0.0)(1195, 0.0)(1196, 0.0)(1197, 0.0)(1198, 0.0)(1199, 0.0)(1200, 0.0)(1201, 0.0)(1202, 0.0)(1203, 0.0)(1204, 0.0)(1205, 0.0)(1206, 0.0)(1207, 0.0)(1208, 0.0)(1209, 0.0)(1210, 0.0)(1211, 0.0)(1212, 0.0)(1213, 0.0)(1214, 0.0)(1215, 0.0)(1216, 0.0)(1217, 0.0)(1218, 0.0)(1219, 0.0)(1220, 0.0)(1221, 0.0)(1222, 0.0)(1223, 0.0)(1224, 0.0)(1225, 0.0)(1226, 0.0)(1227, 0.0)(1228, 0.0)(1229, 0.0)(1230, 0.0)(1231, 0.0)(1232, 0.0)(1233, 0.0)(1234, 0.0)(1235, 0.0)(1236, 0.0)(1237, 0.0)(1238, 0.0)(1239, 0.0)(1240, 0.0)(1241, 0.0)(1242, 0.0)(1243, 0.0)(1244, 0.0)(1245, 0.0)(1246, 0.0)(1247, 0.0)(1248, 0.0)(1249, 0.0)(1250, 0.0)(1251, 0.0)(1252, 0.0)(1253, 0.0)(1254, 0.0)(1255, 0.0)(1256, 0.0)(1257, 0.0)(1258, 0.0)(1259, 0.0)(1260, 0.0)(1261, 0.0)(1262, 0.0)(1263, 0.0)(1264, 0.0)(1265, 0.0)(1266, 0.0)(1267, 0.0)(1268, 0.0)(1269, 0.0)(1270, 0.0)(1271, 0.0)(1272, 0.0)(1273, 0.0)(1274, 0.0)(1275, 0.0)(1276, 0.0)(1277, 0.0)(1278, 0.0)(1279, 0.0)(1280, 0.0)(1281, 0.0)(1282, 0.0)(1283, 0.0)(1284, 0.0)(1285, 0.0)(1286, 0.0)(1287, 0.0)(1288, 0.0)(1289, 0.0)(1290, 0.0)(1291, 0.0)(1292, 0.0)(1293, 0.0)(1294, 0.0)(1295, 0.0)(1296, 0.0)(1297, 0.0)(1298, 0.0)(1299, 0.0)(1300, 0.0)(1301, 0.0)(1302, 0.0)(1303, 0.0)(1304, 0.0)(1305, 0.0)(1306, 0.0)(1307, 0.0)(1308, 0.0)(1309, 0.0)(1310, 0.0)(1311, 0.0)(1312, 0.0)(1313, 0.0)(1314, 0.0)(1315, 0.0)(1316, 0.0)(1317, 0.0)(1318, 0.0)(1319, 0.0)(1320, 0.0)(1321, 0.0)(1322, 0.0)(1323, 0.0)(1324, 0.0)(1325, 0.0)(1326, 0.0)(1327, 0.0)(1328, 0.0)(1329, 0.0)(1330, 0.0)(1331, 0.0)(1332, 0.0)(1333, 0.0)(1334, 0.0)(1335, 0.0)(1336, 0.0)(1337, 0.0)(1338, 0.0)(1339, 0.0)(1340, 0.0)(1341, 0.0)(1342, 0.0)(1343, 0.0)(1344, 0.0)(1345, 0.0)(1346, 0.0)(1347, 0.0)(1348, 0.0)(1349, 0.0)(1350, 0.0)(1351, 0.0)(1352, 0.0)(1353, 0.0)(1354, 0.0)(1355, 0.0)(1356, 0.0)(1357, 0.0)(1358, 0.0)(1359, 0.0)(1360, 0.0)(1361, 0.0)(1362, 0.0)(1363, 0.0)(1364, 0.0)(1365, 0.0)(1366, 0.0)(1367, 0.0)(1368, 0.0)(1369, 0.0)(1370, 0.0)(1371, 0.0)(1372, 0.0)(1373, 0.0)(1374, 0.0)(1375, 0.0)(1376, 0.0)(1377, 0.0)(1378, 0.0)(1379, 0.0)(1380, 0.0)(1381, 0.0)(1382, 0.0)(1383, 0.0)(1384, 0.0)(1385, 0.0)(1386, 0.0)(1387, 0.0)(1388, 0.0)(1389, 0.0)(1390, 0.0)(1391, 0.0)(1392, 0.0)(1393, 0.0)(1394, 0.0)(1395, 0.0)(1396, 0.0)(1397, 0.0)(1398, 0.0)(1399, 0.0)(1400, 0.0)(1401, 0.0)(1402, 0.0)(1403, 0.0)(1404, 0.0)(1405, 0.0)(1406, 0.0)(1407, 0.0)(1408, 0.0)(1409, 0.0)(1410, 0.0)(1411, 0.0)(1412, 0.0)(1413, 0.0)(1414, 0.0)(1415, 0.0)(1416, 0.0)(1417, 0.0)(1418, 0.0)(1419, 0.0)(1420, 0.0)(1421, 0.0)(1422, 0.0)(1423, 0.0)(1424, 0.0)(1425, 0.0)(1426, 0.0)(1427, 0.0)(1428, 0.0)(1429, 0.0)(1430, 0.0)(1431, 0.0)(1432, 0.0)(1433, 0.0)(1434, 0.0)(1435, 0.0)(1436, 0.0)(1437, 0.0)(1438, 0.0)(1439, 0.0)(1440, 0.0)(1441, 0.0)(1442, 0.0)(1443, 0.0)(1444, 0.0)(1445, 0.0)(1446, 0.0)(1447, 0.0)(1448, 0.0)(1449, 0.0)(1450, 0.0)(1451, 0.0)(1452, 0.0)(1453, 0.0)(1454, 0.0)(1455, 0.0)(1456, 0.0)(1457, 0.0)(1458, 0.0)(1459, 0.0)(1460, 0.0)(1461, 0.0)(1462, 0.0)(1463, 0.0)(1464, 0.0)(1465, 0.0)(1466, 0.0)(1467, 0.0)(1468, 0.0)(1469, 0.0)(1470, 0.0)(1471, 0.0)(1472, 0.0)(1473, 0.0)(1474, 0.0)(1475, 0.0)(1476, 0.0)(1477, 0.0)(1478, 0.0)(1479, 0.0)(1480, 0.0)(1481, 0.0)(1482, 0.0)(1483, 0.0)(1484, 0.0)(1485, 0.0)(1486, 0.0)(1487, 0.0)(1488, 0.0)(1489, 0.0)(1490, 0.0)(1491, 0.0)(1492, 0.0)(1493, 0.0)(1494, 0.0)(1495, 0.0)(1496, 0.0)(1497, 0.0)(1498, 0.0)(1499, 0.0)(1500, 0.0)(1501, 0.0)(1502, 0.0)(1503, 0.0)(1504, 0.0)(1505, 0.0)(1506, 0.0)(1507, 0.0)(1508, 0.0)(1509, 0.0)(1510, 0.0)(1511, 0.0)(1512, 0.0)(1513, 0.0)(1514, 0.0)(1515, 0.0)(1516, 0.0)(1517, 0.0)(1518, 0.0)(1519, 0.0)};
            \addplot+[name path=C,black,line width=.1pt] coordinates {(0, 0.0)(1, 0.0)(2, 0.0)(3, 0.0)(4, 0.0)(5, 0.0)(6, 0.0)(7, 0.0)(8, 0.0)(9, 0.0)(10, 0.0)(11, 0.0)(12, 0.0)(13, 0.0)(14, 0.0)(15, 0.0)(16, 0.0)(17, 0.0)(18, 0.0)(19, 0.0)(20, 0.0)(21, 0.0)(22, 0.0)(23, 0.0)(24, 0.0)(25, 0.0)(26, 0.0)(27, 0.0)(28, 0.0)(29, 0.0)(30, 0.0)(31, 0.0)(32, 0.0)(33, 0.0)(34, 0.0)(35, 0.0)(36, 0.0)(37, 0.0)(38, 0.0)(39, 0.0)(40, 0.0)(41, 0.0)(42, 0.0)(43, 0.0)(44, 0.0)(45, 0.0)(46, 0.0)(47, 0.0)(48, 0.0)(49, 0.0)(50, 0.0)(51, 0.0)(52, 0.0)(53, 0.0)(54, 0.0)(55, 0.0)(56, 0.0)(57, 0.0)(58, 0.0)(59, 0.0)(60, 0.0)(61, 0.0)(62, 0.0)(63, 0.0)(64, 0.0)(65, 0.0)(66, 0.0)(67, 0.0)(68, 0.0)(69, 0.0)(70, 0.0)(71, 0.0)(72, 0.0)(73, 0.0)(74, 0.0)(75, 0.0)(76, 0.0)(77, 0.0)(78, 0.0)(79, 0.0)(80, 0.0)(81, 0.0)(82, 0.0)(83, 0.0)(84, 0.0)(85, 0.0)(86, 0.0)(87, 0.0)(88, 0.0)(89, 0.0)(90, 0.0)(91, 0.0)(92, 0.0)(93, 0.0)(94, 0.0)(95, 0.0)(96, 0.0)(97, 0.0)(98, 0.0)(99, 0.0)(100, 0.0)(101, 0.0)(102, 0.0)(103, 0.0)(104, 0.0)(105, 0.0)(106, 0.0)(107, 0.0)(108, 0.0)(109, 0.0)(110, 0.0)(111, 0.0)(112, 0.0)(113, 0.0)(114, 0.0)(115, 0.0)(116, 0.0)(117, 0.0)(118, 0.0)(119, 0.0)(120, 0.0)(121, 0.0)(122, 0.0)(123, 0.0)(124, 0.0)(125, 0.0)(126, 0.0)(127, 0.0)(128, 0.0)(129, 0.0)(130, 0.0)(131, 0.0)(132, 0.0)(133, 0.0)(134, 0.0)(135, 0.0)(136, 0.0)(137, 0.0)(138, 0.0)(139, 0.0)(140, 0.0)(141, 0.0)(142, 0.0)(143, 0.0)(144, 0.0)(145, 0.0)(146, 0.0)(147, 0.0)(148, 0.0)(149, 0.0)(150, 0.0)(151, 0.0)(152, 0.0)(153, 0.0)(154, 0.0)(155, 0.0)(156, 0.0)(157, 0.0)(158, 0.0)(159, 0.0)(160, 0.0)(161, 0.0)(162, 0.0)(163, 0.0)(164, 0.0)(165, 0.0)(166, 0.0)(167, 0.0)(168, 0.0)(169, 0.0)(170, 0.0)(171, 0.0)(172, 0.0)(173, 0.0)(174, 0.0)(175, 0.0)(176, 0.0)(177, 0.0)(178, 0.0)(179, 0.0)(180, 0.0)(181, 0.0)(182, 0.0)(183, 0.0)(184, 0.0)(185, 0.0)(186, 0.0)(187, 0.0)(188, 0.0)(189, 0.0)(190, 0.0)(191, 0.0)(192, 0.0)(193, 0.0)(194, 0.0)(195, 0.0)(196, 0.0)(197, 0.0)(198, 0.0)(199, 0.0)(200, 0.0)(201, 0.0)(202, 0.0)(203, 0.0)(204, 0.0)(205, 0.0)(206, 0.0)(207, 0.0)(208, 0.0)(209, 0.0)(210, 0.0)(211, 0.0)(212, 0.0)(213, 0.0)(214, 0.0)(215, 0.0)(216, 0.0)(217, 0.0)(218, 0.0)(219, 0.0)(220, 0.0)(221, 0.0)(222, 0.0)(223, 0.0)(224, 0.0)(225, 0.0)(226, 0.0)(227, 0.0)(228, 0.0)(229, 0.0)(230, 0.0)(231, 0.0)(232, 0.0)(233, 0.0)(234, 0.0)(235, 0.0)(236, 0.0)(237, 0.0)(238, 0.0)(239, 0.0)(240, 0.0)(241, 0.0)(242, 0.0)(243, 0.0)(244, 0.0)(245, 0.0)(246, 0.0)(247, 0.0)(248, 0.0)(249, 0.0)(250, 0.0)(251, 0.0)(252, 0.0)(253, 0.0)(254, 0.0)(255, 0.0)(256, 0.0)(257, 0.0)(258, 0.0)(259, 0.0)(260, 0.0)(261, 0.0)(262, 0.0)(263, 0.0)(264, 0.0)(265, 0.0)(266, 0.0)(267, 0.0)(268, 0.0)(269, 0.0)(270, 0.0)(271, 0.0)(272, 0.0)(273, 0.0)(274, 0.0)(275, 0.0)(276, 0.0)(277, 0.0)(278, 0.0)(279, 0.0)(280, 0.0)(281, 0.0)(282, 0.0)(283, 0.0)(284, 0.0)(285, 0.0)(286, 0.0)(287, 0.0)(288, 0.0)(289, 0.0)(290, 0.0)(291, 0.0)(292, 0.0)(293, 0.0)(294, 0.0)(295, 0.0)(296, 0.0)(297, 0.0)(298, 0.0)(299, 0.0)(300, 0.0)(301, 0.0)(302, 0.0)(303, 0.0)(304, 0.0)(305, 0.0)(306, 0.0)(307, 0.0)(308, 0.0)(309, 0.0)(310, 0.0)(311, 0.0)(312, 0.0)(313, 0.0)(314, 0.0)(315, 0.0)(316, 0.0)(317, 0.0)(318, 0.0)(319, 0.0)(320, 0.0)(321, 0.0)(322, 0.0)(323, 0.0)(324, 0.0)(325, 0.0)(326, 0.0)(327, 0.0)(328, 0.0)(329, 0.0)(330, 0.0)(331, 0.0)(332, 0.0)(333, 0.0)(334, 0.0)(335, 0.0)(336, 0.0)(337, 0.0)(338, 0.0)(339, 0.0)(340, 0.0)(341, 0.0)(342, 0.0)(343, 0.0)(344, 0.0)(345, 0.0)(346, 0.0)(347, 0.0)(348, 0.0)(349, 0.0)(350, 0.0)(351, 0.0)(352, 0.0)(353, 0.0)(354, 0.0)(355, 0.0)(356, 0.0)(357, 0.0)(358, 0.0)(359, 0.0)(360, 0.0)(361, 0.0)(362, 0.0)(363, 0.0)(364, 0.0)(365, 0.0)(366, 0.0)(367, 0.0)(368, 0.0)(369, 0.0)(370, 0.0)(371, 0.0)(372, 0.0)(373, 0.0)(374, 0.0)(375, 0.0)(376, 0.0)(377, 0.0)(378, 0.0)(379, 0.0)(380, 0.0)(381, 0.0)(382, 0.0)(383, 0.0)(384, 0.0)(385, 0.0)(386, 0.0)(387, 0.0)(388, 0.0)(389, 0.0)(390, 0.0)(391, 0.0)(392, 0.0)(393, 0.0)(394, 0.0)(395, 0.0)(396, 0.0)(397, 0.0)(398, 0.0)(399, 0.0)(400, 0.0)(401, 0.0)(402, 0.0)(403, 0.0)(404, 0.0)(405, 0.0)(406, 0.0)(407, 0.0)(408, 0.0)(409, 0.0)(410, 0.0)(411, 0.0)(412, 0.0)(413, 0.0)(414, 0.0)(415, 0.0)(416, 0.0)(417, 0.0)(418, 0.0)(419, 0.0)(420, 0.0)(421, 0.0)(422, 0.0)(423, 0.0)(424, 0.0)(425, 0.0)(426, 0.0)(427, 0.0)(428, 0.0)(429, 0.0)(430, 0.0)(431, 0.0)(432, 0.0)(433, 0.0)(434, 0.0)(435, 0.0)(436, 0.0)(437, 0.0)(438, 0.0)(439, 0.0)(440, 0.0)(441, 0.0)(442, 0.0)(443, 0.0)(444, 0.0)(445, 0.0)(446, 0.0)(447, 0.0)(448, 0.0)(449, 0.0)(450, 0.0)(451, 0.0)(452, 0.0)(453, 0.0)(454, 0.0)(455, 0.0)(456, 0.0)(457, 0.0)(458, 0.0)(459, 0.0)(460, 0.0)(461, 0.0)(462, 0.0)(463, 0.0)(464, 0.0)(465, 0.0)(466, 0.0)(467, 0.0)(468, 0.0)(469, 0.0)(470, 0.0)(471, 0.0)(472, 0.0)(473, 0.0)(474, 0.0)(475, 0.0)(476, 0.0)(477, 0.0)(478, 0.0)(479, 0.0)(480, 0.0)(481, 0.0)(482, 0.0)(483, 0.0)(484, 0.0)(485, 0.0)(486, 0.0)(487, 0.0)(488, 0.0)(489, 0.0)(490, 0.0)(491, 0.0)(492, 0.0)(493, 0.0)(494, 0.0)(495, 0.0)(496, 0.0)(497, 0.0)(498, 0.0)(499, 0.0)(500, 0.0)(501, 0.0)(502, 0.0)(503, 0.0)(504, 0.0)(505, 0.0)(506, 0.0)(507, 0.0)(508, 0.0)(509, 0.0)(510, 0.0)(511, 0.0)(512, 0.0)(513, 0.0)(514, 0.0)(515, 0.0)(516, 0.0)(517, 0.0)(518, 0.0)(519, 0.0)(520, 0.0)(521, 0.0)(522, 0.0)(523, 0.0)(524, 0.0)(525, 0.0)(526, 0.0)(527, 0.0)(528, 0.0)(529, 0.0)(530, 0.0)(531, 0.0)(532, 0.0)(533, 0.0)(534, 0.0)(535, 0.0)(536, 0.0)(537, 0.0)(538, 0.0)(539, 0.0)(540, 0.0)(541, 0.0)(542, 0.0)(543, 0.0)(544, 0.0)(545, 0.0)(546, 0.0)(547, 0.0)(548, 0.0)(549, 0.0)(550, 0.0)(551, 0.0)(552, 0.0)(553, 0.0)(554, 0.0)(555, 0.0)(556, 0.0)(557, 0.0)(558, 0.0)(559, 0.0)(560, 0.0)(561, 0.0)(562, 0.0)(563, 0.0)(564, 0.0)(565, 0.0)(566, 0.0)(567, 0.0)(568, 0.0)(569, 0.0)(570, 0.0)(571, 0.0)(572, 0.0)(573, 0.0)(574, 0.0)(575, 0.0)(576, 0.0)(577, 0.0)(578, 0.0)(579, 0.0)(580, 0.0)(581, 0.0)(582, 0.0)(583, 0.0)(584, 0.0)(585, 0.0)(586, 0.0)(587, 0.0)(588, 0.0)(589, 0.0)(590, 0.0)(591, 0.0)(592, 0.0)(593, 0.0)(594, 0.0)(595, 0.0)(596, 0.0)(597, 0.0)(598, 0.0)(599, 0.0)(600, 0.0)(601, 0.0)(602, 0.0)(603, 0.0)(604, 0.0)(605, 0.0)(606, 0.0)(607, 0.0)(608, 0.0)(609, 0.0)(610, 0.0)(611, 0.0)(612, 0.0)(613, 0.0)(614, 0.0)(615, 0.0)(616, 0.0)(617, 0.0)(618, 0.0)(619, 0.0)(620, 0.0)(621, 0.0)(622, 0.0)(623, 0.0)(624, 0.0)(625, 0.0)(626, 0.0)(627, 0.0)(628, 0.0)(629, 0.0)(630, 0.0)(631, 0.0)(632, 0.0)(633, 0.0)(634, 0.0)(635, 0.0)(636, 0.0)(637, 0.0)(638, 0.0)(639, 0.0)(640, 0.0)(641, 0.0)(642, 0.0)(643, 0.0)(644, 0.0)(645, 0.0)(646, 0.0)(647, 0.0)(648, 0.0)(649, 0.0)(650, 0.0)(651, 0.0)(652, 0.0)(653, 0.0)(654, 0.0)(655, 0.0)(656, 0.0)(657, 0.0)(658, 0.0)(659, 0.0)(660, 0.0)(661, 0.0)(662, 0.0)(663, 0.0)(664, 0.0)(665, 0.0)(666, 0.0)(667, 0.0)(668, 0.0)(669, 0.0)(670, 0.0)(671, 0.0)(672, 0.0)(673, 0.0)(674, 0.0)(675, 0.0)(676, 0.0)(677, 0.0)(678, 0.0)(679, 0.0)(680, 0.0)(681, 0.0)(682, 0.0)(683, 0.0)(684, 0.0)(685, 0.0)(686, 0.0)(687, 0.0)(688, 0.0)(689, 0.0)(690, 0.0)(691, 0.0)(692, 0.0)(693, 0.0)(694, 0.0)(695, 0.0)(696, 0.0)(697, 0.0)(698, 0.0)(699, 0.0)(700, 0.0)(701, 0.0)(702, 0.0)(703, 0.0)(704, 0.0)(705, 0.0)(706, 0.0)(707, 0.0)(708, 0.0)(709, 0.0)(710, 0.0)(711, 0.0)(712, 0.0)(713, 0.0)(714, 0.0)(715, 0.0)(716, 0.0)(717, 0.0)(718, 0.0)(719, 0.0)(720, 0.0)(721, 0.0)(722, 0.0)(723, 0.0)(724, 0.0)(725, 0.0)(726, 0.0)(727, 0.0)(728, 0.0)(729, 0.0)(730, 0.0)(731, 0.0)(732, 0.0)(733, 0.0)(734, 0.0)(735, 0.0)(736, 0.0)(737, 0.0)(738, 0.0)(739, 0.0)(740, 0.0)(741, 0.0)(742, 0.0)(743, 0.0)(744, 0.0)(745, 0.0)(746, 0.0)(747, 0.0)(748, 0.0)(749, 0.0)(750, 0.0)(751, 0.0)(752, 0.0)(753, 0.0)(754, 0.0)(755, 0.0)(756, 0.0)(757, 0.0)(758, 0.0)(759, 0.0)(760, 0.0)(761, 0.0)(762, 0.0)(763, 0.0)(764, 0.0)(765, 0.0)(766, 0.0)(767, 0.0)(768, 0.0)(769, 0.0)(770, 0.0)(771, 0.0)(772, 0.0)(773, 0.0)(774, 0.0)(775, 0.0)(776, 0.0)(777, 0.0)(778, 0.0)(779, 0.0)(780, 0.0)(781, 0.0)(782, 0.0)(783, 0.0)(784, 0.0)(785, 0.0)(786, 0.0)(787, 0.0)(788, 0.0)(789, 0.0)(790, 0.0)(791, 0.0)(792, 0.0)(793, 0.0)(794, 0.0)(795, 0.0)(796, 0.0)(797, 0.0)(798, 0.0)(799, 0.0)(800, 0.0)(801, 0.0)(802, 0.0)(803, 0.0)(804, 0.0)(805, 0.0)(806, 0.0)(807, 0.0)(808, 0.0)(809, 0.0)(810, 0.0)(811, 0.0)(812, 0.0)(813, 0.0)(814, 0.0)(815, 0.0)(816, 0.0)(817, 0.0)(818, 0.0)(819, 0.0)(820, 0.0)(821, 0.0)(822, 0.0)(823, 0.0)(824, 0.0)(825, 0.0)(826, 0.0)(827, 0.0)(828, 0.0)(829, 0.0)(830, 0.0)(831, 0.0)(832, 0.0)(833, 0.0)(834, 0.0)(835, 0.0)(836, 0.0)(837, 0.0)(838, 0.0)(839, 0.0)(840, 0.0)(841, 0.0)(842, 0.0)(843, 0.0)(844, 0.0)(845, 0.0)(846, 0.0)(847, 0.0)(848, 0.0)(849, 0.0)(850, 0.0)(851, 0.0)(852, 0.0)(853, 0.0)(854, 0.0)(855, 0.0)(856, 0.0)(857, 0.0)(858, 0.0)(859, 0.0)(860, 0.0)(861, 0.0)(862, 0.0)(863, 0.0)(864, 0.0)(865, 0.0)(866, 0.0)(867, 0.0)(868, 0.0)(869, 0.0)(870, 0.0)(871, 0.0)(872, 0.0)(873, 0.0)(874, 0.0)(875, 0.0)(876, 0.0)(877, 0.0)(878, 0.0)(879, 0.0)(880, 0.0)(881, 0.0)(882, 0.0)(883, 0.0)(884, 0.0)(885, 0.0)(886, 0.0)(887, 0.0)(888, 0.0)(889, 0.0)(890, 0.0)(891, 0.0)(892, 0.0)(893, 0.0)(894, 0.0)(895, 0.0)(896, 0.0)(897, 0.0)(898, 0.0)(899, 0.0)(900, 0.0)(901, 0.0)(902, 0.0)(903, 0.0)(904, 0.0)(905, 0.0)(906, 0.0)(907, 0.0)(908, 0.0)(909, 0.0)(910, 0.0)(911, 0.0)(912, 0.0)(913, 0.0)(914, 0.0)(915, 0.0)(916, 0.0)(917, 0.0)(918, 0.0)(919, 0.0)(920, 0.0)(921, 0.0)(922, 0.0)(923, 0.0)(924, 0.0)(925, 0.0)(926, 0.0)(927, 0.0)(928, 0.0)(929, 0.0)(930, 0.0)(931, 0.0)(932, 0.0)(933, 0.0)(934, 0.0)(935, 0.0)(936, 0.0)(937, 0.0)(938, 0.0)(939, 0.0)(940, 0.0)(941, 0.0)(942, 0.0)(943, 0.0)(944, 0.0)(945, 0.0)(946, 0.0)(947, 0.0)(948, 0.0)(949, 0.0)(950, 0.0)(951, 0.0)(952, 0.0)(953, 0.0)(954, 0.0)(955, 0.0)(956, 0.0)(957, 0.0)(958, 0.0)(959, 0.0)(960, 0.0)(961, 0.0)(962, 0.0)(963, 0.0)(964, 0.0)(965, 0.0)(966, 0.0)(967, 0.0)(968, 0.0)(969, 0.0)(970, 0.0)(971, 0.0)(972, 0.0)(973, 0.0)(974, 0.0)(975, 0.0)(976, 0.0)(977, 0.0)(978, 0.0)(979, 0.0)(980, 0.0)(981, 0.0)(982, 0.0)(983, 0.0)(984, 0.0)(985, 0.0)(986, 0.0)(987, 0.0)(988, 0.0)(989, 0.0)(990, 0.0)(991, 0.0)(992, 0.0)(993, 0.0)(994, 0.0)(995, 0.0)(996, 0.0)(997, 0.0)(998, 0.0)(999, 0.0)(1000, 0.0)(1001, 0.0)(1002, 0.0)(1003, 0.0)(1004, 0.0)(1005, 0.0)(1006, 0.0)(1007, 0.0)(1008, 0.0)(1009, 0.0)(1010, 0.0)(1011, 0.0)(1012, 0.0)(1013, 0.0)(1014, 0.0)(1015, 0.0)(1016, 0.0)(1017, 0.0)(1018, 0.0)(1019, 0.0)(1020, 0.0)(1021, 0.0)(1022, 0.0)(1023, 0.0)(1024, 0.0)(1025, 0.0)(1026, 0.0)(1027, 0.0)(1028, 0.0)(1029, 0.0)(1030, 0.0)(1031, 0.0)(1032, 0.0)(1033, 0.0)(1034, 0.0)(1035, 0.0)(1036, 0.0)(1037, 0.0)(1038, 0.0)(1039, 0.0)(1040, 0.0)(1041, 0.0)(1042, 0.0)(1043, 0.0)(1044, 0.0)(1045, 0.0)(1046, 0.0)(1047, 0.0)(1048, 0.0)(1049, 0.0)(1050, 0.0)(1051, 0.0)(1052, 0.0)(1053, 0.0)(1054, 0.0)(1055, 0.0)(1056, 0.0)(1057, 0.0)(1058, 0.0)(1059, 0.0)(1060, 0.0)(1061, 0.0)(1062, 0.0)(1063, 0.0)(1064, 0.0)(1065, 0.0)(1066, 0.0)(1067, 0.0)(1068, 0.0)(1069, 0.0)(1070, 0.0)(1071, 0.0)(1072, 0.0)(1073, 0.0)(1074, 0.0)(1075, 0.0)(1076, 0.0)(1077, 0.0)(1078, 0.0)(1079, 0.0)(1080, 0.0)(1081, 0.0)(1082, 0.0)(1083, 0.0)(1084, 0.0)(1085, 0.0)(1086, 0.0)(1087, 0.0)(1088, 0.0)(1089, 0.0)(1090, 0.0)(1091, 0.0)(1092, 0.0)(1093, 0.0)(1094, 0.0)(1095, 0.0)(1096, 0.0)(1097, 0.0)(1098, 0.0)(1099, 0.0)(1100, 0.0)(1101, 0.0)(1102, 0.0)(1103, 0.0)(1104, 0.0)(1105, 0.0)(1106, 0.0)(1107, 0.0)(1108, 0.0)(1109, 0.0)(1110, 0.0)(1111, 0.0)(1112, 0.0)(1113, 0.0)(1114, 0.0)(1115, 0.0)(1116, 0.0)(1117, 0.0)(1118, 0.0)(1119, 0.0)(1120, 0.0)(1121, 0.0)(1122, 0.0)(1123, 0.0)(1124, 0.0)(1125, 0.0)(1126, 0.0)(1127, 0.0)(1128, 0.0)(1129, 0.0)(1130, 0.0)(1131, 0.0)(1132, 0.0)(1133, 0.0)(1134, 0.0)(1135, 0.0)(1136, 0.0)(1137, 0.0)(1138, 0.0)(1139, 0.0)(1140, 0.0)(1141, 0.0)(1142, 0.0)(1143, 0.0)(1144, 0.0)(1145, 0.0)(1146, 0.0)(1147, 0.0)(1148, 0.0)(1149, 0.0)(1150, 0.0)(1151, 0.0)(1152, 0.0)(1153, 0.0)(1154, 0.0)(1155, 0.0)(1156, 0.0)(1157, 0.0)(1158, 0.0)(1159, 0.0)(1160, 0.0)(1161, 0.0)(1162, 0.0)(1163, 0.0)(1164, 0.0)(1165, 0.0)(1166, 0.0)(1167, 0.0)(1168, 0.0)(1169, 0.0)(1170, 0.0)(1171, 0.0)(1172, 0.0)(1173, 0.0)(1174, 0.0)(1175, 0.0)(1176, 0.0)(1177, 0.0)(1178, 0.0)(1179, 0.0)(1180, 0.0)(1181, 0.0)(1182, 0.0)(1183, 0.0)(1184, 0.0)(1185, 0.0)(1186, 0.0)(1187, 0.0)(1188, 0.0)(1189, 0.0)(1190, 0.0)(1191, 0.0)(1192, 0.0)(1193, 0.0)(1194, 0.0)(1195, 0.0)(1196, 0.0)(1197, 0.0)(1198, 0.0)(1199, 0.0)(1200, 0.0)(1201, 0.0)(1202, 0.0)(1203, 0.0)(1204, 0.0)(1205, 0.0)(1206, 0.0)(1207, 0.0)(1208, 0.0)(1209, 0.0)(1210, 0.0)(1211, 0.0)(1212, 0.0)(1213, 0.0)(1214, 0.0)(1215, 0.0)(1216, 0.0)(1217, 0.0)(1218, 0.0)(1219, 0.0)(1220, 0.0)(1221, 0.0)(1222, 0.0)(1223, 0.0)(1224, 0.0)(1225, 0.0)(1226, 0.0)(1227, 0.0)(1228, 0.0)(1229, 0.0)(1230, 0.0)(1231, 0.0)(1232, 0.0)(1233, 0.0)(1234, 0.0)(1235, 0.0)(1236, 0.0)(1237, 0.0)(1238, 0.0)(1239, 0.0)(1240, 0.0)(1241, 0.0)(1242, 0.0)(1243, 0.0)(1244, 0.0)(1245, 0.0)(1246, 0.0)(1247, 0.0)(1248, 0.0)(1249, 0.0)(1250, 0.0)(1251, 0.0)(1252, 0.0)(1253, 0.0)(1254, 0.0)(1255, 0.0)(1256, 0.0)(1257, 0.0)(1258, 0.0)(1259, 0.0)(1260, 0.0)(1261, 0.0)(1262, 0.0)(1263, 0.0)(1264, 0.0)(1265, 0.0)(1266, 0.0)(1267, 0.0)(1268, 0.0)(1269, 0.0)(1270, 0.0)(1271, 0.0)(1272, 0.0)(1273, 0.0)(1274, 0.0)(1275, 0.0)(1276, 0.0)(1277, 0.0)(1278, 0.0)(1279, 0.0)(1280, 0.0)(1281, 0.0)(1282, 0.0)(1283, 0.0)(1284, 0.0)(1285, 0.0)(1286, 0.0)(1287, 0.0)(1288, 0.0)(1289, 0.0)(1290, 0.0)(1291, 0.0)(1292, 0.0)(1293, 0.0)(1294, 0.0)(1295, 0.0)(1296, 0.0)(1297, 0.0)(1298, 0.0)(1299, 0.0)(1300, 0.0)(1301, 0.0)(1302, 0.0)(1303, 0.0)(1304, 0.0)(1305, 0.0)(1306, 0.0)(1307, 0.0)(1308, 0.0)(1309, 0.0)(1310, 0.0)(1311, 0.0)(1312, 0.0)(1313, 0.0)(1314, 0.0)(1315, 0.0)(1316, 0.0)(1317, 0.0)(1318, 0.0)(1319, 0.0)(1320, 0.0)(1321, 0.0)(1322, 0.0)(1323, 0.0)(1324, 0.0)(1325, 0.0)(1326, 0.0)(1327, 0.0)(1328, 0.0)(1329, 0.0)(1330, 0.0)(1331, 0.0)(1332, 0.0)(1333, 0.0)(1334, 0.0)(1335, 0.0)(1336, 0.0)(1337, 0.0)(1338, 0.0)(1339, 0.0)(1340, 0.0)(1341, 0.0)(1342, 0.0)(1343, 0.0)(1344, 0.0)(1345, 0.0)(1346, 0.0)(1347, 0.0)(1348, 0.0)(1349, 0.0)(1350, 0.0)(1351, 0.0)(1352, 0.0)(1353, 0.0)(1354, 0.0)(1355, 0.0)(1356, 0.0)(1357, 0.0)(1358, 0.0)(1359, 0.0)(1360, 0.0)(1361, 0.0)(1362, 0.0)(1363, 0.0)(1364, 0.0)(1365, 0.0)(1366, 0.0)(1367, 0.0)(1368, 0.0)(1369, 0.0)(1370, 0.0)(1371, 0.0)(1372, 0.0)(1373, 0.0)(1374, 0.0)(1375, 0.0)(1376, 0.0)(1377, 0.0)(1378, 0.0)(1379, 0.0)(1380, 0.0)(1381, 0.0)(1382, 0.0)(1383, 0.0)(1384, 0.0)(1385, 0.0)(1386, 0.0)(1387, 0.0)(1388, 0.0)(1389, 0.0)(1390, 0.0)(1391, 0.0)(1392, 0.0)(1393, 0.0)(1394, 0.0)(1395, 0.0)(1396, 0.0)(1397, 0.0)(1398, 0.0)(1399, 0.0)(1400, 0.0)(1401, 0.0)(1402, 0.0)(1403, 0.0)(1404, 0.0)(1405, 0.0)(1406, 0.0)(1407, 0.0)(1408, 0.0)(1409, 0.0)(1410, 0.0)(1411, 0.0)(1412, 0.0)(1413, 0.0)(1414, 0.0)(1415, 0.0)(1416, 0.0)(1417, 0.0)(1418, 0.0)(1419, 0.0)(1420, 0.0)(1421, 0.0)(1422, 0.0)(1423, 0.0)(1424, 0.0)(1425, 0.0)(1426, 0.0)(1427, 0.0)(1428, 0.0)(1429, 0.0)(1430, 0.0)(1431, 0.0)(1432, 0.0)(1433, 0.0)(1434, 0.0)(1435, 0.0)(1436, 0.0)(1437, 0.0)(1438, 0.0)(1439, 0.0)(1440, 0.0)(1441, 0.0)(1442, 0.0)(1443, 0.0)(1444, 0.0)(1445, 0.0)(1446, 0.0)(1447, 0.0)(1448, 0.0)(1449, 0.0)(1450, 0.0)(1451, 0.0)(1452, 0.0)(1453, 0.0)(1454, 0.0)(1455, 0.0)(1456, 0.0)(1457, 0.0)(1458, 0.0)(1459, 0.0)(1460, 0.0)(1461, 0.0)(1462, 0.0)(1463, 0.0)(1464, 0.0)(1465, 0.0)(1466, 0.0)(1467, 0.0)(1468, 0.0)(1469, 0.0)(1470, 0.0)(1471, 0.0)(1472, 0.0)(1473, 0.0)(1474, 0.0)(1475, 0.0)(1476, 0.0)(1477, 0.0)(1478, 0.0)(1479, 0.0)(1480, 0.0)(1481, 0.0)(1482, 0.0)(1483, 0.0)(1484, 0.0)(1485, 0.0)(1486, 0.0)(1487, 0.0)(1488, 0.0)(1489, 0.0)(1490, 0.0)(1491, 0.0)(1492, 0.0)(1493, 0.0)(1494, 0.0)(1495, 0.0)(1496, 0.0)(1497, 0.0)(1498, 0.0)(1499, 0.0)(1500, 0.0)(1501, 0.0)(1502, 0.0)(1503, 0.0)(1504, 0.0)(1505, 0.0)(1506, 0.0)(1507, 0.0)(1508, 0.0)(1509, 0.0)(1510, 0.0)(1511, 0.0)(1512, 0.0)(1513, 0.0)(1514, 0.0)(1515, 0.0)(1516, 0.0)(1517, 0.0)(1518, 0.0)(1519, 0.0)};
            \addplot+[name path=D,black,line width=.1pt] coordinates {(0, 0.0)(1, 0.0)(2, 0.0)(3, 0.0)(4, 0.0)(5, 0.0)(6, 0.0)(7, 0.0)(8, 0.0)(9, 0.0)(10, 0.0)(11, 0.0)(12, 0.0)(13, 0.0)(14, 0.0)(15, 0.0)(16, 0.0)(17, 0.0)(18, 0.0)(19, 0.0)(20, 0.0)(21, 0.0)(22, 0.0)(23, 0.0)(24, 0.0)(25, 0.0)(26, 0.0)(27, 0.0)(28, 0.0)(29, 0.0)(30, 0.0)(31, 0.0)(32, 0.0)(33, 0.0)(34, 0.0)(35, 0.0)(36, 0.0)(37, 0.0)(38, 0.0)(39, 0.0)(40, 0.0)(41, 0.0)(42, 0.0)(43, 0.0)(44, 0.0)(45, 0.0)(46, 0.0)(47, 0.0)(48, 0.0)(49, 0.0)(50, 0.0)(51, 0.0)(52, 0.0)(53, 0.0)(54, 0.0)(55, 0.0)(56, 0.0)(57, 0.0)(58, 0.0)(59, 0.0)(60, 0.0)(61, 0.0)(62, 0.0)(63, 0.0)(64, 0.0)(65, 0.0)(66, 0.0)(67, 0.0)(68, 0.0)(69, 0.0)(70, 0.0)(71, 0.0)(72, 0.0)(73, 0.0)(74, 0.0)(75, 0.0)(76, 0.0)(77, 0.0)(78, 0.0)(79, 0.0)(80, 0.0)(81, 0.0)(82, 0.0)(83, 0.0)(84, 0.0)(85, 0.0)(86, 0.0)(87, 0.0)(88, 0.0)(89, 0.0)(90, 0.0)(91, 0.0)(92, 0.0)(93, 0.0)(94, 0.0)(95, 0.0)(96, 0.0)(97, 0.0)(98, 0.0)(99, 0.0)(100, 0.0)(101, 0.0)(102, 0.0)(103, 0.0)(104, 0.0)(105, 0.0)(106, 0.0)(107, 0.0)(108, 0.0)(109, 0.0)(110, 0.0)(111, 0.0)(112, 0.0)(113, 0.0)(114, 0.0)(115, 0.0)(116, 0.0)(117, 0.0)(118, 0.0)(119, 0.0)(120, 0.0)(121, 0.0)(122, 0.0)(123, 0.0)(124, 0.0)(125, 0.0)(126, 0.0)(127, 0.0)(128, 0.0)(129, 0.0)(130, 0.0)(131, 0.0)(132, 0.0)(133, 0.0)(134, 0.0)(135, 0.0)(136, 0.0)(137, 0.0)(138, 0.0)(139, 0.0)(140, 0.0)(141, 0.0)(142, 0.0)(143, 0.0)(144, 0.0)(145, 0.0)(146, 0.0)(147, 0.0)(148, 0.0)(149, 0.0)(150, 0.0)(151, 0.0)(152, 0.0)(153, 0.0)(154, 0.0)(155, 0.0)(156, 0.0)(157, 0.0)(158, 0.0)(159, 0.0)(160, 0.0)(161, 0.0)(162, 0.0)(163, 0.0)(164, 0.0)(165, 0.0)(166, 0.0)(167, 0.0)(168, 0.0)(169, 0.0)(170, 0.0)(171, 0.0)(172, 0.0)(173, 0.0)(174, 0.0)(175, 0.0)(176, 0.0)(177, 0.0)(178, 0.0)(179, 0.0)(180, 0.0)(181, 0.0)(182, 0.0)(183, 0.0)(184, 0.0)(185, 0.0)(186, 0.0)(187, 0.0)(188, 0.0)(189, 0.0)(190, 0.0)(191, 0.0)(192, 0.0)(193, 0.0)(194, 0.0)(195, 0.0)(196, 0.0)(197, 0.0)(198, 0.0)(199, 0.0)(200, 0.0)(201, 0.0)(202, 0.0)(203, 0.0)(204, 0.0)(205, 0.0)(206, 0.0)(207, 0.0)(208, 0.0)(209, 0.0)(210, 0.0)(211, 0.0)(212, 0.0)(213, 0.0)(214, 0.0)(215, 0.0)(216, 0.0)(217, 0.0)(218, 0.0)(219, 0.0)(220, 0.0)(221, 0.0)(222, 0.0)(223, 0.0)(224, 0.0)(225, 0.0)(226, 0.0)(227, 0.0)(228, 0.0)(229, 0.0)(230, 0.0)(231, 0.0)(232, 0.0)(233, 0.0)(234, 0.0)(235, 0.0)(236, 0.0)(237, 0.0)(238, 0.0)(239, 0.0)(240, 0.0)(241, 0.0)(242, 0.0)(243, 0.0)(244, 0.0)(245, 0.0)(246, 0.0)(247, 0.0)(248, 0.0)(249, 0.0)(250, 0.0)(251, 0.0)(252, 0.0)(253, 0.0)(254, 0.0)(255, 0.0)(256, 0.0)(257, 0.0)(258, 0.0)(259, 0.0)(260, 0.0)(261, 0.0)(262, 0.0)(263, 0.0)(264, 0.0)(265, 0.0)(266, 0.0)(267, 0.0)(268, 0.0)(269, 0.0)(270, 0.0)(271, 0.0)(272, 0.0)(273, 0.0)(274, 0.0)(275, 0.0)(276, 0.0)(277, 0.0)(278, 0.0)(279, 0.0)(280, 0.0)(281, 0.0)(282, 0.0)(283, 0.0)(284, 0.0)(285, 0.0)(286, 0.0)(287, 0.0)(288, 0.0)(289, 0.0)(290, 0.0)(291, 0.0)(292, 0.0)(293, 0.0)(294, 0.0)(295, 0.0)(296, 0.0)(297, 0.0)(298, 0.0)(299, 0.0)(300, 0.0)(301, 0.0)(302, 0.0)(303, 0.0)(304, 0.0)(305, 0.0)(306, 0.0)(307, 0.0)(308, 0.0)(309, 0.0)(310, 0.0)(311, 0.0)(312, 0.0)(313, 0.0)(314, 0.0)(315, 0.0)(316, 0.0)(317, 0.0)(318, 0.0)(319, 0.0)(320, 0.0)(321, 0.0)(322, 0.0)(323, 0.0)(324, 0.0)(325, 0.0)(326, 0.0)(327, 0.0)(328, 0.0)(329, 0.0)(330, 0.0)(331, 0.0)(332, 0.0)(333, 0.0)(334, 0.0)(335, 0.0)(336, 0.0)(337, 0.0)(338, 0.0)(339, 0.0)(340, 0.0)(341, 0.0)(342, 0.0)(343, 0.0)(344, 0.0)(345, 0.0)(346, 0.0)(347, 0.0)(348, 0.0)(349, 0.0)(350, 0.0)(351, 0.0)(352, 0.0)(353, 0.0)(354, 0.0)(355, 0.0)(356, 0.0)(357, 0.0)(358, 0.0)(359, 0.0)(360, 0.0)(361, 0.0)(362, 0.0)(363, 0.0)(364, 0.0)(365, 0.0)(366, 0.0)(367, 0.0)(368, 0.0)(369, 0.0)(370, 0.0)(371, 0.0)(372, 0.0)(373, 0.0)(374, 0.0)(375, 0.0)(376, 0.0)(377, 0.0)(378, 0.0)(379, 0.0)(380, 0.0)(381, 0.0)(382, 0.0)(383, 0.0)(384, 0.0)(385, 0.0)(386, 0.0)(387, 0.0)(388, 0.0)(389, 0.0)(390, 0.0)(391, 0.0)(392, 0.0)(393, 0.0)(394, 0.0)(395, 0.0)(396, 0.0)(397, 0.0)(398, 0.0)(399, 0.0)(400, 0.0)(401, 0.0)(402, 0.0)(403, 0.0)(404, 0.0)(405, 0.0)(406, 0.0)(407, 0.0)(408, 0.0)(409, 0.0)(410, 0.0)(411, 0.0)(412, 0.0)(413, 0.0)(414, 0.0)(415, 0.0)(416, 0.0)(417, 0.0)(418, 0.0)(419, 0.0)(420, 0.0)(421, 0.0)(422, 0.0)(423, 0.0)(424, 0.0)(425, 0.0)(426, 0.0)(427, 0.0)(428, 0.0)(429, 0.0)(430, 0.0)(431, 0.0)(432, 0.0)(433, 0.0)(434, 0.0)(435, 0.0)(436, 0.0)(437, 0.0)(438, 0.0)(439, 0.0)(440, 0.0)(441, 0.0)(442, 0.0)(443, 0.0)(444, 0.0)(445, 0.0)(446, 0.0)(447, 0.0)(448, 0.0)(449, 0.0)(450, 0.0)(451, 0.0)(452, 0.0)(453, 0.0)(454, 0.0)(455, 0.0)(456, 0.0)(457, 0.0)(458, 0.0)(459, 0.0)(460, 0.0)(461, 0.0)(462, 0.0)(463, 0.0)(464, 0.0)(465, 0.0)(466, 0.0)(467, 0.0)(468, 0.0)(469, 0.0)(470, 0.0)(471, 0.0)(472, 0.0)(473, 0.0)(474, 0.0)(475, 0.0)(476, 0.0)(477, 0.0)(478, 0.0)(479, 0.0)(480, 0.0)(481, 0.0)(482, 0.0)(483, 0.0)(484, 0.0)(485, 0.0)(486, 0.0)(487, 0.0)(488, 0.0)(489, 0.0)(490, 0.0)(491, 0.0)(492, 0.0)(493, 0.0)(494, 0.0)(495, 0.0)(496, 0.0)(497, 0.0)(498, 0.0)(499, 0.0)(500, 0.0)(501, 0.0)(502, 0.0)(503, 0.0)(504, 0.0)(505, 0.0)(506, 0.0)(507, 0.0)(508, 0.0)(509, 0.0)(510, 0.0)(511, 0.0)(512, 0.0)(513, 0.0)(514, 0.0)(515, 0.0)(516, 0.0)(517, 0.0)(518, 0.0)(519, 0.0)(520, 0.0)(521, 0.0)(522, 0.0)(523, 0.0)(524, 0.0)(525, 0.0)(526, 0.0)(527, 0.0)(528, 0.0)(529, 0.0)(530, 0.0)(531, 0.0)(532, 0.0)(533, 0.0)(534, 0.0)(535, 0.0)(536, 0.0)(537, 0.0)(538, 0.0)(539, 0.0)(540, 0.0)(541, 0.0)(542, 0.0)(543, 0.0)(544, 0.0)(545, 0.0)(546, 0.0)(547, 0.0)(548, 0.0)(549, 0.0)(550, 0.0)(551, 0.0)(552, 0.0)(553, 0.0)(554, 0.0)(555, 0.0)(556, 0.0)(557, 0.0)(558, 0.0)(559, 0.0)(560, 0.0)(561, 0.0)(562, 0.0)(563, 0.0)(564, 0.0)(565, 0.0)(566, 0.0)(567, 0.0)(568, 0.0)(569, 0.0)(570, 0.0)(571, 0.0)(572, 0.0)(573, 0.0)(574, 0.0)(575, 0.0)(576, 0.0)(577, 0.0)(578, 0.0)(579, 0.0)(580, 0.0)(581, 0.0)(582, 0.0)(583, 0.0)(584, 0.0)(585, 0.0)(586, 0.0)(587, 0.0)(588, 0.0)(589, 0.0)(590, 0.0)(591, 0.0)(592, 0.0)(593, 0.0)(594, 0.0)(595, 0.0)(596, 0.0)(597, 0.0)(598, 0.0)(599, 0.0)(600, 0.0)(601, 0.0)(602, 0.0)(603, 0.0)(604, 0.0)(605, 0.0)(606, 0.0)(607, 0.0)(608, 0.0)(609, 0.0)(610, 0.0)(611, 0.0)(612, 0.0)(613, 0.0)(614, 0.0)(615, 0.0)(616, 0.0)(617, 0.0)(618, 0.0)(619, 0.0)(620, 0.0)(621, 0.0)(622, 0.0)(623, 0.0)(624, 0.0)(625, 0.0)(626, 0.0)(627, 0.0)(628, 0.0)(629, 0.0)(630, 0.0)(631, 0.0)(632, 0.0)(633, 0.0)(634, 0.0)(635, 0.0)(636, 0.0)(637, 0.0)(638, 0.0)(639, 0.0)(640, 0.0)(641, 0.0)(642, 0.0)(643, 0.0)(644, 0.0)(645, 0.0)(646, 0.0)(647, 0.0)(648, 0.0)(649, 0.0)(650, 0.0)(651, 0.0)(652, 0.0)(653, 0.0)(654, 0.0)(655, 0.0)(656, 0.0)(657, 0.0)(658, 0.0)(659, 0.0)(660, 0.0)(661, 0.0)(662, 0.0)(663, 0.0)(664, 0.0)(665, 0.0)(666, 0.0)(667, 0.0)(668, 0.0)(669, 0.0)(670, 0.0)(671, 0.0)(672, 0.0)(673, 0.0)(674, 0.0)(675, 0.0)(676, 0.0)(677, 0.0)(678, 0.0)(679, 0.0)(680, 0.0)(681, 0.0)(682, 0.0)(683, 0.0)(684, 0.0)(685, 0.0)(686, 0.0)(687, 0.0)(688, 0.0)(689, 0.0)(690, 0.0)(691, 0.0)(692, 0.0)(693, 0.0)(694, 0.0)(695, 0.0)(696, 0.0)(697, 0.0)(698, 0.0)(699, 0.0)(700, 0.0)(701, 0.0)(702, 0.0)(703, 0.0)(704, 0.0)(705, 0.0)(706, 0.0)(707, 0.0)(708, 0.0)(709, 0.0)(710, 0.0)(711, 0.0)(712, 0.0)(713, 0.0)(714, 0.0)(715, 0.0)(716, 0.0)(717, 0.0)(718, 0.0)(719, 0.0)(720, 0.0)(721, 0.0)(722, 0.0)(723, 0.0)(724, 0.0)(725, 0.0)(726, 0.0)(727, 0.0)(728, 0.0)(729, 0.0)(730, 0.0)(731, 0.0)(732, 0.0)(733, 0.0)(734, 0.0)(735, 0.0)(736, 0.0)(737, 0.0)(738, 0.0)(739, 0.0)(740, 0.0)(741, 0.0)(742, 0.0)(743, 0.0)(744, 0.0)(745, 0.0)(746, 0.0)(747, 0.0)(748, 0.0)(749, 0.0)(750, 0.0)(751, 0.0)(752, 0.0)(753, 0.0)(754, 0.0)(755, 0.0)(756, 0.0)(757, 0.0)(758, 0.0)(759, 0.0)(760, 0.0)(761, 0.0)(762, 0.0)(763, 0.0)(764, 0.0)(765, 0.0)(766, 0.0)(767, 0.0)(768, 0.0)(769, 0.0)(770, 0.0)(771, 0.0)(772, 0.0)(773, 0.0)(774, 0.0)(775, 0.0)(776, 0.0)(777, 0.0)(778, 0.0)(779, 0.0)(780, 0.0)(781, 0.0)(782, 0.0)(783, 0.0)(784, 0.0)(785, 0.0)(786, 0.0)(787, 0.0)(788, 0.0)(789, 0.0)(790, 0.0)(791, 0.0)(792, 0.0)(793, 0.0)(794, 0.0)(795, 0.0)(796, 0.0)(797, 0.0)(798, 0.0)(799, 0.0)(800, 0.0)(801, 0.0)(802, 0.0)(803, 0.0)(804, 0.0)(805, 0.0)(806, 0.0)(807, 0.0)(808, 0.0)(809, 0.0)(810, 0.0)(811, 0.0)(812, 0.0)(813, 0.0)(814, 0.0)(815, 0.0)(816, 0.0)(817, 0.0)(818, 0.0)(819, 0.0)(820, 0.0)(821, 0.0)(822, 0.0)(823, 0.0)(824, 0.0)(825, 0.0)(826, 0.0)(827, 0.0)(828, 0.0)(829, 0.0)(830, 0.0)(831, 0.0)(832, 0.0)(833, 0.0)(834, 0.0)(835, 0.0)(836, 0.0)(837, 0.0)(838, 0.0)(839, 0.0)(840, 0.0)(841, 0.0)(842, 0.0)(843, 0.0)(844, 0.0)(845, 0.0)(846, 0.0)(847, 0.0)(848, 0.0)(849, 0.0)(850, 0.0)(851, 0.0)(852, 0.0)(853, 0.0)(854, 0.0)(855, 0.0)(856, 0.0)(857, 0.0)(858, 0.0)(859, 0.0)(860, 0.0)(861, 0.0)(862, 0.0)(863, 0.0)(864, 0.0)(865, 0.0)(866, 0.0)(867, 0.0)(868, 0.0)(869, 0.0)(870, 0.0)(871, 0.0)(872, 0.0)(873, 0.0)(874, 0.0)(875, 0.0)(876, 0.0)(877, 0.0)(878, 0.0)(879, 0.0)(880, 0.0)(881, 0.0)(882, 0.0)(883, 0.0)(884, 0.0)(885, 0.0)(886, 0.0)(887, 0.0)(888, 0.0)(889, 0.0)(890, 0.0)(891, 0.0)(892, 0.0)(893, 0.0)(894, 0.0)(895, 0.0)(896, 0.0)(897, 0.0)(898, 0.0)(899, 0.0)(900, 0.0)(901, 0.0)(902, 0.0)(903, 0.0)(904, 0.0)(905, 0.0)(906, 0.0)(907, 0.0)(908, 0.0)(909, 0.0)(910, 0.0)(911, 0.0)(912, 0.0)(913, 0.0)(914, 0.0)(915, 0.0)(916, 0.0)(917, 0.0)(918, 0.0)(919, 0.0)(920, 0.0)(921, 0.0)(922, 0.0)(923, 0.0)(924, 0.0)(925, 0.0)(926, 0.0)(927, 0.0)(928, 0.0)(929, 0.0)(930, 0.0)(931, 0.0)(932, 0.0)(933, 0.0)(934, 0.0)(935, 0.0)(936, 0.0)(937, 0.0)(938, 0.0)(939, 0.0)(940, 0.0)(941, 0.0)(942, 0.0)(943, 0.0)(944, 0.0)(945, 0.0)(946, 0.0)(947, 0.0)(948, 0.0)(949, 0.0)(950, 0.0)(951, 0.0)(952, 0.0)(953, 0.0)(954, 0.0)(955, 0.0)(956, 0.0)(957, 0.0)(958, 0.0)(959, 0.0)(960, 0.0)(961, 0.0)(962, 0.0)(963, 0.0)(964, 0.0)(965, 0.0)(966, 0.0)(967, 0.0)(968, 0.0)(969, 0.0)(970, 0.0)(971, 0.0)(972, 0.0)(973, 0.0)(974, 0.0)(975, 0.0)(976, 0.0)(977, 0.0)(978, 0.0)(979, 0.0)(980, 0.0)(981, 0.0)(982, 0.0)(983, 0.0)(984, 0.0)(985, 0.0)(986, 0.0)(987, 0.0)(988, 0.0)(989, 0.0)(990, 0.0)(991, 0.0)(992, 0.0)(993, 0.0)(994, 0.0)(995, 0.0)(996, 0.0)(997, 0.0)(998, 0.0)(999, 0.0)(1000, 0.0)(1001, 0.0)(1002, 0.0)(1003, 0.0)(1004, 0.0)(1005, 0.0)(1006, 0.0)(1007, 0.0)(1008, 0.0)(1009, 0.0)(1010, 0.0)(1011, 0.0)(1012, 0.0)(1013, 0.0)(1014, 0.0)(1015, 0.0)(1016, 0.0)(1017, 0.0)(1018, 0.0)(1019, 0.0)(1020, 0.0)(1021, 0.0)(1022, 0.0)(1023, 0.0)(1024, 0.0)(1025, 0.0)(1026, 0.0)(1027, 0.0)(1028, 0.0)(1029, 0.0)(1030, 0.0)(1031, 0.0)(1032, 0.0)(1033, 0.0)(1034, 0.0)(1035, 0.0)(1036, 0.0)(1037, 0.0)(1038, 0.0)(1039, 0.0)(1040, 0.0)(1041, 0.0)(1042, 0.0)(1043, 0.0)(1044, 0.0)(1045, 0.0)(1046, 0.0)(1047, 0.0)(1048, 0.0)(1049, 0.0)(1050, 0.0)(1051, 0.0)(1052, 0.0)(1053, 0.0)(1054, 0.0)(1055, 0.0)(1056, 0.0)(1057, 0.0)(1058, 0.0)(1059, 0.0)(1060, 0.0)(1061, 0.0)(1062, 0.0)(1063, 0.0)(1064, 0.0)(1065, 0.0)(1066, 0.0)(1067, 0.0)(1068, 0.0)(1069, 0.0)(1070, 0.0)(1071, 0.0)(1072, 0.0)(1073, 0.0)(1074, 0.0)(1075, 0.0)(1076, 0.0)(1077, 0.0)(1078, 0.0)(1079, 0.0)(1080, 0.0)(1081, 0.0)(1082, 0.0)(1083, 0.0)(1084, 0.0)(1085, 0.0)(1086, 0.0)(1087, 0.0)(1088, 0.0)(1089, 0.0)(1090, 0.0)(1091, 0.0)(1092, 0.0)(1093, 0.0)(1094, 0.0)(1095, 0.0)(1096, 0.0)(1097, 0.0)(1098, 0.0)(1099, 0.0)(1100, 0.0)(1101, 0.0)(1102, 0.0)(1103, 0.0)(1104, 0.0)(1105, 0.0)(1106, 0.0)(1107, 0.0)(1108, 0.0)(1109, 0.0)(1110, 0.0)(1111, 0.0)(1112, 0.0)(1113, 0.0)(1114, 0.0)(1115, 0.0)(1116, 0.0)(1117, 0.0)(1118, 0.0)(1119, 0.0)(1120, 0.0)(1121, 0.0)(1122, 0.0)(1123, 0.0)(1124, 0.0)(1125, 0.0)(1126, 0.0)(1127, 0.0)(1128, 0.0)(1129, 0.0)(1130, 0.0)(1131, 0.0)(1132, 0.0)(1133, 0.0)(1134, 0.0)(1135, 0.0)(1136, 0.0)(1137, 0.0)(1138, 0.0)(1139, 0.0)(1140, 0.0)(1141, 0.0)(1142, 0.0)(1143, 0.0)(1144, 0.0)(1145, 0.0)(1146, 0.0)(1147, 0.0)(1148, 0.0)(1149, 0.0)(1150, 0.0)(1151, 0.0)(1152, 0.0)(1153, 0.0)(1154, 0.0)(1155, 0.0)(1156, 0.0)(1157, 0.0)(1158, 0.0)(1159, 0.0)(1160, 0.0)(1161, 0.0)(1162, 0.0)(1163, 0.0)(1164, 0.0)(1165, 0.0)(1166, 0.0)(1167, 0.0)(1168, 0.0)(1169, 0.0)(1170, 0.0)(1171, 0.0)(1172, 0.0)(1173, 0.0)(1174, 0.0)(1175, 0.0)(1176, 0.0)(1177, 0.0)(1178, 0.0)(1179, 0.0)(1180, 0.0)(1181, 0.0)(1182, 0.0)(1183, 0.0)(1184, 0.0)(1185, 0.0)(1186, 0.0)(1187, 0.0)(1188, 0.0)(1189, 0.0)(1190, 0.0)(1191, 0.0)(1192, 0.0)(1193, 0.0)(1194, 0.0)(1195, 0.0)(1196, 0.0)(1197, 0.0)(1198, 0.0)(1199, 0.0)(1200, 0.0)(1201, 0.0)(1202, 0.0)(1203, 0.0)(1204, 0.0)(1205, 0.0)(1206, 0.0)(1207, 0.0)(1208, 0.0)(1209, 0.0)(1210, 0.0)(1211, 0.0)(1212, 0.0)(1213, 0.0)(1214, 0.0)(1215, 0.0)(1216, 0.0)(1217, 0.0)(1218, 0.0)(1219, 0.0)(1220, 0.0)(1221, 0.0)(1222, 0.0)(1223, 0.0)(1224, 0.0)(1225, 0.0)(1226, 0.0)(1227, 0.0)(1228, 0.0)(1229, 0.0)(1230, 0.0)(1231, 0.0)(1232, 0.0)(1233, 0.0)(1234, 0.0)(1235, 0.0)(1236, 0.0)(1237, 0.0)(1238, 0.0)(1239, 0.0)(1240, 0.0)(1241, 0.0)(1242, 0.0)(1243, 0.0)(1244, 0.0)(1245, 0.0)(1246, 0.0)(1247, 0.0)(1248, 0.0)(1249, 0.0)(1250, 0.0)(1251, 0.0)(1252, 0.0)(1253, 0.0)(1254, 0.0)(1255, 0.0)(1256, 0.0)(1257, 0.0)(1258, 0.0)(1259, 0.0)(1260, 0.0)(1261, 0.0)(1262, 0.0)(1263, 0.0)(1264, 0.0)(1265, 0.0)(1266, 0.0)(1267, 0.0)(1268, 0.0)(1269, 0.0)(1270, 0.0)(1271, 0.0)(1272, 0.0)(1273, 0.0)(1274, 0.0)(1275, 0.0)(1276, 0.0)(1277, 0.0)(1278, 0.0)(1279, 0.0)(1280, 0.0)(1281, 0.0)(1282, 0.0)(1283, 0.0)(1284, 0.0)(1285, 0.0)(1286, 0.0)(1287, 0.0)(1288, 0.0)(1289, 0.0)(1290, 0.0)(1291, 0.0)(1292, 0.0)(1293, 0.0)(1294, 0.0)(1295, 0.0)(1296, 0.0)(1297, 0.0)(1298, 0.0)(1299, 0.0)(1300, 0.0)(1301, 0.0)(1302, 0.0)(1303, 0.0)(1304, 0.0)(1305, 0.0)(1306, 0.0)(1307, 0.0)(1308, 0.0)(1309, 0.0)(1310, 0.0)(1311, 0.0)(1312, 0.0)(1313, 0.0)(1314, 0.0)(1315, 0.0)(1316, 0.0)(1317, 0.0)(1318, 0.0)(1319, 0.0)(1320, 0.0)(1321, 0.0)(1322, 0.0)(1323, 0.0)(1324, 0.0)(1325, 0.0)(1326, 0.0)(1327, 0.0)(1328, 0.0)(1329, 0.0)(1330, 0.0)(1331, 0.0)(1332, 0.0)(1333, 0.0)(1334, 0.0)(1335, 0.0)(1336, 0.0)(1337, 0.0)(1338, 0.0)(1339, 0.0)(1340, 0.0)(1341, 0.0)(1342, 0.0)(1343, 0.0)(1344, 0.0)(1345, 0.0)(1346, 0.0)(1347, 0.0)(1348, 0.0)(1349, 0.0)(1350, 0.0)(1351, 0.0)(1352, 0.0)(1353, 0.0)(1354, 0.0)(1355, 0.0)(1356, 0.0)(1357, 0.0)(1358, 0.0)(1359, 0.0)(1360, 0.0)(1361, 0.0)(1362, 0.0)(1363, 0.0)(1364, 0.0)(1365, 0.0)(1366, 0.0)(1367, 0.0)(1368, 0.0)(1369, 0.0)(1370, 0.0)(1371, 0.0)(1372, 0.0)(1373, 0.0)(1374, 0.0)(1375, 0.0)(1376, 0.0)(1377, 0.0)(1378, 0.0)(1379, 0.0)(1380, 0.0)(1381, 0.0)(1382, 0.0)(1383, 0.0)(1384, 0.0)(1385, 0.0)(1386, 0.0)(1387, 0.0)(1388, 0.0)(1389, 0.0)(1390, 0.0)(1391, 0.0)(1392, 0.0)(1393, 0.0)(1394, 0.0)(1395, 0.0)(1396, 0.0)(1397, 0.0)(1398, 0.0)(1399, 0.0)(1400, 0.0)(1401, 0.0)(1402, 0.0)(1403, 0.0)(1404, 0.0)(1405, 0.0)(1406, 0.0)(1407, 0.0)(1408, 0.0)(1409, 0.0)(1410, 0.0)(1411, 0.0)(1412, 0.0)(1413, 0.0)(1414, 0.0)(1415, 0.0)(1416, 0.0)(1417, 0.0)(1418, 0.0)(1419, 0.0)(1420, 0.0)(1421, 0.0)(1422, 0.0)(1423, 0.0)(1424, 0.0)(1425, 0.0)(1426, 0.0)(1427, 0.0)(1428, 0.0)(1429, 0.0)(1430, 0.0)(1431, 0.0)(1432, 0.0)(1433, 0.0)(1434, 0.0)(1435, 0.0)(1436, 0.0)(1437, 0.0)(1438, 0.0)(1439, 0.0)(1440, 0.0)(1441, 0.0)(1442, 0.0)(1443, 0.0)(1444, 0.0)(1445, 0.0)(1446, 0.0)(1447, 0.0)(1448, 0.0)(1449, 0.0)(1450, 0.0)(1451, 0.0)(1452, 0.0)(1453, 0.0)(1454, 0.0)(1455, 0.0)(1456, 0.0)(1457, 0.0)(1458, 0.0)(1459, 0.0)(1460, 0.0)(1461, 0.0)(1462, 0.0)(1463, 0.0)(1464, 0.0)(1465, 0.0)(1466, 0.0)(1467, 0.0)(1468, 0.0)(1469, 0.0)(1470, 0.0)(1471, 0.0)(1472, 0.0)(1473, 0.0)(1474, 0.0)(1475, 0.0)(1476, 0.0)(1477, 0.0)(1478, 0.0)(1479, 0.0)(1480, 0.0)(1481, 0.0)(1482, 0.0)(1483, 0.0)(1484, 0.0)(1485, 0.0)(1486, 0.0)(1487, 0.0)(1488, 0.0)(1489, 0.0)(1490, 0.0)(1491, 0.0)(1492, 0.0)(1493, 0.0)(1494, 0.0)(1495, 0.0)(1496, 0.0)(1497, 0.0)(1498, 0.0)(1499, 0.0)(1500, 0.0)(1501, 0.0)(1502, 0.0)(1503, 0.0)(1504, 0.0)(1505, 0.0)(1506, 0.0)(1507, 0.0)(1508, 0.0)(1509, 0.0)(1510, 0.0)(1511, 0.0)(1512, 0.0)(1513, 0.0)(1514, 0.0)(1515, 0.0)(1516, 0.0)(1517, 0.0)(1518, 0.0)(1519, 0.0)};
            \addplot+[name path=E,black,line width=.1pt] coordinates {(0, 0.0)(1, 0.0)(2, 0.0)(3, 0.0)(4, 0.0)(5, 0.0)(6, 0.0)(7, 0.0)(8, 0.0)(9, 0.0)(10, 0.0)(11, 0.0)(12, 0.0)(13, 0.0)(14, 0.0)(15, 0.0)(16, 0.0)(17, 0.0)(18, 0.0)(19, 0.0)(20, 0.0)(21, 0.0)(22, 0.0)(23, 0.0)(24, 0.0)(25, 0.0)(26, 0.0)(27, 0.0)(28, 0.0)(29, 0.0)(30, 0.0)(31, 0.0)(32, 0.0)(33, 0.0)(34, 0.0)(35, 0.0)(36, 0.0)(37, 0.0)(38, 0.0)(39, 0.0)(40, 0.0)(41, 0.0)(42, 0.0)(43, 0.0)(44, 0.0)(45, 0.0)(46, 0.0)(47, 0.0)(48, 0.0)(49, 0.0)(50, 0.0)(51, 0.0)(52, 0.0)(53, 0.0)(54, 0.0)(55, 0.0)(56, 0.0)(57, 0.0)(58, 0.0)(59, 0.0)(60, 0.0)(61, 0.0)(62, 0.0)(63, 0.0)(64, 0.0)(65, 0.0)(66, 0.0)(67, 0.0)(68, 0.0)(69, 0.0)(70, 0.0)(71, 0.0)(72, 0.0)(73, 0.0)(74, 0.0)(75, 0.0)(76, 0.0)(77, 0.0)(78, 0.0)(79, 0.0)(80, 0.0)(81, 0.0)(82, 0.0)(83, 0.0)(84, 0.0)(85, 0.0)(86, 0.0)(87, 0.0)(88, 0.0)(89, 0.0)(90, 0.0)(91, 0.0)(92, 0.0)(93, 0.0)(94, 0.0)(95, 0.0)(96, 0.0)(97, 0.0)(98, 0.0)(99, 0.0)(100, 0.0)(101, 0.0)(102, 0.0)(103, 0.0)(104, 0.0)(105, 0.0)(106, 0.0)(107, 0.0)(108, 0.0)(109, 0.0)(110, 0.0)(111, 0.0)(112, 0.0)(113, 0.0)(114, 0.0)(115, 0.0)(116, 0.0)(117, 0.0)(118, 0.0)(119, 0.0)(120, 0.0)(121, 0.0)(122, 0.0)(123, 0.0)(124, 0.0)(125, 0.0)(126, 0.0)(127, 0.0)(128, 0.0)(129, 0.0)(130, 0.0)(131, 0.0)(132, 0.0)(133, 0.0)(134, 0.0)(135, 0.0)(136, 0.0)(137, 0.0)(138, 0.0)(139, 0.0)(140, 0.0)(141, 0.0)(142, 0.0)(143, 0.0)(144, 0.0)(145, 0.0)(146, 0.0)(147, 0.0)(148, 0.0)(149, 0.0)(150, 0.0)(151, 0.0)(152, 0.0)(153, 0.0)(154, 0.0)(155, 0.0)(156, 0.0)(157, 0.0)(158, 0.0)(159, 0.0)(160, 0.0)(161, 0.0)(162, 0.0)(163, 0.0)(164, 0.0)(165, 0.0)(166, 0.0)(167, 0.0)(168, 0.0)(169, 0.0)(170, 0.0)(171, 0.0)(172, 0.0)(173, 0.0)(174, 0.0)(175, 0.0)(176, 0.0)(177, 0.0)(178, 0.0)(179, 0.0)(180, 0.0)(181, 0.0)(182, 0.0)(183, 0.0)(184, 0.0)(185, 0.0)(186, 0.0)(187, 0.0)(188, 0.0)(189, 0.0)(190, 0.0)(191, 0.0)(192, 0.0)(193, 0.0)(194, 0.0)(195, 0.0)(196, 0.0)(197, 0.0)(198, 0.0)(199, 0.0)(200, 0.0)(201, 0.0)(202, 0.0)(203, 0.0)(204, 0.0)(205, 0.0)(206, 0.0)(207, 0.0)(208, 0.0)(209, 0.0)(210, 0.0)(211, 0.0)(212, 0.0)(213, 0.0)(214, 0.0)(215, 0.0)(216, 0.0)(217, 0.0)(218, 0.0)(219, 0.0)(220, 0.0)(221, 0.0)(222, 0.0)(223, 0.0)(224, 0.0)(225, 0.0)(226, 0.0)(227, 0.0)(228, 0.0)(229, 0.0)(230, 0.0)(231, 0.0)(232, 0.0)(233, 0.0)(234, 0.0)(235, 0.0)(236, 0.0)(237, 0.0)(238, 0.0)(239, 0.0)(240, 0.0)(241, 0.0)(242, 0.0)(243, 0.0)(244, 0.0)(245, 0.0)(246, 0.0)(247, 0.0)(248, 0.0)(249, 0.0)(250, 0.0)(251, 0.0)(252, 0.0)(253, 0.0)(254, 0.0)(255, 0.0)(256, 0.0)(257, 0.0)(258, 0.0)(259, 0.0)(260, 0.0)(261, 0.0)(262, 0.0)(263, 0.0)(264, 0.0)(265, 0.0)(266, 0.0)(267, 0.0)(268, 0.0)(269, 0.0)(270, 0.0)(271, 0.0)(272, 0.0)(273, 0.0)(274, 0.0)(275, 0.0)(276, 0.0)(277, 0.0)(278, 0.0)(279, 0.0)(280, 0.0)(281, 0.0)(282, 0.0)(283, 0.0)(284, 0.0)(285, 0.0)(286, 0.0)(287, 0.0)(288, 0.0)(289, 0.0)(290, 0.0)(291, 0.0)(292, 0.0)(293, 0.0)(294, 0.0)(295, 0.0)(296, 0.0)(297, 0.0)(298, 0.0)(299, 0.0)(300, 0.0)(301, 0.0)(302, 0.0)(303, 0.0)(304, 0.0)(305, 0.0)(306, 0.0)(307, 0.0)(308, 0.0)(309, 0.0)(310, 0.0)(311, 0.0)(312, 0.0)(313, 0.0)(314, 0.0)(315, 0.0)(316, 0.0)(317, 0.0)(318, 0.0)(319, 0.0)(320, 0.0)(321, 0.0)(322, 0.0)(323, 0.0)(324, 0.0)(325, 0.0)(326, 0.0)(327, 0.0)(328, 0.0)(329, 0.0)(330, 0.0)(331, 0.0)(332, 0.0)(333, 0.0)(334, 0.0)(335, 0.0)(336, 0.0)(337, 0.0)(338, 0.0)(339, 0.0)(340, 0.0)(341, 0.0)(342, 0.0)(343, 0.0)(344, 0.0)(345, 0.0)(346, 0.0)(347, 0.0)(348, 0.0)(349, 0.0)(350, 0.0)(351, 0.0)(352, 0.0)(353, 0.0)(354, 0.0)(355, 0.0)(356, 0.0)(357, 0.0)(358, 0.0)(359, 0.0)(360, 0.0)(361, 0.0)(362, 0.0)(363, 0.0)(364, 0.0)(365, 0.0)(366, 0.0)(367, 0.0)(368, 0.0)(369, 0.0)(370, 0.0)(371, 0.0)(372, 0.0)(373, 0.0)(374, 0.0)(375, 0.0)(376, 0.0)(377, 0.0)(378, 0.0)(379, 0.0)(380, 0.0)(381, 0.0)(382, 0.0)(383, 0.0)(384, 0.0)(385, 0.0)(386, 0.0)(387, 0.0)(388, 0.0)(389, 0.0)(390, 0.0)(391, 0.0)(392, 0.0)(393, 0.0)(394, 0.0)(395, 0.0)(396, 0.0)(397, 0.0)(398, 0.0)(399, 0.0)(400, 0.0)(401, 0.0)(402, 0.0)(403, 0.0)(404, 0.0)(405, 0.0)(406, 0.0)(407, 0.0)(408, 0.0)(409, 0.0)(410, 0.0)(411, 0.0)(412, 0.0)(413, 0.0)(414, 0.0)(415, 0.0)(416, 0.0)(417, 0.0)(418, 0.0)(419, 0.0)(420, 0.0)(421, 0.0)(422, 0.0)(423, 0.0)(424, 0.0)(425, 0.0)(426, 0.0)(427, 0.0)(428, 0.0)(429, 0.0)(430, 0.0)(431, 0.0)(432, 0.0)(433, 0.0)(434, 0.0)(435, 0.0)(436, 0.0)(437, 0.0)(438, 0.0)(439, 0.0)(440, 0.0)(441, 0.0)(442, 0.0)(443, 0.0)(444, 0.0)(445, 0.0)(446, 0.0)(447, 0.0)(448, 0.0)(449, 0.0)(450, 0.0)(451, 0.0)(452, 0.0)(453, 0.0)(454, 0.0)(455, 0.0)(456, 0.0)(457, 0.0)(458, 0.0)(459, 0.0)(460, 0.0)(461, 0.0)(462, 0.0)(463, 0.0)(464, 0.0)(465, 0.0)(466, 0.0)(467, 0.0)(468, 0.0)(469, 0.0)(470, 0.0)(471, 0.0)(472, 0.0)(473, 0.0)(474, 0.0)(475, 0.0)(476, 0.0)(477, 0.0)(478, 0.0)(479, 0.0)(480, 0.0)(481, 0.0)(482, 0.0)(483, 0.0)(484, 0.0)(485, 0.0)(486, 0.0)(487, 0.0)(488, 0.0)(489, 0.0)(490, 0.0)(491, 0.0)(492, 0.0)(493, 0.0)(494, 0.0)(495, 0.0)(496, 0.0)(497, 0.0)(498, 0.0)(499, 0.0)(500, 0.0)(501, 0.0)(502, 0.0)(503, 0.0)(504, 0.0)(505, 0.0)(506, 0.0)(507, 0.0)(508, 0.0)(509, 0.0)(510, 0.0)(511, 0.0)(512, 0.0)(513, 0.0)(514, 0.0)(515, 0.0)(516, 0.0)(517, 0.0)(518, 0.0)(519, 0.0)(520, 0.0)(521, 0.0)(522, 0.0)(523, 0.0)(524, 0.0)(525, 0.0)(526, 0.0)(527, 0.0)(528, 0.0)(529, 0.0)(530, 0.0)(531, 0.0)(532, 0.0)(533, 0.0)(534, 0.0)(535, 0.0)(536, 0.0)(537, 0.0)(538, 0.0)(539, 0.0)(540, 0.0)(541, 0.0)(542, 0.0)(543, 0.0)(544, 0.0)(545, 0.0)(546, 0.0)(547, 0.0)(548, 0.0)(549, 0.0)(550, 0.0)(551, 0.0)(552, 0.0)(553, 0.0)(554, 0.0)(555, 0.0)(556, 0.0)(557, 0.0)(558, 0.0)(559, 0.0)(560, 0.0)(561, 0.0)(562, 0.0)(563, 0.0)(564, 0.0)(565, 0.0)(566, 0.0)(567, 0.0)(568, 0.0)(569, 0.0)(570, 0.0)(571, 0.0)(572, 0.0)(573, 0.0)(574, 0.0)(575, 0.0)(576, 0.0)(577, 0.0)(578, 0.0)(579, 0.0)(580, 0.0)(581, 0.0)(582, 0.0)(583, 0.0)(584, 0.0)(585, 0.0)(586, 0.0)(587, 0.0)(588, 0.0)(589, 0.0)(590, 0.0)(591, 0.0)(592, 0.0)(593, 0.0)(594, 0.0)(595, 0.0)(596, 0.0)(597, 0.0)(598, 0.0)(599, 0.0)(600, 0.0)(601, 0.0)(602, 0.0)(603, 0.0)(604, 0.0)(605, 0.0)(606, 0.0)(607, 0.0)(608, 0.0)(609, 0.0)(610, 0.0)(611, 0.0)(612, 0.0)(613, 0.0)(614, 0.0)(615, 0.0)(616, 0.0)(617, 0.0)(618, 0.0)(619, 0.0)(620, 0.0)(621, 0.0)(622, 0.0)(623, 0.0)(624, 0.0)(625, 0.0)(626, 0.0)(627, 0.0)(628, 0.0)(629, 0.0)(630, 0.0)(631, 0.0)(632, 0.0)(633, 0.0)(634, 0.0)(635, 0.0)(636, 0.0)(637, 0.0)(638, 0.0)(639, 0.0)(640, 0.0)(641, 0.0)(642, 0.0)(643, 0.0)(644, 0.0)(645, 0.0)(646, 0.0)(647, 0.0)(648, 0.0)(649, 0.0)(650, 0.0)(651, 0.0)(652, 0.0)(653, 0.0)(654, 0.0)(655, 0.0)(656, 0.0)(657, 0.0)(658, 0.0)(659, 0.0)(660, 0.0)(661, 0.0)(662, 0.0)(663, 0.0)(664, 0.0)(665, 0.0)(666, 0.0)(667, 0.0)(668, 0.0)(669, 0.0)(670, 0.0)(671, 0.0)(672, 0.0)(673, 0.0)(674, 0.0)(675, 0.0)(676, 0.0)(677, 0.0)(678, 0.0)(679, 0.0)(680, 0.0)(681, 0.0)(682, 0.0)(683, 0.0)(684, 0.0)(685, 0.0)(686, 0.0)(687, 0.0)(688, 0.0)(689, 0.0)(690, 0.0)(691, 0.0)(692, 0.0)(693, 0.0)(694, 0.0)(695, 0.0)(696, 0.0)(697, 0.0)(698, 0.0)(699, 0.0)(700, 0.0)(701, 0.0)(702, 0.0)(703, 0.0)(704, 0.0)(705, 0.0)(706, 0.0)(707, 0.0)(708, 0.0)(709, 0.0)(710, 0.0)(711, 0.0)(712, 0.0)(713, 0.0)(714, 0.0)(715, 0.0)(716, 0.0)(717, 0.0)(718, 0.0)(719, 0.0)(720, 0.0)(721, 0.0)(722, 0.0)(723, 0.0)(724, 0.0)(725, 0.0)(726, 0.0)(727, 0.0)(728, 0.0)(729, 0.0)(730, 0.0)(731, 0.0)(732, 0.0)(733, 0.0)(734, 0.0)(735, 0.0)(736, 0.0)(737, 0.0)(738, 0.0)(739, 0.0)(740, 0.0)(741, 0.0)(742, 0.0)(743, 0.0)(744, 0.0)(745, 0.0)(746, 0.0)(747, 0.0)(748, 0.0)(749, 0.0)(750, 0.0)(751, 0.0)(752, 0.0)(753, 0.0)(754, 0.0)(755, 0.0)(756, 0.0)(757, 0.0)(758, 0.0)(759, 0.0)(760, 0.0)(761, 0.0)(762, 0.0)(763, 0.0)(764, 0.0)(765, 0.0)(766, 0.0)(767, 0.0)(768, 0.0)(769, 0.0)(770, 0.0)(771, 0.0)(772, 0.0)(773, 0.0)(774, 0.0)(775, 0.0)(776, 0.0)(777, 0.0)(778, 0.0)(779, 0.0)(780, 0.0)(781, 0.0)(782, 0.0)(783, 0.0)(784, 0.0)(785, 0.0)(786, 0.0)(787, 0.0)(788, 0.0)(789, 0.0)(790, 0.0)(791, 0.0)(792, 0.0)(793, 0.0)(794, 0.0)(795, 0.0)(796, 0.0)(797, 0.0)(798, 0.0)(799, 0.0)(800, 0.0)(801, 0.0)(802, 0.0)(803, 0.0)(804, 0.0)(805, 0.0)(806, 0.0)(807, 0.0)(808, 0.0)(809, 0.0)(810, 0.0)(811, 0.0)(812, 0.0)(813, 0.0)(814, 0.0)(815, 0.0)(816, 0.0)(817, 0.0)(818, 0.0)(819, 0.0)(820, 0.0)(821, 0.0)(822, 0.0)(823, 0.0)(824, 0.0)(825, 0.0)(826, 0.0)(827, 0.0)(828, 0.0)(829, 0.0)(830, 0.0)(831, 0.0)(832, 0.0)(833, 0.0)(834, 0.0)(835, 0.0)(836, 0.0)(837, 0.0)(838, 0.0)(839, 0.0)(840, 0.0)(841, 0.0)(842, 0.0)(843, 0.0)(844, 0.0)(845, 0.0)(846, 0.0)(847, 0.0)(848, 0.0)(849, 0.0)(850, 0.0)(851, 0.0)(852, 0.0)(853, 0.0)(854, 0.0)(855, 0.0)(856, 0.0)(857, 0.0)(858, 0.0)(859, 0.0)(860, 0.0)(861, 0.0)(862, 0.0)(863, 0.0)(864, 0.0)(865, 0.0)(866, 0.0)(867, 0.0)(868, 0.0)(869, 0.0)(870, 0.0)(871, 0.0)(872, 0.0)(873, 0.0)(874, 0.0)(875, 0.0)(876, 0.0)(877, 0.0)(878, 0.0)(879, 0.0)(880, 0.0)(881, 0.0)(882, 0.0)(883, 0.0)(884, 0.0)(885, 0.0)(886, 0.0)(887, 0.0)(888, 0.0)(889, 0.0)(890, 0.0)(891, 0.0)(892, 0.0)(893, 0.0)(894, 0.0)(895, 0.0)(896, 0.0)(897, 0.0)(898, 0.0)(899, 0.0)(900, 0.0)(901, 0.0)(902, 0.0)(903, 0.0)(904, 0.0)(905, 0.0)(906, 0.0)(907, 0.0)(908, 0.0)(909, 0.0)(910, 0.0)(911, 0.0)(912, 0.0)(913, 0.0)(914, 0.0)(915, 0.0)(916, 0.0)(917, 0.0)(918, 0.0)(919, 0.0)(920, 0.0)(921, 0.0)(922, 0.0)(923, 0.0)(924, 0.0)(925, 0.0)(926, 0.0)(927, 0.0)(928, 0.0)(929, 0.0)(930, 0.0)(931, 0.0)(932, 0.0)(933, 0.0)(934, 0.0)(935, 0.0)(936, 0.0)(937, 0.0)(938, 0.0)(939, 0.0)(940, 0.0)(941, 0.0)(942, 0.0)(943, 0.0)(944, 0.0)(945, 0.0)(946, 0.0)(947, 0.0)(948, 0.0)(949, 0.0)(950, 0.0)(951, 0.0)(952, 0.0)(953, 0.0)(954, 0.0)(955, 0.0)(956, 0.0)(957, 0.0)(958, 0.0)(959, 0.0)(960, 0.0)(961, 0.0)(962, 0.0)(963, 0.0)(964, 0.0)(965, 0.0)(966, 0.0)(967, 0.0)(968, 0.0)(969, 0.0)(970, 0.0)(971, 0.0)(972, 0.0)(973, 0.0)(974, 0.0)(975, 0.0)(976, 0.0)(977, 0.0)(978, 0.0)(979, 0.0)(980, 0.0)(981, 0.0)(982, 0.0)(983, 0.0)(984, 0.0)(985, 0.0)(986, 0.0)(987, 0.0)(988, 0.0)(989, 0.0)(990, 0.0)(991, 0.0)(992, 0.0)(993, 0.0)(994, 0.0)(995, 0.0)(996, 0.0)(997, 0.0)(998, 0.0)(999, 0.0)(1000, 0.0)(1001, 0.0)(1002, 0.0)(1003, 0.0)(1004, 0.0)(1005, 0.0)(1006, 0.0)(1007, 0.0)(1008, 0.0)(1009, 0.0)(1010, 0.0)(1011, 0.0)(1012, 0.0)(1013, 0.0)(1014, 0.0)(1015, 0.0)(1016, 0.0)(1017, 0.0)(1018, 0.0)(1019, 0.0)(1020, 0.0)(1021, 0.0)(1022, 0.0)(1023, 0.0)(1024, 0.0)(1025, 0.0)(1026, 0.0)(1027, 0.0)(1028, 0.0)(1029, 0.0)(1030, 0.0)(1031, 0.0)(1032, 0.0)(1033, 0.0)(1034, 0.0)(1035, 0.0)(1036, 0.0)(1037, 0.0)(1038, 0.0)(1039, 0.0)(1040, 0.0)(1041, 0.0)(1042, 0.0)(1043, 0.0)(1044, 0.0)(1045, 0.0)(1046, 0.0)(1047, 0.0)(1048, 0.0)(1049, 0.0)(1050, 0.0)(1051, 0.0)(1052, 0.0)(1053, 0.0)(1054, 0.0)(1055, 0.0)(1056, 0.0)(1057, 0.0)(1058, 0.0)(1059, 0.0)(1060, 0.0)(1061, 0.0)(1062, 0.0)(1063, 0.0)(1064, 0.0)(1065, 0.0)(1066, 0.0)(1067, 0.0)(1068, 0.0)(1069, 0.0)(1070, 0.0)(1071, 0.0)(1072, 0.0)(1073, 0.0)(1074, 0.0)(1075, 0.0)(1076, 0.0)(1077, 0.0)(1078, 0.0)(1079, 0.0)(1080, 0.0)(1081, 0.0)(1082, 0.0)(1083, 0.0)(1084, 0.0)(1085, 0.0)(1086, 0.0)(1087, 0.0)(1088, 0.0)(1089, 0.0)(1090, 0.0)(1091, 0.0)(1092, 0.0)(1093, 0.0)(1094, 0.0)(1095, 0.0)(1096, 0.0)(1097, 0.0)(1098, 0.0)(1099, 0.0)(1100, 0.0)(1101, 0.0)(1102, 0.0)(1103, 0.0)(1104, 0.0)(1105, 0.0)(1106, 0.0)(1107, 0.0)(1108, 0.0)(1109, 0.0)(1110, 0.0)(1111, 0.0)(1112, 0.0)(1113, 0.0)(1114, 0.0)(1115, 0.0)(1116, 0.0)(1117, 0.0)(1118, 0.0)(1119, 0.0)(1120, 0.0)(1121, 0.0)(1122, 0.0)(1123, 0.0)(1124, 0.0)(1125, 0.0)(1126, 0.0)(1127, 0.0)(1128, 0.0)(1129, 0.0)(1130, 0.0)(1131, 0.0)(1132, 0.0)(1133, 0.0)(1134, 0.0)(1135, 0.0)(1136, 0.0)(1137, 0.0)(1138, 0.0)(1139, 0.0)(1140, 0.0)(1141, 0.0)(1142, 0.0)(1143, 0.0)(1144, 0.0)(1145, 0.0)(1146, 0.0)(1147, 0.0)(1148, 0.0)(1149, 0.0)(1150, 0.0)(1151, 0.0)(1152, 0.0)(1153, 0.0)(1154, 0.0)(1155, 0.0)(1156, 0.0)(1157, 0.0)(1158, 0.0)(1159, 0.0)(1160, 0.0)(1161, 0.0)(1162, 0.0)(1163, 0.0)(1164, 0.0)(1165, 0.0)(1166, 0.0)(1167, 0.0)(1168, 0.0)(1169, 0.0)(1170, 0.0)(1171, 0.0)(1172, 0.0)(1173, 0.0)(1174, 0.0)(1175, 0.0)(1176, 0.0)(1177, 0.0)(1178, 0.0)(1179, 0.0)(1180, 0.0)(1181, 0.0)(1182, 0.0)(1183, 0.0)(1184, 0.0)(1185, 0.0)(1186, 0.0)(1187, 0.0)(1188, 0.0)(1189, 0.0)(1190, 0.0)(1191, 0.0)(1192, 0.0)(1193, 0.0)(1194, 0.0)(1195, 0.0)(1196, 0.0)(1197, 0.0)(1198, 0.0)(1199, 0.0)(1200, 0.0)(1201, 0.0)(1202, 0.0)(1203, 0.0)(1204, 0.0)(1205, 0.0)(1206, 0.0)(1207, 0.0)(1208, 0.0)(1209, 0.0)(1210, 0.0)(1211, 0.0)(1212, 0.0)(1213, 0.0)(1214, 0.0)(1215, 0.0)(1216, 0.0)(1217, 0.0)(1218, 0.0)(1219, 0.0)(1220, 0.0)(1221, 0.0)(1222, 0.0)(1223, 0.0)(1224, 0.0)(1225, 0.0)(1226, 0.0)(1227, 0.0)(1228, 0.0)(1229, 0.0)(1230, 0.0)(1231, 0.0)(1232, 0.0)(1233, 0.0)(1234, 0.0)(1235, 0.0)(1236, 0.0)(1237, 0.0)(1238, 0.0)(1239, 0.0)(1240, 0.0)(1241, 0.0)(1242, 0.0)(1243, 0.0)(1244, 0.0)(1245, 0.0)(1246, 0.0)(1247, 0.0)(1248, 0.0)(1249, 0.0)(1250, 0.0)(1251, 0.0)(1252, 0.0)(1253, 0.0)(1254, 0.0)(1255, 0.0)(1256, 0.0)(1257, 0.0)(1258, 0.0)(1259, 0.0)(1260, 0.0)(1261, 0.0)(1262, 0.0)(1263, 0.0)(1264, 0.0)(1265, 0.0)(1266, 0.0)(1267, 0.0)(1268, 0.0)(1269, 0.0)(1270, 0.0)(1271, 0.0)(1272, 0.0)(1273, 0.0)(1274, 0.0)(1275, 0.0)(1276, 0.0)(1277, 0.0)(1278, 0.0)(1279, 0.0)(1280, 0.0)(1281, 0.0)(1282, 0.0)(1283, 0.0)(1284, 0.0)(1285, 0.0)(1286, 0.0)(1287, 0.0)(1288, 0.0)(1289, 0.0)(1290, 0.0)(1291, 0.0)(1292, 0.0)(1293, 0.0)(1294, 0.0)(1295, 0.0)(1296, 0.0)(1297, 0.0)(1298, 0.0)(1299, 0.0)(1300, 0.0)(1301, 0.0)(1302, 0.0)(1303, 0.0)(1304, 0.0)(1305, 0.0)(1306, 0.0)(1307, 0.0)(1308, 0.0)(1309, 0.0)(1310, 0.0)(1311, 0.0)(1312, 0.0)(1313, 0.0)(1314, 0.0)(1315, 0.0)(1316, 0.0)(1317, 0.0)(1318, 0.0)(1319, 0.0)(1320, 0.0)(1321, 0.0)(1322, 0.0)(1323, 0.0)(1324, 0.0)(1325, 0.0)(1326, 0.0)(1327, 0.0)(1328, 0.0)(1329, 0.0)(1330, 0.0)(1331, 0.0)(1332, 0.0)(1333, 0.0)(1334, 0.0)(1335, 0.0)(1336, 0.0)(1337, 0.0)(1338, 0.0)(1339, 0.0)(1340, 0.0)(1341, 0.0)(1342, 0.0)(1343, 0.0)(1344, 0.0)(1345, 0.0)(1346, 0.0)(1347, 0.0)(1348, 0.0)(1349, 0.0)(1350, 0.0)(1351, 0.0)(1352, 0.0)(1353, 0.0)(1354, 0.0)(1355, 0.0)(1356, 0.0)(1357, 0.0)(1358, 0.0)(1359, 0.0)(1360, 0.0)(1361, 0.0)(1362, 0.0)(1363, 0.0)(1364, 0.0)(1365, 0.0)(1366, 0.0)(1367, 0.0)(1368, 0.0)(1369, 0.0)(1370, 0.0)(1371, 0.0)(1372, 0.0)(1373, 0.0)(1374, 0.0)(1375, 0.0)(1376, 0.0)(1377, 0.0)(1378, 0.0)(1379, 0.0)(1380, 0.0)(1381, 0.0)(1382, 0.0)(1383, 0.0)(1384, 0.0)(1385, 0.0)(1386, 0.0)(1387, 0.0)(1388, 0.0)(1389, 0.0)(1390, 0.0)(1391, 0.0)(1392, 0.0)(1393, 0.0)(1394, 0.0)(1395, 0.0)(1396, 0.0)(1397, 0.0)(1398, 0.0)(1399, 0.0)(1400, 0.0)(1401, 0.0)(1402, 0.0)(1403, 0.0)(1404, 0.0)(1405, 0.0)(1406, 0.0)(1407, 0.0)(1408, 0.0)(1409, 0.0)(1410, 0.0)(1411, 0.0)(1412, 0.0)(1413, 0.0)(1414, 0.0)(1415, 0.0)(1416, 0.0)(1417, 0.0)(1418, 0.0)(1419, 0.0)(1420, 0.0)(1421, 0.0)(1422, 0.0)(1423, 0.0)(1424, 0.0)(1425, 0.0)(1426, 0.0)(1427, 0.0)(1428, 0.0)(1429, 0.0)(1430, 0.0)(1431, 0.0)(1432, 0.0)(1433, 0.0)(1434, 0.0)(1435, 0.0)(1436, 0.0)(1437, 0.0)(1438, 0.0)(1439, 0.0)(1440, 0.0)(1441, 0.0)(1442, 0.0)(1443, 0.0)(1444, 0.0)(1445, 0.0)(1446, 0.0)(1447, 0.0)(1448, 0.0)(1449, 0.0)(1450, 0.0)(1451, 0.0)(1452, 0.0)(1453, 0.0)(1454, 0.0)(1455, 0.0)(1456, 0.0)(1457, 0.0)(1458, 0.0)(1459, 0.0)(1460, 0.0)(1461, 0.0)(1462, 0.0)(1463, 0.0)(1464, 0.0)(1465, 0.0)(1466, 0.0)(1467, 0.0)(1468, 0.0)(1469, 0.0)(1470, 0.0)(1471, 0.0)(1472, 0.0)(1473, 0.0)(1474, 0.0)(1475, 0.0)(1476, 0.0)(1477, 0.0)(1478, 0.0)(1479, 0.0)(1480, 0.0)(1481, 0.0)(1482, 0.0)(1483, 0.0)(1484, 0.0)(1485, 0.0)(1486, 0.0)(1487, 0.0)(1488, 0.0)(1489, 0.0)(1490, 0.0)(1491, 0.0)(1492, 0.0)(1493, 0.0)(1494, 0.0)(1495, 0.0)(1496, 0.0)(1497, 0.0)(1498, 0.0)(1499, 0.0)(1500, 0.0)(1501, 0.0)(1502, 0.0)(1503, 0.0)(1504, 0.0)(1505, 0.0)(1506, 0.0)(1507, 0.0)(1508, 0.0)(1509, 0.0)(1510, 0.0)(1511, 0.0)(1512, 0.0)(1513, 0.0)(1514, 0.0)(1515, 0.0)(1516, 0.0)(1517, 0.0)(1518, 0.0)(1519, 0.0)};

            \addplot+[name path=F,black,line width=.1pt] coordinates {(0, 0.0)(1, 0.0)(2, 0.0)(3, 0.0)(4, 0.0)(5, 0.0)(6, 0.0)(7, 0.0)(8, 0.0)(9, 0.0)(10, 0.0)(11, 0.0)(12, 0.0)(13, 0.0)(14, 0.0)(15, 0.0)(16, 0.0)(17, 0.0)(18, 0.0)(19, 0.0)(20, 0.0)(21, 0.0)(22, 0.0)(23, 0.0)(24, 0.0)(25, 0.0)(26, 0.0)(27, 0.0)(28, 0.0)(29, 0.0)(30, 0.0)(31, 0.0)(32, 0.0)(33, 0.0)(34, 0.0)(35, 0.0)(36, 0.0)(37, 0.0)(38, 0.0)(39, 0.0)(40, 0.0)(41, 0.0)(42, 0.0)(43, 0.0)(44, 0.0)(45, 0.0)(46, 0.0)(47, 0.0)(48, 0.0)(49, 0.0)(50, 0.0)(51, 0.0)(52, 0.0)(53, 0.0)(54, 0.0)(55, 0.0)(56, 0.0)(57, 0.0)(58, 0.0)(59, 0.0)(60, 0.0)(61, 0.0)(62, 0.0)(63, 0.0)(64, 0.0)(65, 0.0)(66, 0.0)(67, 0.0)(68, 0.0)(69, 0.0)(70, 0.0)(71, 0.0)(72, 0.0)(73, 0.0)(74, 0.0)(75, 0.0)(76, 0.0)(77, 0.0)(78, 0.0)(79, 0.0)(80, 0.0)(81, 0.0)(82, 0.0)(83, 0.0)(84, 0.0)(85, 0.0)(86, 0.0)(87, 0.0)(88, 0.0)(89, 0.0)(90, 0.0)(91, 0.0)(92, 0.0)(93, 0.0)(94, 0.0)(95, 0.0)(96, 0.0)(97, 0.0)(98, 0.0)(99, 0.0)(100, 0.0)(101, 0.0)(102, 0.0)(103, 0.0)(104, 0.0)(105, 0.0)(106, 0.0)(107, 0.0)(108, 0.0)(109, 0.0)(110, 0.0)(111, 0.0)(112, 0.0)(113, 0.0)(114, 0.0)(115, 0.0)(116, 0.0)(117, 0.0)(118, 0.0)(119, 0.0)(120, 0.0)(121, 0.0)(122, 0.0)(123, 0.0)(124, 0.0)(125, 0.0)(126, 0.0)(127, 0.0)(128, 0.0)(129, 0.0)(130, 0.0)(131, 0.0)(132, 0.0)(133, 0.0)(134, 0.0)(135, 0.0)(136, 0.0)(137, 0.0)(138, 0.0)(139, 0.0)(140, 0.0)(141, 0.0)(142, 0.0)(143, 0.0)(144, 0.0)(145, 0.0)(146, 0.0)(147, 0.0)(148, 0.0)(149, 0.0)(150, 0.0)(151, 0.0)(152, 0.0)(153, 0.0)(154, 0.0)(155, 0.0)(156, 0.0)(157, 0.0)(158, 0.0)(159, 0.0)(160, 0.0)(161, 0.0)(162, 0.0)(163, 0.0)(164, 0.0)(165, 0.0)(166, 0.0)(167, 0.0)(168, 0.0)(169, 0.0)(170, 0.0)(171, 0.0)(172, 0.0)(173, 0.0)(174, 0.0)(175, 0.0)(176, 0.0)(177, 0.0)(178, 0.0)(179, 0.0)(180, 0.0)(181, 0.0)(182, 0.0)(183, 0.0)(184, 0.0)(185, 0.0)(186, 0.0)(187, 0.0)(188, 0.0)(189, 0.0)(190, 0.0)(191, 0.0)(192, 0.0)(193, 0.0)(194, 0.0)(195, 0.0)(196, 0.0)(197, 0.0)(198, 0.0)(199, 0.0)(200, 0.0)(201, 0.0)(202, 0.0)(203, 0.0)(204, 0.0)(205, 0.0)(206, 0.0)(207, 0.0)(208, 0.0)(209, 0.0)(210, 0.0)(211, 0.0)(212, 0.0)(213, 0.0)(214, 0.0)(215, 0.0)(216, 0.0)(217, 0.0)(218, 0.0)(219, 0.0)(220, 0.0)(221, 0.0)(222, 0.0)(223, 0.0)(224, 0.0)(225, 0.0)(226, 0.0)(227, 0.0)(228, 0.0)(229, 0.0)(230, 0.0)(231, 0.0)(232, 0.0)(233, 0.0)(234, 0.0)(235, 0.0)(236, 0.0)(237, 0.0)(238, 0.0)(239, 0.0)(240, 0.0)(241, 0.0)(242, 0.0)(243, 0.0)(244, 0.0)(245, 0.0)(246, 0.0)(247, 0.0)(248, 0.0)(249, 0.0)(250, 0.0)(251, 0.0)(252, 0.0)(253, 0.0)(254, 0.0)(255, 0.0)(256, 0.0)(257, 0.0)(258, 0.0)(259, 0.0)(260, 0.0)(261, 0.0)(262, 0.0)(263, 0.0)(264, 0.0)(265, 0.0)(266, 0.0)(267, 0.0)(268, 0.0)(269, 0.0)(270, 0.0)(271, 0.0)(272, 0.0)(273, 0.0)(274, 0.0)(275, 0.0)(276, 0.0)(277, 0.0)(278, 0.0)(279, 0.0)(280, 0.0)(281, 0.0)(282, 0.0)(283, 0.0)(284, 0.0)(285, 0.0)(286, 0.0)(287, 0.0)(288, 0.0)(289, 0.0)(290, 0.0)(291, 0.0)(292, 0.0)(293, 0.0)(294, 0.0)(295, 0.0)(296, 0.0)(297, 0.0)(298, 0.0)(299, 0.0)(300, 0.0)(301, 0.0)(302, 0.0)(303, 0.0)(304, 0.0)(305, 0.0)(306, 0.0)(307, 0.0)(308, 0.0)(309, 0.0)(310, 0.0)(311, 0.0)(312, 0.0)(313, 0.0)(314, 0.0)(315, 0.0)(316, 0.0)(317, 0.0)(318, 0.0)(319, 0.0)(320, 0.0)(321, 0.0)(322, 0.0)(323, 0.0)(324, 0.0)(325, 0.0)(326, 0.0)(327, 0.0)(328, 0.0)(329, 0.0)(330, 0.0)(331, 0.0)(332, 0.0)(333, 0.0)(334, 0.0)(335, 0.0)(336, 0.0)(337, 0.0)(338, 0.0)(339, 0.0)(340, 0.0)(341, 0.0)(342, 0.0)(343, 0.0)(344, 0.0)(345, 0.0)(346, 0.0)(347, 0.0)(348, 0.0)(349, 0.0)(350, 0.0)(351, 0.0)(352, 0.0)(353, 0.0)(354, 0.0)(355, 0.0)(356, 0.0)(357, 0.0)(358, 0.0)(359, 0.0)(360, 0.0)(361, 0.0)(362, 0.0)(363, 0.0)(364, 0.0)(365, 0.0)(366, 0.0)(367, 0.0)(368, 0.0)(369, 0.0)(370, 0.0)(371, 0.0)(372, 0.0)(373, 0.0)(374, 0.0)(375, 0.0)(376, 0.0)(377, 0.0)(378, 0.0)(379, 0.0)(380, 0.0)(381, 0.0)(382, 0.0)(383, 0.0)(384, 0.0)(385, 0.0)(386, 0.0)(387, 0.0)(388, 0.0)(389, 0.0)(390, 0.0)(391, 0.0)(392, 0.0)(393, 0.0)(394, 0.0)(395, 0.0)(396, 0.0)(397, 0.0)(398, 0.0)(399, 0.0)(400, 0.0)(401, 0.0)(402, 0.0)(403, 0.0)(404, 0.0)(405, 0.0)(406, 0.0)(407, 0.0)(408, 0.0)(409, 0.0)(410, 0.0)(411, 0.0)(412, 0.0)(413, 0.0)(414, 0.0)(415, 0.0)(416, 0.0)(417, 0.0)(418, 0.0)(419, 0.0)(420, 0.0)(421, 0.0)(422, 0.0)(423, 0.0)(424, 0.0)(425, 0.0)(426, 0.0)(427, 0.0)(428, 0.0)(429, 0.0)(430, 0.0)(431, 0.0)(432, 0.0)(433, 0.0)(434, 0.0)(435, 0.0)(436, 0.0)(437, 0.0)(438, 0.0)(439, 0.0)(440, 0.0)(441, 0.0)(442, 0.0)(443, 0.0)(444, 0.0)(445, 0.0)(446, 0.0)(447, 0.0)(448, 0.0)(449, 0.0)(450, 0.0)(451, 0.0)(452, 0.0)(453, 0.0)(454, 0.0)(455, 0.0)(456, 0.0)(457, 0.0)(458, 0.0)(459, 0.0)(460, 0.0)(461, 0.0)(462, 0.0)(463, 0.0)(464, 0.0)(465, 0.0)(466, 0.0)(467, 0.0)(468, 0.0)(469, 0.0)(470, 0.0)(471, 0.0)(472, 0.0)(473, 0.0)(474, 0.0)(475, 0.0)(476, 0.0)(477, 0.0)(478, 0.0)(479, 0.0)(480, 0.0)(481, 0.0)(482, 0.0)(483, 0.0)(484, 0.0)(485, 0.0)(486, 0.0)(487, 0.0)(488, 0.0)(489, 0.0)(490, 0.0)(491, 0.0)(492, 0.0)(493, 0.0)(494, 0.0)(495, 0.0)(496, 0.0)(497, 0.0)(498, 0.0)(499, 0.0)(500, 0.0)(501, 0.0)(502, 0.0)(503, 0.0)(504, 0.0)(505, 0.0)(506, 0.0)(507, 0.0)(508, 0.0)(509, 0.0)(510, 0.0)(511, 0.0)(512, 0.0)(513, 0.0)(514, 0.0)(515, 0.0)(516, 0.0)(517, 0.0)(518, 0.0)(519, 0.0)(520, 0.0)(521, 0.0)(522, 0.0)(523, 0.0)(524, 0.0)(525, 0.0)(526, 0.0)(527, 0.0)(528, 0.0)(529, 0.0)(530, 0.0)(531, 0.0)(532, 0.0)(533, 0.0)(534, 0.0)(535, 0.0)(536, 0.0)(537, 0.0)(538, 0.0)(539, 0.0)(540, 0.0)(541, 0.0)(542, 0.0)(543, 0.0)(544, 0.0)(545, 0.0)(546, 0.0)(547, 0.0)(548, 0.0)(549, 0.0)(550, 0.0)(551, 0.0)(552, 0.0)(553, 0.0)(554, 0.0)(555, 0.0)(556, 0.0)(557, 0.0)(558, 0.0)(559, 0.0)(560, 0.0)(561, 0.0)(562, 0.0)(563, 0.0)(564, 0.0)(565, 0.0)(566, 0.0)(567, 0.0)(568, 0.0)(569, 0.0)(570, 0.0)(571, 0.0)(572, 0.0)(573, 0.0)(574, 0.0)(575, 0.0)(576, 0.0)(577, 0.0)(578, 0.0)(579, 0.0)(580, 0.0)(581, 0.0)(582, 0.0)(583, 0.0)(584, 0.0)(585, 0.0)(586, 0.0)(587, 0.0)(588, 0.0)(589, 0.0)(590, 0.0)(591, 0.0)(592, 0.0)(593, 0.0)(594, 0.0)(595, 0.0)(596, 0.0)(597, 0.0)(598, 0.0)(599, 0.0)(600, 0.0)(601, 0.0)(602, 0.0)(603, 0.0)(604, 0.0)(605, 0.0)(606, 0.0)(607, 0.0)(608, 0.0)(609, 0.0)(610, 0.0)(611, 0.0)(612, 0.0)(613, 0.0)(614, 0.0)(615, 0.0)(616, 0.0)(617, 0.0)(618, 0.0)(619, 0.0)(620, 0.0)(621, 0.0)(622, 0.0)(623, 0.0)(624, 0.0)(625, 0.0)(626, 0.0)(627, 0.0)(628, 0.0)(629, 0.0)(630, 0.0)(631, 0.0)(632, 0.0)(633, 0.0)(634, 0.0)(635, 0.0)(636, 0.0)(637, 0.0)(638, 0.0)(639, 0.0)(640, 0.0)(641, 0.0)(642, 0.0)(643, 0.0)(644, 0.0)(645, 0.0)(646, 0.0)(647, 0.0)(648, 0.0)(649, 0.0)(650, 0.0)(651, 0.0)(652, 0.0)(653, 0.0)(654, 0.0)(655, 0.0)(656, 0.0)(657, 0.0)(658, 0.0)(659, 0.0)(660, 0.0)(661, 0.0)(662, 0.0)(663, 0.0)(664, 0.0)(665, 0.0)(666, 0.0)(667, 0.0)(668, 0.0)(669, 0.0)(670, 0.0)(671, 0.0)(672, 0.0)(673, 0.0)(674, 0.0)(675, 0.0)(676, 0.0)(677, 0.0)(678, 0.0)(679, 0.0)(680, 0.0)(681, 0.0)(682, 0.0)(683, 0.0)(684, 0.0)(685, 0.0)(686, 0.0)(687, 0.0)(688, 0.0)(689, 0.0)(690, 0.0)(691, 0.0)(692, 0.0)(693, 0.0)(694, 0.0)(695, 0.0)(696, 0.0)(697, 0.0)(698, 0.0)(699, 0.0)(700, 0.0)(701, 0.0)(702, 0.0)(703, 0.0)(704, 0.0)(705, 0.0)(706, 0.0)(707, 0.0)(708, 0.0)(709, 0.0)(710, 0.0)(711, 0.0)(712, 0.0)(713, 0.0)(714, 0.0)(715, 0.0)(716, 0.0)(717, 0.0)(718, 0.0)(719, 0.0)(720, 0.0)(721, 0.0)(722, 0.0)(723, 0.0)(724, 0.0)(725, 0.0)(726, 0.0)(727, 0.0)(728, 0.0)(729, 0.0)(730, 0.0)(731, 0.0)(732, 0.0)(733, 0.0)(734, 0.0)(735, 0.0)(736, 0.0)(737, 0.0)(738, 0.0)(739, 0.0)(740, 0.0)(741, 0.0)(742, 0.0)(743, 0.0)(744, 0.0)(745, 0.0)(746, 0.0)(747, 0.0)(748, 0.0)(749, 0.0)(750, 0.0)(751, 0.0)(752, 0.0)(753, 0.0)(754, 0.0)(755, 0.0)(756, 0.0)(757, 0.0)(758, 0.0)(759, 0.0)(760, 0.0)(761, 0.0)(762, 0.0)(763, 0.0)(764, 0.0)(765, 0.0)(766, 0.0)(767, 0.0)(768, 0.0)(769, 0.0)(770, 0.0)(771, 0.0)(772, 0.0)(773, 0.0)(774, 0.0)(775, 0.0)(776, 0.0)(777, 0.0)(778, 0.0)(779, 0.0)(780, 0.0)(781, 0.0)(782, 0.0)(783, 0.0)(784, 0.0)(785, 0.0)(786, 0.0)(787, 0.0)(788, 0.0)(789, 0.0)(790, 0.0)(791, 0.0)(792, 0.0)(793, 0.0)(794, 0.0)(795, 0.0)(796, 0.0)(797, 0.0)(798, 0.0)(799, 0.0)(800, 0.0)(801, 0.0)(802, 0.0)(803, 0.0)(804, 0.0)(805, 0.0)(806, 0.0)(807, 0.0)(808, 0.0)(809, 0.0)(810, 0.0)(811, 0.0)(812, 0.0)(813, 0.0)(814, 0.0)(815, 0.0)(816, 0.0)(817, 0.0)(818, 0.0)(819, 0.0)(820, 0.0)(821, 0.0)(822, 0.0)(823, 0.0)(824, 0.0)(825, 0.0)(826, 0.0)(827, 0.0)(828, 0.0)(829, 0.0)(830, 0.0)(831, 0.0)(832, 0.0)(833, 0.0)(834, 0.0)(835, 0.0)(836, 0.0)(837, 0.0)(838, 0.0)(839, 0.0)(840, 0.0)(841, 0.0)(842, 0.0)(843, 0.0)(844, 0.0)(845, 0.0)(846, 0.0)(847, 0.0)(848, 0.0)(849, 0.0)(850, 0.0)(851, 0.0)(852, 0.0)(853, 0.0)(854, 0.0)(855, 0.0)(856, 0.0)(857, 0.0)(858, 0.0)(859, 0.0)(860, 0.0)(861, 0.0)(862, 0.0)(863, 0.0)(864, 0.0)(865, 0.0)(866, 0.0)(867, 0.0)(868, 0.0)(869, 0.0)(870, 0.0)(871, 0.0)(872, 0.0)(873, 0.0)(874, 0.0)(875, 0.0)(876, 0.0)(877, 0.0)(878, 0.0)(879, 0.0)(880, 0.0)(881, 0.0)(882, 0.0)(883, 0.0)(884, 0.0)(885, 0.0)(886, 0.0)(887, 0.0)(888, 0.0)(889, 0.0)(890, 0.0)(891, 0.0)(892, 0.0)(893, 0.0)(894, 0.0)(895, 0.0)(896, 0.0)(897, 0.0)(898, 0.0)(899, 0.0)(900, 0.0)(901, 0.0)(902, 0.0)(903, 0.0)(904, 0.0)(905, 0.0)(906, 0.0)(907, 0.0)(908, 0.0)(909, 0.0)(910, 0.0)(911, 0.0)(912, 0.0)(913, 0.0)(914, 0.0)(915, 0.0)(916, 0.0)(917, 0.0)(918, 0.0)(919, 0.0)(920, 0.0)(921, 0.0)(922, 0.0)(923, 0.0)(924, 0.0)(925, 0.0)(926, 0.0)(927, 0.0)(928, 0.0)(929, 0.0)(930, 0.0)(931, 0.0)(932, 0.0)(933, 0.0)(934, 0.0)(935, 0.0)(936, 0.0)(937, 0.0)(938, 0.0)(939, 0.0)(940, 0.0)(941, 0.0)(942, 0.0)(943, 0.0)(944, 0.0)(945, 0.0)(946, 0.0)(947, 0.0)(948, 0.0)(949, 0.0)(950, 0.0)(951, 0.0)(952, 0.0)(953, 0.0)(954, 0.0)(955, 0.0)(956, 0.0)(957, 0.0)(958, 0.0)(959, 0.0)(960, 0.0)(961, 0.0)(962, 0.0)(963, 0.0)(964, 0.0)(965, 0.0)(966, 0.0)(967, 0.0)(968, 0.0)(969, 0.0)(970, 0.0)(971, 0.0)(972, 0.0)(973, 0.0)(974, 0.0)(975, 0.0)(976, 0.0)(977, 0.0)(978, 0.0)(979, 0.0)(980, 0.0)(981, 0.0)(982, 0.0)(983, 0.0)(984, 0.0)(985, 0.0)(986, 0.0)(987, 0.0)(988, 0.0)(989, 0.0)(990, 0.0)(991, 0.0)(992, 0.0)(993, 0.0)(994, 0.0)(995, 0.0)(996, 0.0)(997, 0.0)(998, 0.0)(999, 0.0)(1000, 0.0)(1001, 0.0)(1002, 0.0)(1003, 0.0)(1004, 0.0)(1005, 0.0)(1006, 0.0)(1007, 0.0)(1008, 0.0)(1009, 0.0)(1010, 0.0)(1011, 0.0)(1012, 0.0)(1013, 0.0)(1014, 0.0)(1015, 0.0)(1016, 0.0)(1017, 0.0)(1018, 0.0)(1019, 0.0)(1020, 0.0)(1021, 0.0)(1022, 0.0)(1023, 0.0)(1024, 0.0)(1025, 0.0)(1026, 0.0)(1027, 0.0)(1028, 0.0)(1029, 0.0)(1030, 0.0)(1031, 0.0)(1032, 0.0)(1033, 0.0)(1034, 0.0)(1035, 0.0)(1036, 0.0)(1037, 0.0)(1038, 0.0)(1039, 0.0)(1040, 0.0)(1041, 0.0)(1042, 0.0)(1043, 0.0)(1044, 0.0)(1045, 0.0)(1046, 0.0)(1047, 0.0)(1048, 0.0)(1049, 0.0)(1050, 0.0)(1051, 0.0)(1052, 0.0)(1053, 0.0)(1054, 0.0)(1055, 0.0)(1056, 0.0)(1057, 0.0)(1058, 0.0)(1059, 0.0)(1060, 0.0)(1061, 0.0)(1062, 0.0)(1063, 0.0)(1064, 0.0)(1065, 0.0)(1066, 0.0)(1067, 0.0)(1068, 0.0)(1069, 0.0)(1070, 0.0)(1071, 0.0)(1072, 0.0)(1073, 0.0)(1074, 0.0)(1075, 0.0)(1076, 0.0)(1077, 0.0)(1078, 0.0)(1079, 0.0)(1080, 0.0)(1081, 0.0)(1082, 0.0)(1083, 0.0)(1084, 0.0)(1085, 0.0)(1086, 0.0)(1087, 0.0)(1088, 0.0)(1089, 0.0)(1090, 0.0)(1091, 0.0)(1092, 0.0)(1093, 0.0)(1094, 0.0)(1095, 0.0)(1096, 0.0)(1097, 0.0)(1098, 0.0)(1099, 0.0)(1100, 0.0)(1101, 0.0)(1102, 0.0)(1103, 0.0)(1104, 0.0)(1105, 0.0)(1106, 0.0)(1107, 0.0)(1108, 0.0)(1109, 0.0)(1110, 0.0)(1111, 0.0)(1112, 0.0)(1113, 0.0)(1114, 0.0)(1115, 0.0)(1116, 0.0)(1117, 0.0)(1118, 0.0)(1119, 0.0)(1120, 0.0)(1121, 0.0)(1122, 0.0)(1123, 0.0)(1124, 0.0)(1125, 0.0)(1126, 0.0)(1127, 0.0)(1128, 0.0)(1129, 0.0)(1130, 0.0)(1131, 0.0)(1132, 0.0)(1133, 0.0)(1134, 0.0)(1135, 0.0)(1136, 0.0)(1137, 0.0)(1138, 0.0)(1139, 0.0)(1140, 0.0)(1141, 0.0)(1142, 0.0)(1143, 0.0)(1144, 0.0)(1145, 0.0)(1146, 0.0)(1147, 0.0)(1148, 0.0)(1149, 0.0)(1150, 0.0)(1151, 0.0)(1152, 0.0)(1153, 0.0)(1154, 0.0)(1155, 0.0)(1156, 0.0)(1157, 0.0)(1158, 0.0)(1159, 0.0)(1160, 0.0)(1161, 0.0)(1162, 0.0)(1163, 0.0)(1164, 0.0)(1165, 0.0)(1166, 0.0)(1167, 0.0)(1168, 0.0)(1169, 0.0)(1170, 0.0)(1171, 0.0)(1172, 0.0)(1173, 0.0)(1174, 0.0)(1175, 0.0)(1176, 0.0)(1177, 0.0)(1178, 0.0)(1179, 0.0)(1180, 0.0)(1181, 0.0)(1182, 0.0)(1183, 0.0)(1184, 0.0)(1185, 0.0)(1186, 0.0)(1187, 0.0)(1188, 0.0)(1189, 0.0)(1190, 0.0)(1191, 0.0)(1192, 0.0)(1193, 0.0)(1194, 0.0)(1195, 0.0)(1196, 0.0)(1197, 0.0)(1198, 0.0)(1199, 0.0)(1200, 0.0)(1201, 0.0)(1202, 0.0)(1203, 0.0)(1204, 0.0)(1205, 0.0)(1206, 0.0)(1207, 0.0)(1208, 0.0)(1209, 0.0)(1210, 0.0)(1211, 0.0)(1212, 0.0)(1213, 0.0)(1214, 0.0)(1215, 0.0)(1216, 0.0)(1217, 0.0)(1218, 0.0)(1219, 0.0)(1220, 0.0)(1221, 0.0)(1222, 0.0)(1223, 0.0)(1224, 0.0)(1225, 0.0)(1226, 0.0)(1227, 0.0)(1228, 0.0)(1229, 0.0)(1230, 0.0)(1231, 0.0)(1232, 0.0)(1233, 0.0)(1234, 0.0)(1235, 0.0)(1236, 0.0)(1237, 0.0)(1238, 0.0)(1239, 0.0)(1240, 0.0)(1241, 0.0)(1242, 0.0)(1243, 0.0)(1244, 0.0)(1245, 0.0)(1246, 0.0)(1247, 0.0)(1248, 0.0)(1249, 0.0)(1250, 0.0)(1251, 0.0)(1252, 0.0)(1253, 0.0)(1254, 0.0)(1255, 0.0)(1256, 0.0)(1257, 0.0)(1258, 0.0)(1259, 0.0)(1260, 0.0)(1261, 0.0)(1262, 0.0)(1263, 0.0)(1264, 0.0)(1265, 0.0)(1266, 0.0)(1267, 0.0)(1268, 0.0)(1269, 0.0)(1270, 0.0)(1271, 0.0)(1272, 0.0)(1273, 0.0)(1274, 0.0)(1275, 0.0)(1276, 0.0)(1277, 0.0)(1278, 0.0)(1279, 0.0)(1280, 0.0)(1281, 0.0)(1282, 0.0)(1283, 0.0)(1284, 0.0)(1285, 0.0)(1286, 0.0)(1287, 0.0)(1288, 0.0)(1289, 0.0)(1290, 0.0)(1291, 0.0)(1292, 0.0)(1293, 0.0)(1294, 0.0)(1295, 0.0)(1296, 0.0)(1297, 0.0)(1298, 0.0)(1299, 0.0)(1300, 0.0)(1301, 0.0)(1302, 0.0)(1303, 0.0)(1304, 0.0)(1305, 0.0)(1306, 0.0)(1307, 0.0)(1308, 0.0)(1309, 0.0)(1310, 0.0)(1311, 0.0)(1312, 0.0)(1313, 0.0)(1314, 0.0)(1315, 0.0)(1316, 0.0)(1317, 0.0)(1318, 0.0)(1319, 0.0)(1320, 0.0)(1321, 0.0)(1322, 0.0)(1323, 0.0)(1324, 0.0)(1325, 0.0)(1326, 0.0)(1327, 0.0)(1328, 0.0)(1329, 0.0)(1330, 0.0)(1331, 0.0)(1332, 0.0)(1333, 0.0)(1334, 0.0)(1335, 0.0)(1336, 0.0)(1337, 0.0)(1338, 0.0)(1339, 0.0)(1340, 0.0)(1341, 0.0)(1342, 0.0)(1343, 0.0)(1344, 0.0)(1345, 0.0)(1346, 0.0)(1347, 0.0)(1348, 0.0)(1349, 0.0)(1350, 0.0)(1351, 0.0)(1352, 0.0)(1353, 0.0)(1354, 0.0)(1355, 0.0)(1356, 0.0)(1357, 0.0)(1358, 0.0)(1359, 0.0)(1360, 0.0)(1361, 0.0)(1362, 0.0)(1363, 0.0)(1364, 0.0)(1365, 0.0)(1366, 0.0)(1367, 0.0)(1368, 0.0)(1369, 0.0)(1370, 0.0)(1371, 0.0)(1372, 0.0)(1373, 0.0)(1374, 0.0)(1375, 0.0)(1376, 0.0)(1377, 0.0)(1378, 0.0)(1379, 0.0)(1380, 0.0)(1381, 0.0)(1382, 0.0)(1383, 0.0)(1384, 0.0)(1385, 0.0)(1386, 0.0)(1387, 0.0)(1388, 0.0)(1389, 0.0)(1390, 0.0)(1391, 0.0)(1392, 0.0)(1393, 0.0)(1394, 0.0)(1395, 0.0)(1396, 0.0)(1397, 0.0)(1398, 0.0)(1399, 0.0)(1400, 0.0)(1401, 0.0)(1402, 0.0)(1403, 0.0)(1404, 0.0)(1405, 0.0)(1406, 0.0)(1407, 0.0)(1408, 0.0)(1409, 0.0)(1410, 0.0)(1411, 0.0)(1412, 0.0)(1413, 0.0)(1414, 0.0)(1415, 0.0)(1416, 0.0)(1417, 0.0)(1418, 0.0)(1419, 0.0)(1420, 0.0)(1421, 0.0)(1422, 0.0)(1423, 0.0)(1424, 0.0)(1425, 0.0)(1426, 0.0)(1427, 0.0)(1428, 0.0)(1429, 0.0)(1430, 0.0)(1431, 0.0)(1432, 0.0)(1433, 0.0)(1434, 0.0)(1435, 0.0)(1436, 0.0)(1437, 0.0)(1438, 0.0)(1439, 0.0)(1440, 0.0)(1441, 0.0)(1442, 0.0)(1443, 0.0)(1444, 0.0)(1445, 0.0)(1446, 0.0)(1447, 0.0)(1448, 0.0)(1449, 0.0)(1450, 0.0)(1451, 0.0)(1452, 0.0)(1453, 0.0)(1454, 0.0)(1455, 0.0)(1456, 0.0)(1457, 0.0)(1458, 0.0)(1459, 0.0)(1460, 0.0)(1461, 0.0)(1462, 0.0)(1463, 0.0)(1464, 0.0)(1465, 0.0)(1466, 0.0)(1467, 0.0)(1468, 0.0)(1469, 0.0)(1470, 0.0)(1471, 0.0)(1472, 0.0)(1473, 0.0)(1474, 0.0)(1475, 0.0)(1476, 0.0)(1477, 0.0)(1478, 0.0)(1479, 0.0)(1480, 0.0)(1481, 0.0)(1482, 0.0)(1483, 0.0)(1484, 0.0)(1485, 0.0)(1486, 0.0)(1487, 0.0)(1488, 0.0)(1489, 0.0)(1490, 0.0)(1491, 0.0)(1492, 0.0)(1493, 0.0)(1494, 0.0)(1495, 0.0)(1496, 0.0)(1497, 0.0)(1498, 0.0)(1499, 0.0)(1500, 0.0)(1501, 0.0)(1502, 0.0)(1503, 0.0)(1504, 0.0)(1505, 0.0)(1506, 0.0)(1507, 0.0)(1508, 0.0)(1509, 0.0)(1510, 0.0)(1511, 0.0)(1512, 0.0)(1513, 0.0)(1514, 0.0)(1515, 0.0)(1516, 0.0)(1517, 0.0)(1518, 0.0)(1519, 0.0)};
            \addplot+[name path=G,black,line width=.1pt] coordinates {(0, 0.0)(1, 0.0)(2, 0.0)(3, 0.0)(4, 0.0)(5, 0.0)(6, 0.0)(7, 0.0)(8, 0.0)(9, 0.0)(10, 0.0)(11, 0.0)(12, 0.0)(13, 0.0)(14, 0.0)(15, 0.0)(16, 0.0)(17, 0.0)(18, 0.0)(19, 0.0)(20, 0.0)(21, 0.0)(22, 0.0)(23, 0.0)(24, 0.0)(25, 0.0)(26, 0.0)(27, 0.0)(28, 0.0)(29, 0.0)(30, 0.0)(31, 0.0)(32, 0.0)(33, 0.0)(34, 0.0)(35, 0.0)(36, 0.0)(37, 0.0)(38, 0.0)(39, 0.0)(40, 0.0)(41, 0.0)(42, 0.0)(43, 0.0)(44, 0.0)(45, 0.0)(46, 0.0)(47, 0.0)(48, 0.0)(49, 0.0)(50, 0.0)(51, 0.0)(52, 0.0)(53, 0.0)(54, 0.0)(55, 0.0)(56, 0.0)(57, 0.0)(58, 0.0)(59, 0.0)(60, 0.0)(61, 0.0)(62, 0.0)(63, 0.0)(64, 0.0)(65, 0.0)(66, 0.0)(67, 0.0)(68, 0.0)(69, 0.0)(70, 0.0)(71, 0.0)(72, 0.0)(73, 0.0)(74, 0.0)(75, 0.0)(76, 0.0)(77, 0.0)(78, 0.0)(79, 0.0)(80, 0.0)(81, 0.0)(82, 0.0)(83, 0.0)(84, 0.0)(85, 0.0)(86, 0.0)(87, 0.0)(88, 0.0)(89, 0.0)(90, 0.0)(91, 0.0)(92, 0.0)(93, 0.0)(94, 0.0)(95, 0.0)(96, 0.0)(97, 0.0)(98, 0.0)(99, 0.0)(100, 0.0)(101, 0.0)(102, 0.0)(103, 0.0)(104, 0.0)(105, 0.0)(106, 0.0)(107, 0.0)(108, 0.0)(109, 0.0)(110, 0.0)(111, 0.0)(112, 0.0)(113, 0.0)(114, 0.0)(115, 0.0)(116, 0.0)(117, 0.0)(118, 0.0)(119, 0.0)(120, 0.0)(121, 0.0)(122, 0.0)(123, 0.0)(124, 0.0)(125, 0.0)(126, 0.0)(127, 0.0)(128, 0.0)(129, 0.0)(130, 0.0)(131, 0.0)(132, 0.0)(133, 0.0)(134, 0.0)(135, 0.0)(136, 0.0)(137, 0.0)(138, 0.0)(139, 0.0)(140, 0.0)(141, 0.0)(142, 0.0)(143, 0.0)(144, 0.0)(145, 0.0)(146, 0.0)(147, 0.0)(148, 0.0)(149, 0.0)(150, 0.0)(151, 0.0)(152, 0.0)(153, 0.0)(154, 0.0)(155, 0.0)(156, 0.0)(157, 0.0)(158, 0.0)(159, 0.0)(160, 0.0)(161, 0.0)(162, 0.0)(163, 0.0)(164, 0.0)(165, 0.0)(166, 0.0)(167, 0.0)(168, 0.0)(169, 0.0)(170, 0.0)(171, 0.0)(172, 0.0)(173, 0.0)(174, 0.0)(175, 0.0)(176, 0.0)(177, 0.0)(178, 0.0)(179, 0.0)(180, 0.0)(181, 0.0)(182, 0.0)(183, 0.0)(184, 0.0)(185, 0.0)(186, 0.0)(187, 0.0)(188, 0.0)(189, 0.0)(190, 0.0)(191, 0.0)(192, 0.0)(193, 0.0)(194, 0.0)(195, 0.0)(196, 0.0)(197, 0.0)(198, 0.0)(199, 0.0)(200, 0.0)(201, 0.0)(202, 0.0)(203, 0.0)(204, 0.0)(205, 0.0)(206, 0.0)(207, 0.0)(208, 0.0)(209, 0.0)(210, 0.0)(211, 0.0)(212, 0.0)(213, 0.0)(214, 0.0)(215, 0.0)(216, 0.0)(217, 0.0)(218, 0.0)(219, 0.0)(220, 0.0)(221, 0.0)(222, 0.0)(223, 0.0)(224, 0.0)(225, 0.0)(226, 0.0)(227, 0.0)(228, 0.0)(229, 0.0)(230, 0.0)(231, 0.0)(232, 0.0)(233, 0.0)(234, 0.0)(235, 0.0)(236, 0.0)(237, 0.0)(238, 0.0)(239, 0.0)(240, 0.0)(241, 0.0)(242, 0.0)(243, 0.0)(244, 0.0)(245, 0.0)(246, 0.0)(247, 0.0)(248, 0.0)(249, 0.0)(250, 0.0)(251, 0.0)(252, 0.0)(253, 0.0)(254, 0.0)(255, 0.0)(256, 0.0)(257, 0.0)(258, 0.0)(259, 0.0)(260, 0.0)(261, 0.0)(262, 0.0)(263, 0.0)(264, 0.0)(265, 0.0)(266, 0.0)(267, 0.0)(268, 0.0)(269, 0.0)(270, 0.0)(271, 0.0)(272, 0.0)(273, 0.0)(274, 0.0)(275, 0.0)(276, 0.0)(277, 0.0)(278, 0.0)(279, 0.0)(280, 0.0)(281, 0.0)(282, 0.0)(283, 0.0)(284, 0.0)(285, 0.0)(286, 0.0)(287, 0.0)(288, 0.0)(289, 0.0)(290, 0.0)(291, 0.0)(292, 0.0)(293, 0.0)(294, 0.0)(295, 0.0)(296, 0.0)(297, 0.0)(298, 0.0)(299, 0.0)(300, 0.0)(301, 0.0)(302, 0.0)(303, 0.0)(304, 0.0)(305, 0.0)(306, 0.0)(307, 0.0)(308, 0.0)(309, 0.0)(310, 0.0)(311, 0.0)(312, 0.0)(313, 0.0)(314, 0.0)(315, 0.0)(316, 0.0)(317, 0.0)(318, 0.0)(319, 0.0)(320, 0.0)(321, 0.0)(322, 0.0)(323, 0.0)(324, 0.0)(325, 0.0)(326, 0.0)(327, 0.0)(328, 0.0)(329, 0.0)(330, 0.0)(331, 0.0)(332, 0.0)(333, 0.0)(334, 0.0)(335, 0.0)(336, 0.0)(337, 0.0)(338, 0.0)(339, 0.0)(340, 0.0)(341, 0.0)(342, 0.0)(343, 0.0)(344, 0.0)(345, 0.0)(346, 0.0)(347, 0.0)(348, 0.0)(349, 0.0)(350, 0.0)(351, 0.0)(352, 0.0)(353, 0.0)(354, 0.0)(355, 0.0)(356, 0.0)(357, 0.0)(358, 0.0)(359, 0.0)(360, 0.0)(361, 0.0)(362, 0.0)(363, 0.0)(364, 0.0)(365, 0.0)(366, 0.0)(367, 0.0)(368, 0.0)(369, 0.0)(370, 0.0)(371, 0.0)(372, 0.0)(373, 0.0)(374, 0.0)(375, 0.0)(376, 0.0)(377, 0.0)(378, 0.0)(379, 0.0)(380, 0.0)(381, 0.0)(382, 0.0)(383, 0.0)(384, 0.0)(385, 0.0)(386, 0.0)(387, 0.0)(388, 0.0)(389, 0.0)(390, 0.0)(391, 0.0)(392, 0.0)(393, 0.0)(394, 0.0)(395, 0.0)(396, 0.0)(397, 0.0)(398, 0.0)(399, 0.0)(400, 0.0)(401, 0.0)(402, 0.0)(403, 0.0)(404, 0.0)(405, 0.0)(406, 0.0)(407, 0.0)(408, 0.0)(409, 0.0)(410, 0.0)(411, 0.0)(412, 0.0)(413, 0.0)(414, 0.0)(415, 0.0)(416, 0.0)(417, 0.0)(418, 0.0)(419, 0.0)(420, 0.0)(421, 0.0)(422, 0.0)(423, 0.0)(424, 0.0)(425, 0.0)(426, 0.0)(427, 0.0)(428, 0.0)(429, 0.0)(430, 0.0)(431, 0.0)(432, 0.0)(433, 0.0)(434, 0.0)(435, 0.0)(436, 0.0)(437, 0.0)(438, 0.0)(439, 0.0)(440, 0.0)(441, 0.0)(442, 0.0)(443, 0.0)(444, 0.0)(445, 0.0)(446, 0.0)(447, 0.0)(448, 0.0)(449, 0.0)(450, 0.0)(451, 0.0)(452, 0.0)(453, 0.0)(454, 0.0)(455, 0.0)(456, 0.0)(457, 0.0)(458, 0.0)(459, 0.0)(460, 0.0)(461, 0.0)(462, 0.0)(463, 0.0)(464, 0.0)(465, 0.0)(466, 0.0)(467, 0.0)(468, 0.0)(469, 0.0)(470, 0.0)(471, 0.0)(472, 0.0)(473, 0.0)(474, 0.0)(475, 0.0)(476, 0.0)(477, 0.0)(478, 0.0)(479, 0.0)(480, 0.0)(481, 0.0)(482, 0.0)(483, 0.0)(484, 0.0)(485, 0.0)(486, 0.0)(487, 0.0)(488, 0.0)(489, 0.0)(490, 0.0)(491, 0.0)(492, 0.0)(493, 0.0)(494, 0.0)(495, 0.0)(496, 0.0)(497, 0.0)(498, 0.0)(499, 0.0)(500, 0.0)(501, 0.0)(502, 0.0)(503, 0.0)(504, 0.0)(505, 0.0)(506, 0.0)(507, 0.0)(508, 0.0)(509, 0.0)(510, 0.0)(511, 0.0)(512, 0.0)(513, 0.0)(514, 0.0)(515, 0.0)(516, 0.0)(517, 0.0)(518, 0.0)(519, 0.0)(520, 0.0)(521, 0.0)(522, 0.0)(523, 0.0)(524, 0.0)(525, 0.0)(526, 0.0)(527, 0.0)(528, 0.0)(529, 0.0)(530, 0.0)(531, 0.0)(532, 0.0)(533, 0.0)(534, 0.0)(535, 0.0)(536, 0.0)(537, 0.0)(538, 0.0)(539, 0.0)(540, 0.0)(541, 0.0)(542, 0.0)(543, 0.0)(544, 0.0)(545, 0.0)(546, 0.0)(547, 0.0)(548, 0.0)(549, 0.0)(550, 0.0)(551, 0.0)(552, 0.0)(553, 0.0)(554, 0.0)(555, 0.0)(556, 0.0)(557, 0.0)(558, 0.0)(559, 0.0)(560, 0.0)(561, 0.0)(562, 0.0)(563, 0.0)(564, 0.0)(565, 0.0)(566, 0.0)(567, 0.0)(568, 0.0)(569, 0.0)(570, 0.0)(571, 0.0)(572, 0.0)(573, 0.0)(574, 0.0)(575, 0.0)(576, 0.0)(577, 0.0)(578, 0.0)(579, 0.0)(580, 0.0)(581, 0.0)(582, 0.0)(583, 0.0)(584, 0.0)(585, 0.0)(586, 0.0)(587, 0.0)(588, 0.0)(589, 0.0)(590, 0.0)(591, 0.0)(592, 0.0)(593, 0.0)(594, 0.0)(595, 0.0)(596, 0.0)(597, 0.0)(598, 0.0)(599, 0.0)(600, 0.0)(601, 0.0)(602, 0.0)(603, 0.0)(604, 0.0)(605, 0.0)(606, 0.0)(607, 0.0)(608, 0.0)(609, 0.0)(610, 0.0)(611, 0.0)(612, 0.0)(613, 0.0)(614, 0.0)(615, 0.0)(616, 0.0)(617, 0.0)(618, 0.0)(619, 0.0)(620, 0.0)(621, 0.0)(622, 0.0)(623, 0.0)(624, 0.0)(625, 0.0)(626, 0.0)(627, 0.0)(628, 0.0)(629, 0.0)(630, 0.0)(631, 0.0)(632, 0.0)(633, 0.0)(634, 0.0)(635, 0.0)(636, 0.0)(637, 0.0)(638, 0.0)(639, 0.0)(640, 0.0)(641, 0.0)(642, 0.0)(643, 0.0)(644, 0.0)(645, 0.0)(646, 0.0)(647, 0.0)(648, 0.0)(649, 0.0)(650, 0.0)(651, 0.0)(652, 0.0)(653, 0.0)(654, 0.0)(655, 0.0)(656, 0.0)(657, 0.0)(658, 0.0)(659, 0.0)(660, 0.0)(661, 0.0)(662, 0.0)(663, 0.0)(664, 0.0)(665, 0.0)(666, 0.0)(667, 0.0)(668, 0.0)(669, 0.0)(670, 0.0)(671, 0.0)(672, 0.0)(673, 0.0)(674, 0.0)(675, 0.0)(676, 0.0)(677, 0.0)(678, 0.0)(679, 0.0)(680, 0.0)(681, 0.0)(682, 0.0)(683, 0.0)(684, 0.0)(685, 0.0)(686, 0.0)(687, 0.0)(688, 0.0)(689, 0.0)(690, 0.0)(691, 0.0)(692, 0.0)(693, 0.0)(694, 0.0)(695, 0.0)(696, 0.0)(697, 0.0)(698, 0.0)(699, 0.0)(700, 0.0)(701, 0.0)(702, 0.0)(703, 0.0)(704, 0.0)(705, 0.0)(706, 0.0)(707, 0.0)(708, 0.0)(709, 0.0)(710, 0.0)(711, 0.0)(712, 0.0)(713, 0.0)(714, 0.0)(715, 0.0)(716, 0.0)(717, 0.0)(718, 0.0)(719, 0.0)(720, 0.0)(721, 0.0)(722, 0.0)(723, 0.0)(724, 0.0)(725, 0.0)(726, 0.0)(727, 0.0)(728, 0.0)(729, 0.0)(730, 0.0)(731, 0.0)(732, 0.0)(733, 0.0)(734, 0.0)(735, 0.0)(736, 0.0)(737, 0.0)(738, 0.0)(739, 0.0)(740, 0.0)(741, 0.0)(742, 0.0)(743, 0.0)(744, 0.0)(745, 0.0)(746, 0.0)(747, 0.0)(748, 0.0)(749, 0.0)(750, 0.0)(751, 0.0)(752, 0.0)(753, 0.0)(754, 0.0)(755, 0.0)(756, 0.0)(757, 0.0)(758, 0.0)(759, 0.0)(760, 0.0)(761, 0.0)(762, 0.0)(763, 0.0)(764, 0.0)(765, 0.0)(766, 0.0)(767, 0.0)(768, 0.0)(769, 0.0)(770, 0.0)(771, 0.0)(772, 0.0)(773, 0.0)(774, 0.0)(775, 0.0)(776, 0.0)(777, 0.0)(778, 0.0)(779, 0.0)(780, 0.0)(781, 0.0)(782, 0.0)(783, 0.0)(784, 0.0)(785, 0.0)(786, 0.0)(787, 0.0)(788, 0.0)(789, 0.0)(790, 0.0)(791, 0.0)(792, 0.0)(793, 0.0)(794, 0.0)(795, 0.0)(796, 0.0)(797, 0.0)(798, 0.0)(799, 0.0)(800, 0.0)(801, 0.0)(802, 0.0)(803, 0.0)(804, 0.0)(805, 0.0)(806, 0.0)(807, 0.0)(808, 0.0)(809, 0.0)(810, 0.0)(811, 0.0)(812, 0.0)(813, 0.0)(814, 0.0)(815, 0.0)(816, 0.0)(817, 0.0)(818, 0.0)(819, 0.0)(820, 0.0)(821, 0.0)(822, 0.0)(823, 0.0)(824, 0.0)(825, 0.0)(826, 0.0)(827, 0.0)(828, 0.0)(829, 0.0)(830, 0.0)(831, 0.0)(832, 0.0)(833, 0.0)(834, 0.0)(835, 0.0)(836, 0.0)(837, 0.0)(838, 0.0)(839, 0.0)(840, 0.0)(841, 0.0)(842, 0.0)(843, 0.0)(844, 0.0)(845, 0.0)(846, 0.0)(847, 0.0)(848, 0.0)(849, 0.0)(850, 0.0)(851, 0.0)(852, 0.0)(853, 0.0)(854, 0.0)(855, 0.0)(856, 0.0)(857, 0.0)(858, 0.0)(859, 0.0)(860, 0.0)(861, 0.0)(862, 0.0)(863, 0.0)(864, 0.0)(865, 0.0)(866, 0.0)(867, 0.0)(868, 0.0)(869, 0.0)(870, 0.0)(871, 0.0)(872, 0.0)(873, 0.0)(874, 0.0)(875, 0.0)(876, 0.0)(877, 0.0)(878, 0.0)(879, 0.0)(880, 0.0)(881, 0.0)(882, 0.0)(883, 0.0)(884, 0.0)(885, 0.0)(886, 0.0)(887, 0.0)(888, 0.0)(889, 0.0)(890, 0.0)(891, 0.0)(892, 0.0)(893, 0.0)(894, 0.0)(895, 0.0)(896, 0.0)(897, 0.0)(898, 0.0)(899, 0.0)(900, 0.0)(901, 0.0)(902, 0.0)(903, 0.0)(904, 0.0)(905, 0.0)(906, 0.0)(907, 0.0)(908, 0.0)(909, 0.0)(910, 0.0)(911, 0.0)(912, 0.0)(913, 0.0)(914, 0.0)(915, 0.0)(916, 0.0)(917, 0.0)(918, 0.0)(919, 0.0)(920, 0.0)(921, 0.0)(922, 0.0)(923, 0.0)(924, 0.0)(925, 0.0)(926, 0.0)(927, 0.0)(928, 0.0)(929, 0.0)(930, 0.0)(931, 0.0)(932, 0.0)(933, 0.0)(934, 0.0)(935, 0.0)(936, 0.0)(937, 0.0)(938, 0.0)(939, 0.0)(940, 0.0)(941, 0.0)(942, 0.0)(943, 0.0)(944, 0.0)(945, 0.0)(946, 0.0)(947, 0.0)(948, 0.0)(949, 0.0)(950, 0.0)(951, 0.0)(952, 0.0)(953, 0.0)(954, 0.0)(955, 0.0)(956, 0.0)(957, 0.0)(958, 0.0)(959, 0.0)(960, 0.0)(961, 0.0)(962, 0.0)(963, 0.0)(964, 0.0)(965, 0.0)(966, 0.0)(967, 0.0)(968, 0.0)(969, 0.0)(970, 0.0)(971, 0.0)(972, 0.0)(973, 0.0)(974, 0.0)(975, 0.0)(976, 0.0)(977, 0.0)(978, 0.0)(979, 0.0)(980, 0.0)(981, 0.0)(982, 0.0)(983, 0.0)(984, 0.0)(985, 0.0)(986, 0.0)(987, 0.0)(988, 0.0)(989, 0.0)(990, 0.0)(991, 0.0)(992, 0.0)(993, 0.0)(994, 0.0)(995, 0.0)(996, 0.0)(997, 0.0)(998, 0.0)(999, 0.0)(1000, 0.0)(1001, 0.0)(1002, 0.0)(1003, 0.0)(1004, 0.0)(1005, 0.0)(1006, 0.0)(1007, 0.0)(1008, 0.0)(1009, 0.0)(1010, 0.0)(1011, 0.0)(1012, 0.0)(1013, 0.0)(1014, 0.0)(1015, 0.0)(1016, 0.0)(1017, 0.0)(1018, 0.0)(1019, 0.0)(1020, 0.0)(1021, 0.0)(1022, 0.0)(1023, 0.0)(1024, 0.0)(1025, 0.0)(1026, 0.0)(1027, 0.0)(1028, 0.0)(1029, 0.0)(1030, 0.0)(1031, 0.0)(1032, 0.0)(1033, 0.0)(1034, 0.0)(1035, 0.0)(1036, 0.0)(1037, 0.0)(1038, 0.0)(1039, 0.0)(1040, 0.0)(1041, 0.0)(1042, 0.0)(1043, 0.0)(1044, 0.0)(1045, 0.0)(1046, 0.0)(1047, 0.0)(1048, 0.0)(1049, 0.0)(1050, 0.0)(1051, 0.0)(1052, 0.0)(1053, 0.0)(1054, 0.0)(1055, 0.0)(1056, 0.0)(1057, 0.0)(1058, 0.0)(1059, 0.0)(1060, 0.0)(1061, 0.0)(1062, 0.0)(1063, 0.0)(1064, 0.0)(1065, 0.0)(1066, 0.0)(1067, 0.0)(1068, 0.0)(1069, 0.0)(1070, 0.0)(1071, 0.0)(1072, 0.0)(1073, 0.0)(1074, 0.0)(1075, 0.0)(1076, 0.0)(1077, 0.0)(1078, 0.0)(1079, 0.0)(1080, 0.0)(1081, 0.0)(1082, 0.0)(1083, 0.0)(1084, 0.0)(1085, 0.0)(1086, 0.0)(1087, 0.0)(1088, 0.0)(1089, 0.0)(1090, 0.0)(1091, 0.0)(1092, 0.0)(1093, 0.0)(1094, 0.0)(1095, 0.0)(1096, 0.0)(1097, 0.0)(1098, 0.0)(1099, 0.0)(1100, 0.0)(1101, 0.0)(1102, 0.0)(1103, 0.0)(1104, 0.0)(1105, 0.0)(1106, 0.0)(1107, 0.0)(1108, 0.0)(1109, 0.0)(1110, 0.0)(1111, 0.0)(1112, 0.0)(1113, 0.0)(1114, 0.0)(1115, 0.0)(1116, 0.0)(1117, 0.0)(1118, 0.0)(1119, 0.0)(1120, 0.0)(1121, 0.0)(1122, 0.0)(1123, 0.0)(1124, 0.0)(1125, 0.0)(1126, 0.0)(1127, 0.0)(1128, 0.0)(1129, 0.0)(1130, 0.0)(1131, 0.0)(1132, 0.0)(1133, 0.0)(1134, 0.0)(1135, 0.0)(1136, 0.0)(1137, 0.0)(1138, 0.0)(1139, 0.0)(1140, 0.0)(1141, 0.0)(1142, 0.0)(1143, 0.0)(1144, 0.0)(1145, 0.0)(1146, 0.0)(1147, 0.0)(1148, 0.0)(1149, 0.0)(1150, 0.0)(1151, 0.0)(1152, 0.0)(1153, 0.0)(1154, 0.0)(1155, 0.0)(1156, 0.0)(1157, 0.0)(1158, 0.0)(1159, 0.0)(1160, 0.0)(1161, 0.0)(1162, 0.0)(1163, 0.0)(1164, 0.0)(1165, 0.0)(1166, 0.0)(1167, 0.0)(1168, 0.0)(1169, 0.0)(1170, 0.0)(1171, 0.0)(1172, 0.0)(1173, 0.0)(1174, 0.0)(1175, 0.0)(1176, 0.0)(1177, 0.0)(1178, 0.0)(1179, 0.0)(1180, 0.0)(1181, 0.0)(1182, 0.0)(1183, 0.0)(1184, 0.0)(1185, 0.0)(1186, 0.0)(1187, 0.0)(1188, 0.0)(1189, 0.0)(1190, 0.0)(1191, 0.0)(1192, 0.0)(1193, 0.0)(1194, 0.0)(1195, 0.0)(1196, 0.0)(1197, 0.0)(1198, 0.0)(1199, 0.0)(1200, 0.0)(1201, 0.0)(1202, 0.0)(1203, 0.0)(1204, 0.0)(1205, 0.0)(1206, 0.0)(1207, 0.0)(1208, 0.0)(1209, 0.0)(1210, 0.0)(1211, 0.0)(1212, 0.0)(1213, 0.0)(1214, 0.0)(1215, 0.0)(1216, 0.0)(1217, 0.0)(1218, 0.0)(1219, 0.0)(1220, 0.0)(1221, 0.0)(1222, 0.0)(1223, 0.0)(1224, 0.0)(1225, 0.0)(1226, 0.0)(1227, 0.0)(1228, 0.0)(1229, 0.0)(1230, 0.0)(1231, 0.0)(1232, 0.0)(1233, 0.0)(1234, 0.0)(1235, 0.0)(1236, 0.0)(1237, 0.0)(1238, 0.0)(1239, 0.0)(1240, 0.0)(1241, 0.0)(1242, 0.0)(1243, 0.0)(1244, 0.0)(1245, 0.0)(1246, 0.0)(1247, 0.0)(1248, 0.0)(1249, 0.0)(1250, 0.0)(1251, 0.0)(1252, 0.0)(1253, 0.0)(1254, 0.0)(1255, 0.0)(1256, 0.0)(1257, 0.0)(1258, 0.0)(1259, 0.0)(1260, 0.0)(1261, 0.0)(1262, 0.0)(1263, 0.0)(1264, 0.0)(1265, 0.0)(1266, 0.0)(1267, 0.0)(1268, 0.0)(1269, 0.0)(1270, 0.0)(1271, 0.0)(1272, 0.0)(1273, 0.0)(1274, 0.0)(1275, 0.0)(1276, 0.0)(1277, 0.0)(1278, 0.0)(1279, 0.0)(1280, 0.0)(1281, 0.0)(1282, 0.0)(1283, 0.0)(1284, 0.0)(1285, 0.0)(1286, 0.0)(1287, 0.0)(1288, 0.0)(1289, 0.0)(1290, 0.0)(1291, 0.0)(1292, 0.0)(1293, 0.0)(1294, 0.0)(1295, 0.0)(1296, 0.0)(1297, 0.0)(1298, 0.0)(1299, 0.0)(1300, 0.0)(1301, 0.0)(1302, 0.0)(1303, 0.0)(1304, 0.0)(1305, 0.0)(1306, 0.0)(1307, 0.0)(1308, 0.0)(1309, 0.0)(1310, 0.0)(1311, 0.0)(1312, 0.0)(1313, 0.0)(1314, 0.0)(1315, 0.0)(1316, 0.0)(1317, 0.0)(1318, 0.0)(1319, 0.0)(1320, 0.0)(1321, 0.0)(1322, 0.0)(1323, 0.0)(1324, 0.0)(1325, 0.0)(1326, 0.0)(1327, 0.0)(1328, 0.0)(1329, 0.0)(1330, 0.0)(1331, 0.0)(1332, 0.0)(1333, 0.0)(1334, 0.0)(1335, 0.0)(1336, 0.0)(1337, 0.0)(1338, 0.0)(1339, 0.0)(1340, 0.0)(1341, 0.0)(1342, 0.0)(1343, 0.0)(1344, 0.0)(1345, 0.0)(1346, 0.0)(1347, 0.0)(1348, 0.0)(1349, 0.0)(1350, 0.0)(1351, 0.0)(1352, 0.0)(1353, 0.0)(1354, 0.0)(1355, 0.0)(1356, 0.0)(1357, 0.0)(1358, 0.0)(1359, 0.0)(1360, 0.0)(1361, 0.0)(1362, 0.0)(1363, 0.0)(1364, 0.0)(1365, 0.0)(1366, 0.0)(1367, 0.0)(1368, 0.0)(1369, 0.0)(1370, 0.0)(1371, 0.0)(1372, 0.0)(1373, 0.0)(1374, 0.0)(1375, 0.0)(1376, 0.0)(1377, 0.0)(1378, 0.0)(1379, 0.0)(1380, 0.0)(1381, 0.0)(1382, 0.0)(1383, 0.0)(1384, 0.0)(1385, 0.0)(1386, 0.0)(1387, 0.0)(1388, 0.0)(1389, 0.0)(1390, 0.0)(1391, 0.0)(1392, 0.0)(1393, 0.0)(1394, 0.0)(1395, 0.0)(1396, 0.0)(1397, 0.0)(1398, 0.0)(1399, 0.0)(1400, 0.0)(1401, 0.0)(1402, 0.0)(1403, 0.0)(1404, 0.0)(1405, 0.0)(1406, 0.0)(1407, 0.0)(1408, 0.0)(1409, 0.0)(1410, 0.0)(1411, 0.0)(1412, 0.0)(1413, 0.0)(1414, 0.0)(1415, 0.0)(1416, 0.0)(1417, 0.0)(1418, 0.0)(1419, 0.0)(1420, 0.0)(1421, 0.0)(1422, 0.0)(1423, 0.0)(1424, 0.0)(1425, 0.0)(1426, 0.0)(1427, 0.0)(1428, 0.0)(1429, 0.0)(1430, 0.0)(1431, 0.0)(1432, 0.0)(1433, 0.0)(1434, 0.0)(1435, 0.0)(1436, 0.0)(1437, 0.0)(1438, 0.0)(1439, 0.0)(1440, 0.0)(1441, 0.0)(1442, 0.0)(1443, 0.0)(1444, 0.0)(1445, 0.0)(1446, 0.0)(1447, 0.0)(1448, 0.0)(1449, 0.0)(1450, 0.0)(1451, 0.0)(1452, 0.0)(1453, 0.0)(1454, 0.0)(1455, 0.0)(1456, 0.0)(1457, 0.0)(1458, 0.0)(1459, 0.0)(1460, 0.0)(1461, 0.0)(1462, 0.0)(1463, 0.0)(1464, 0.0)(1465, 0.0)(1466, 0.0)(1467, 0.0)(1468, 0.0)(1469, 0.0)(1470, 0.0)(1471, 0.0)(1472, 0.0)(1473, 0.0)(1474, 0.0)(1475, 0.0)(1476, 0.0)(1477, 0.0)(1478, 0.0)(1479, 0.0)(1480, 0.0)(1481, 0.0)(1482, 0.0)(1483, 0.0)(1484, 0.0)(1485, 0.0)(1486, 0.0)(1487, 0.0)(1488, 0.0)(1489, 0.0)(1490, 0.0)(1491, 0.0)(1492, 0.0)(1493, 0.0)(1494, 0.0)(1495, 0.0)(1496, 0.0)(1497, 0.0)(1498, 0.0)(1499, 0.0)(1500, 0.0)(1501, 0.0)(1502, 0.0)(1503, 0.0)(1504, 0.0)(1505, 0.0)(1506, 0.0)(1507, 0.0)(1508, 0.0)(1509, 0.0)(1510, 0.0)(1511, 0.0)(1512, 0.0)(1513, 0.0)(1514, 0.0)(1515, 0.0)(1516, 0.0)(1517, 0.0)(1518, 0.0)(1519, 0.0)};

            \addplot+[name path=ZZZ,red,line width=.3pt] coordinates {};

            \addplot[blue!50,fill opacity=0.2] fill between[of=A and B];
            \addplot[blue!50,fill opacity=0.2] fill between[of=A and C];
            \addplot[blue!50,fill opacity=0.2] fill between[of=A and D];
            \addplot[blue!50,fill opacity=0.2] fill between[of=A and E];
            \end{axis}
        \end{tikzpicture}
        \end{document}