### 2021-01-05 paper/paper.tex
Recompile paper with new figures

\documentclass[letterpaper]{article}

\usepackage{natbib,alifeconf}  %% The order is important
\usepackage{graphicx}%
\usepackage{blindtext}
\usepackage{subfig}

% *****************
%  Requirements:
% *****************
%
% - All pages sized consistently at 8.5 x 11 inches (US letter size).
% - PDF length <= 8 pages for full papers, <=2 pages for extended
%    abstracts (not including citations).
% - Abstract length <= 250 words.
% - No visible crop marks.
% - Images at no greater than 300 dpi, scaled at 100%.
% - Embedded open type fonts only.
% - All layers flattened.
% - No attachments.
% - All desired links active in the files.

% Note that the PDF file must not exceed 5 MB if it is to be indexed
% by Google Scholar. Additional information about Google Scholar
% can be found here:
% http://www.google.com/intl/en/scholar/inclusion.html.

% If your system does not generate letter format documents by default,
% you can use the following workflow:
% latex example
% bibtex example
% latex example ; latex example
% dvips -o example.ps -t letterSize example.dvi
% ps2pdf example.ps example.pdf

% For pdflatex users:
% The alifeconf style file loads the "graphicx" package, and
% this may lead some users of pdflatex to experience problems.
% These can be fixed by editing the alifeconf.sty file to specify:
% \usepackage[pdftex]{graphicx}
%   instead of
% \usepackage{graphicx}.
% The PDF output generated by pdflatex should match the required
% specifications and obviously the dvips and ps2pdf steps become
% unnecessary.

% Note:  Some laser printers have a serious problem printing TeX
% output. The use of ps type I fonts should avoid this problem.

\title{Towards Ecosystem Management from Greedy Reinforcement Learning in a Predator-Prey Setting}
\title{Emergent Cooperation from Greedy Multi-Agent Reinforcement Learning in a Predator-Prey Setting}
\author{Daniel Ratke$^{2}$, Fabian Ritz$^{1}$, Thomy Phan$^1$ \and Claudia Linnhoff-Popien$^1$ \\
\mbox{}\\
$^1$Mobile and Distributed Systems Group, LMU Munich, Germany \\
$^2$LMU Munich, Germany \\
D.Ratke@campus.lmu.de} % email of corresponding author

% For several authors from the same institution use the same number to
% refer to one address.
%
% If the names do not fit well on one line use
%         Author 1, Author 2 ... \\ {\Large\bf Author n} ...\\ ...
%
% If the title and author information do not fit in the area
% allocated, place \setlength\titlebox{<new height>} after the
% \documentclass line where <new height> is 2.25in

\begin{document}
\maketitle

\begin{abstract}
% Abstract length should not exceed 250 words
    \blindtext[2]
\end{abstract}

\section{Introduction}

\blindtext[4]

\begin{figure}%
    \centering
    \subfloat[][\centering Two predators and multiple prey]{\includegraphics[width=3.8cm]{figures/game_screen2.png} }%
    \subfloat[][\centering A stunned predator]{\includegraphics[width=3.66cm]{figures/stun_move2.png} }%
    \caption{The predator-prey environment, with yellow predators and blue prey. Predators can stun each other.}%
    \label{fig:aquarium}%
\end{figure}

\section{Background}

\subsection{Reinforcement Learning}

\blindtext[2]

\subsection{Multi-Agent Reinforcement Learning}

\blindtext[3]

\subsection{Herding Behaviour}

\blindtext[2]

\subsection{Cooperation}

\blindtext[2]

\subsection{Related Work}

\blindtext[5]

\section{Aquarium Environment}

\blindtext[5]

\section{Experimental Setup}

\blindtext[2]

\section{Results}

\begin{figure}[t]
\begin{center}
\includegraphics[width=8cm]{figures/dqncomp1fish_2.png}
\caption{Sustainability of static, static-wait, DQN Stage II and PPO with initial fish population of 1.}
\label{fig:dqncomp1}
\end{center}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=8cm]{figures/dqncomp2fish_2.png}
\caption{Sustainability of static, static-wait, DQN Stage II and PPO with initial fish population of 2.}
\label{fig:dqncomp2}
\end{center}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=8cm]{figures/fish_pop_chart3.png}
\caption{Left-over fish population per learning episode over 12M steps.}
\label{fig:fish_pop}
\end{center}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=6cm]{figures/shark_speeds.png}
\caption{Distribution of shark speeds over time.}
\label{fig:shark_speeds}
\end{center}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=8cm]{figures/Sterberisiko_vs_Herding_Two_Net_2.png}
\caption{TODO : Placeholder Chart (Idea: Three distributions with fish pop i10,i5,i2)}
\label{fig:herding}
\end{center}
\end{figure}

\begin{figure*}[t]
\begin{center}
\includegraphics[width=\textwidth]{figures/coop_movie_frames2.png}
\caption{From left to right: Two sharks cooperatively catch a fish.}
\label{fig:coop_movie}
\end{center}
\end{figure*}

\begin{figure*}%
    \centering
    \subfloat[][\centering Cooperation rate with initial fish population of 10.]{\includegraphics[width=5.5cm]{figures/coop_chart_split1.png} }%
    \subfloat[][\centering Cooperation rate with initial fish population of 5.]{\includegraphics[width=5.5cm]{figures/coop_chart_split2.png} }%
    \subfloat[][\centering Cooperation rate with initial fish population of 5 and stuns.]{\includegraphics[width=6.5cm]{figures/coop_chart_stuns_split2.png} }%
    \caption{Cooperation rate in different scenarios.}%
    \label{fig:coop}%
\end{figure*}

\begin{figure*}[t]
\begin{center}
\includegraphics[width=14cm]{figures/coop_vs_sterberisiko_5fish.png}
\caption{Cooperation rate vs starving pressure with initial fish population of 5.}
\label{fig:coop_starve}
\end{center}
\end{figure*}

\blindtext[14]

\section{Conclusion}

\blindtext[2]

\footnotesize
\bibliographystyle{apalike}
\bibliography{paper} % replace by the name of your .bib file

\end{document}